### 2016-03-01 JSS/jss-style/article.tex
Initial commit

\documentclass[article,shortnames]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Lindsay Rutter\\Iowa State University \And 
        Susan VanderPlas\\Iowa State University \And
        Di Cook\\Monash University}
\title{\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Lindsay Rutter, Susan VanderPlas, Di Cook} %% comma-separated
\Plaintitle{ggenealogy: An R Package for Visualizing Genealogical Data} %% without formatting
\Shorttitle{\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, and displaying the results. It is possible to draw the genealogy in relation to variables related to the nodes, and to determine and display the shortest path distances between genetic lines. Production of pairwise distance matrices and genealogical diagrams constrained on generation are also available in the visualization toolkit. The software is being tested on datasets with milestone cultivars of soybean (\citealt{soybean}) and barley varieties [2], as well as on data from the \citealt{mgp}, a database of the academic genealogy of mathematicians. It is currently available on the Comprehensive R Archive Network.}

\Keywords{genealogy, data visualization, statistical graphics, exploratory data analysis, interactive, \proglang{R}}
\Plainkeywords{genealogy, data visualization, statistical graphics, exploratory data analysis, interactive, R} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Lindsay Rutter\\
  Bioinformatics and Computational Biology Program\\
  Iowa State University\\
  2014 Molecular Biology Building\\
  Ames, IA, 50011, United States of America\\
  E-mail: \email{lrutter@iastate.edu}\\
  URL: \url{https://github.com/lrutter/}\\

  Susan VanderPlas\\
  Department of Statistics\\
  Iowa State University\\
  2413 Snedecor Hall\\
  Ames, IA, 50011, United States of America\\
  E-mail: \email{srvanderplas@gmail.com}\\
  URL: \url{https://github.com/srvanderplas/}\\

  Di Cook\\
  Department of Econometrics and Business Statistics\\
  Monash University\\
  E869 Menzies Building\\
  20 Chancellors Walk\\
  Clayton, VIC 3800, Australia\\
  E-mail: \email{dicook@monash.edu}\\
  URL: \url{https://github.com/dicook/}
}

%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%% \section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\section{Introduction}

Genealogy is the study of the parent-child relationships. By tracing through lineages of groups of individuals, the history of features that have been modified over time can be studied. Comparative geneticists, computational biologists, and bioinformaticians commonly use these tools to better understand the historical changes that caused novel and desirable traits to arise in lineages. For example, in crops, desirable modifications could include an increase in protein yield or an increase in disease resistance. However, there are also times when lineages of detrimental traits are analyzed, such as to determine the origin of hazardous traits in rapidly-evolving viruses.

Furthermore, genealogical relationships can be applied outside of a strict biological sense, an example being the \citealt{mgp}, which maintains the family lineage of academic mathematicians, with documentation of doctoral advisors, doctoral students, and graduation years of all mathematicians in academia. Such family lineages allow us to understand the position of one member in the larger historical picture, and to accurately preserve past relationships for the knowledge of future generations.

In all these examples, the data structures containing the genealogical relationships can be represented visually. Access to various types of visual plots and diagrams of the lineage can allow scientists and others to more efficiently and accurately explore an otherwise complicated data structure. We introduce here a developing visualization toolkit that is intended to assist users in their exploration and analysis of genealogical relationships. In this paper, we will demonstrate some of the main features of this software package, and summarize any novelty that such features may provide users, using an example lineage dataset of soybean cultivars (\citealt{soybean}).

\section{Available Software}

Publishing in the open source \proglang{R} statistical programming language allows for tools to be distributed and modified at ease, encourages cross-platform collaboration, and provides a foundation for effective and aesthetic data visualization from the grammar of graphics. There are several useful \proglang{R} packages that offer tools for analyzing and visualizing genealogical datasets. Here, we introduce these packages, and emphasize the new features that \proglang{ggenealogy} brings to this collection of work.

The \proglang{R} package \proglang{pedigree} is named after the standardized chart used by genealogists to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}). The package uses a similar data input type as \proglang{ggenealogy}, namely a data frame for parent and child labels. While it does provide tools to perform methods on the data set, such as rapidly determining the generation count for each member in the pedigree, it does not provide any visualization tools.

Another \proglang{R} package called \proglang{kinship2} does produce basic pedigree charts, one from the package vignette of which is provided as an example in Figure \ref{fig:kinshipFig} (\citealt{kin}). This pedigree chart adheres to the standard notations used for visualizing genealogical structures. Even though doing so creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational cross-breeding occurs, as is often the case in agronomical genealogical lineages. We demonstrate how the standardized pedigree charts in the \proglang{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational cross-breeding case in Figure \ref{fig:kinshipFig}. We then later illustrate how new plotting tools in the \proglang{ggenealogy} package resolve this previously unaddressed area in Figure \ref{fig:Lee}.

In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures, as is shown in Figure \ref{fig:Graph} (\citealt{graphvizCit}, \citealt{cytoscapeCit}). These programs allow for visualizations of graphs, which are defined as objects with sets of nodes and edges, where nodes cannot be repeated. However, as is described in Figure \ref{fig:Lee}, for certain genealogical datasets to be non-ambiguously displayed (as is described in ), Hence, as is the case with the aforementioned \proglang{R} packages, these popular software also cannot

\begin{figure}[h]
    \centering
    \includegraphics[width=0.65\textwidth]{kinshipFig}
    \caption{Example pedigree chart output from the \proglang{kinship2} package. A standardized set of symbols are used to represent males with squares and females with circles. Parents are connected to each other by horizontal lines, and to their children by vertical lines. Siblings are connected by horizontal sibship lines. Each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202. We superimposed green-highlighted individual 215 and its corresponding connections onto the pedigree chart for explanatory purposes. Its parents are individuals 201 and 206, which are from generations one and two, respectively, and have a parent-child relationship themselves. As an offspring of a parent-child relationship, individual 215 is both a second generation and third generation individual. Hence, individual 215 should be displayed in both generational positions on the vertical axis. However, standard pedigree tools only allow for an individual to be displayed once. As a result, in cases where inter-generational cross-breading occurs, such as in agronomical applications, standardized tools for visualizing genealogical information cannot non-ambiguously portray the genealogical dataset.}
    \label{fig:kinshipFig}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{Graph}
    \caption{Example genealogical display using popular software like \proglang{GraphViz} and \proglang{Cytoscape}. The generation count of an individual node is constrained by its position on the vertical axis. As was shown in Figure \ref{fig:kinshipFig}, here too we superimpose a green node that has parents from the first and second generations. As an offspring of parents from two different generations, this green node is both a second generation and third generation individual of the original parent. Hence, it should be displayed in both corresponding generation positions on the vertical axis. However, standard graph visualization tools only allow for a given ode to be displayed once. As a result, this green node must be ambiguously positioned as either a second or third generation individual; here, it is positioned as a third generation individual.}
    \label{fig:Graph}
\end{figure}

\section{Demonstration Dataset}

The available data used in the current demonstration of the software is a data frame structure that contains 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties. In total, there are 230 unique soybean varieties present. These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s. They also contain information on the developmental years, as well as the copy number variants, single nucleotide polymorphisms, protein content, and yield, of each of the soybeans.

In this context, the software could ideally be used by bioinformaticians, geneticists, and agronomists who wish to study how soybean varieties are related. By referencing the visualization of the genealogical tree, these scientists may better understand genetic testing results - in this particular dataset, in terms of copy number variants, single nucleotide polymorphisms, protein content, and yield - and use that knowledge in future breeding sessions.

\section{Generating a Graphical Object}

Most functions in the software package require an input parameter of a graph structure. Therefore, as a preprocessing step, we must first convert our original data frame structure into a graph structure. Below, we read in the \proglang{R} data file \code{sbGeneal} that is included in the package as a sample data set of soybean genealogy.

\begin{CodeChunk}
\begin{CodeInput}
R> library("ggenealogy")
R> data(sbGeneal)
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	412 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}

We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{ig} using the function \code{dfToIG()}.

\begin{CodeChunk}
\begin{CodeInput}
R> ig <- dfToIG(sbGeneal)
R> ig
\end{CodeInput}
\begin{CodeOutput}
IGRAPH UNW- 230 340 -- 
+ attr: name (v/c), weight (e/n)
+ edges (vertex names):
 [1] 5601T    --Hutcheson        Adams    --Dunfield        
 [3] A.K.     --A.K. (Harrow)    Altona   --Flambeau        
 [5] Amcor    --Amsoy 71         Adams    --Amsoy           
 [7] Amsoy 71 --C1253            Anderson --Lincoln         
 [9] Bay      --York             Bedford  --Forrest         
[11] Beeson   --Kent             Blackhawk--Richland        
[13] Bonus    --C1266R           Bradley  --J74-39          
[15] Bragg    --Jackson          Bragg    --Bragg x D60-7965
+ ... omitted several edges
\end{CodeOutput}
\end{CodeChunk}

There are many statistics about the \code{sbGeneal} genealogical dataset that we may wish to know that cannot easily be obtained through images and tables. The package function \code{getBasicStatistics()} can be called, using the \code{ig} object as input. This will return a list of common graph theoretical measurements regarding the genealogical graph structure. For instance, is the whole structure connected? If not, how many separated components does it contain? In addition to these statistics, the \code{getBasicStatistics()} function will also return the number of nodes, the number of edges, the average path length, the graph diameter, and other graph theoretical information.

\begin{CodeChunk}
\begin{CodeInput}
R> getBasicStatistics(ig)
\end{CodeInput}
\begin{CodeOutput}
$isConnected
[1] FALSE

$numComponents
[1] 11

$avePathLength
[1] 5.333746

$graphDiameter
[1] 13

$numNodes
[1] 230

$numEdges
[1] 340

$logN
[1] 5.438079
\end{CodeOutput}
\end{CodeChunk}

\section{Plotting a Shortest Path}

As this data set deals with soy bean lineages, it may be useful for agronomists to track how two varieties are related to each other via parent-child relationships. Then, any dramatic changes in protein yield, SNP varieties, and other measures of interest between the two varieties can be traced across their genetic timeline, and pinpointed to certain varieties within their historical lineage. The \pkg{ggenealogy} software allows users to select two varieties of interest, and determine the shortest pathway of parent-child relationships between them, using the \code{getPath()} function. This will return a list path that contains the variety names and their years in the path.

\begin{CodeChunk}
\begin{CodeInput}
R> pathTN <- getPath("Tokyo", "Narow", ig, sbGeneal)
R> pathTN
\end{CodeInput}
\begin{CodeOutput}
$pathVertices
[1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"

$yearVertices
[1] "1907"   "1942"   "1954.5" "1971.5" "1985" 
\end{CodeOutput}
\end{CodeChunk}

The returned path object can then be plotted using the \code{plotPath()} function, as demonstrated below.

\begin{Code}
R> plotPath(pathTN)
\end{Code}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{pathTNZB}
    \caption{Left: The shortest path between varieties Tokyo and Narow is strictly composed of a unidirectional sequence of parent-child relationships. Right: The shortest path between varieties Zane and Bedford is not strictly composed of unidirectional parent-child relationships, but have a cousin-like relationship.}
    \label{fig:pathTNZB}
\end{figure}

This produces a visual that informs users of all the varieties involved in the shortest path between the two varieties of interest, see left half of Figure \ref{fig:pathTNZB}. In this plot, the years of all varieties involved in the path are indicated on the horizontal axis, while the vertical axis has no meaning other than to simply to display the labels evenly spaced vertically. The shortest path between varieties \code{Tokyo} and \code{Narow} is composed of a unidirectional series of parent-child relationships, with \code{Tokyo} as the starting ancestor in the early 1900s, \code{Narow} as the most recent descendent in the mid 1980s, and three varieties in between.

Next, we can run the same set of functions on a different pair of varieties. First, a call to the \pkg{ggenealogy} function \code{getYear()} indicates that variety \code{Bedford} was developed in 1978 and variety \code{Zane} in 1985.

\begin{CodeChunk}
\begin{CodeInput}
R> getYear("Bedford", sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 1978
\end{CodeOutput}
\begin{CodeInput}
R> getYear("Zane", sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 1985
\end{CodeOutput}
\end{CodeChunk}

We can then create a plot showing the shortest path between these two varieties of interest.

\begin{Code}
R> pathZB <- getPath("Zane", "Bedford", ig, sbGeneal)
R> plotPath(pathZB)
\end{Code}

The resulting plot (right half of Figure \ref{fig:pathTNZB}) allows us to quickly determine that \code{Bedford} is not a parent, grandparent, or any great grandparent of \code{Zane}. Instead, we see that these two varieties are not related through a unidirectional parent-child lineage, but have a cousin-like relationship. The oldest common ancestor between \code{Zane} and \code{Bedford} is the variety \code{D55-4090}, which was developed in the mid 1940s.

Furthermore, as determined by the figure, for both \code{Zane} and \code{Bedford}, there are four varieties of unidirectional parent-child relationships between each of them and their common ancestor \code{D55-4090}. Hence, any parameter of interest that differentiates \code{Zane} and \code{Bedford} (protein yield, disease resistance, etc.) can also be examined across these two separate lineage histories.

\section{Superimposing Shortest Path on Tree}

Now that we can create path objects, we may wish to know how those paths are positioned in comparison to the genealogical lineage of the entire data structure. For instance, of the documented soybean cultivar lineage varieties, where does the shortest path between two varieties of interest exist? Are these two varieties comparatively older compared to the overall data structure? Are they newer? Or, do they span the entire structure, and represent two extreme ends of documented time points?

There is a function available in the \pkg{ggenealogy} package \code{plotPathOnAll()} that can allow users to quickly visualize their path of interest superimposed over all varieties and edges present in the whole data structure. Here we will produce a plot of the previously-determined shortest path between varieties \code{Tokyo} and \code{Narow} across the entire dataset:

\begin{Code}
R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, ig, binVector = 1:3,
  pathColor = "red")
R> plotAllImage + ggplot2::theme(axis.text = element_text(size = 20),
  axis.title = element_text(size = 20))
R> plotAllImage <- plotPathOnAll(path = path, geneal = sbGeneal, ig = ig,
  binVector = sample(1:3), edgeCol = "gray84", pathEdgeCol = "blue", nodeSize
  = 1, pathNodeSize = 3)
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{fig3}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
    \label{fig:fig3}
\end{figure}

While the first three explicit parameters to the function \code{plotPathOnAll()} have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation. The motivation of the \code{plotPathOnAll()} is to write variety text labels on a plot, with the center of each variety label constricted on the horizontal axis to its developmental year. As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the text labels. Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have labels with large numbers of characters (width of x coordinate).

For each variety, the x coordinate (year) and width of the x coordinate (text label width) cannot be altered, as they provide useful information. However, for each variety, the y coordinate is arbitrary. Hence, in an attempt to mitigate text overlapping, the \code{plotPathOnAll()} does not randomly assign the y coordinate. Instead, it allows users to partially control the y coordinates with a user-determined number of bins (\code{binVector}).

If the user determines to produce a plot using three bins, as in the example code above, then the varieties are all grouped into three bins based on their years of development. In other words, there will be bin 1 (the ``oldest bin") which includes one-third of the total number of varieties all with the oldest developmental years, bin 2 (the ``middle bin"), and bin 3 (the ``youngest bin").

Then, in order to decrease text overlap, the consecutively increasing y-axis coordinates are alternatively assigned to the three bins (For example: bin 1, then bin 2, then bin 3, then bin 1, then bin 2, then bin 3, ...) repeatedly until all varieties are accounted for. This algorithm means that for any pair of varieties within a given bin constrained to those years on the horizontal axis, there are exactly two other varieties placed between them vertically on the y-axis that come from the two other bins 1 constrained to a different set of year values on the horizontal axis.

Hence, in the code above, the user selected a \code{binVector} value of three, and a \code{plotColor} of red, which produces the plot in Figure \ref{fig:fig3}. We see that edges not on the path of interest are thin and gray by default, whereas edges on the path of interest are bolded and red. We also see that varieties in the path of interest are boldfaced by default.

The plot presents useful information: We immediately gather that the path of interest between does span most of the years of the data structure. In fact, \code{Tokyo} appears to be the oldest variety present in the dataset, and \code{Narow} appears to be one of the youngest varieties. We can also determine that the vast majority of varieties appear to have development years between 1950 and 1970.

However, this plot has significant empty spaces between the noticeably distinct bins, whereas almost all text labels are overlapping, thereby decreasing their readability. To force some variety text labels into these spaces, the user may consider using a larger number of bins. Hence, we next examine a bin size of 6:

\begin{Code}
R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, ig, binVector = 1:6,
  pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{fig4}
    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
    \label{fig:fig4}
\end{figure}

This similar code now outputs the plot shown in Figure \ref{fig:fig4}. We can immediately see that this plot more successfully mitigates text variety label overlap than the previous plot in \ref{fig:fig3}. We can also confirm what we saw in the previous plot that indeed most varieties have development years between 1950 and 1970, and any textual overlap is confined to this range of years.

\section{Future Directions: Fine-tuning Superimposition}

This visualization tool is suitable as part of a data exploration phase, but not for any users seeking publication quality plots due to the remaining textual overlap. We continue to work towards further reducing this textual overlap, although it is impossible to guarantee no textual overlap, especially with larger datasets with dense subgroups of varieties having similar years.

As such, we plan to add a feature to the \pkg{ggenealogy} package that allows users to manually fine-tune the plot they deem best after examining various bin sizes. For example, after comparing several bin sizes (1-12) on the current soy bean data set, we determined that the bin size of 6 produced minimal textual overlap, as seen in Figure 4.

However, there still remained one dozen cases of partial textual overlap. As an example, the labels of \code{Crawford} (1974) and \code{Swift} (1973) overlap at just below the vertical midpoint of the plot. The proposed function would allow users to hardcode the label of either variety and manually assign it to a new vertical coordinate. For instance, a user might select \code{Swift} and slightly decrease its vertical coordinate so that it is drawn halfway between \code{Crawford} and \code{Bradley}.

If the user sequentially fine-tuned the vertical positions of overlapping text labels for the small fraction of labels that remained overlapped after the automated function, and if they monitored the progress by visually inspecting updated plots until there were no more overlaps, then the plot could be used in presentations and publications.

\section{Plotting Ancestors and Descendants by Generation}

The most novel visual tool in \pkg{ggenealogy}, \code{plotAncDes()}, allows users to view the ancestors and descendants of a given variety. The inputted variety is highlighted in the center of the plot, ancestors are displayed to the left of the center, and descendants to the right of the center. The further from the center, the larger the number of generations that particular ancestor/descendant is from the centered variety of interest.

This particular \pkg{ggenealogy} tool is unique because most available graphical software only consider simple graphs, with no repeated nodes. However, in some genealogical lineage datasets, some varieties must be repeated if they are to be visualized by generation counts. This was found to be the case in the soy bean dataset.

As an example, we will create a plot of the ancestors and descendants of the variety \code{Lee}. We specify that the maximum number of ancestor and descendant generations are both 6, and that the text of the variety of interest is highlighted in blue:

\begin{Code}
R> plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
\end{Code}

\begin{figure}%[h]
    \centering
    \includegraphics[width=\textwidth]{Lee}
    \caption{Left: All varieties within three generations of ancestors and five generations of descendants of the variety Lee are shown. The generation distance from the center variety is neatly displayed in this graph; the larger the generation distance from the center variety, the further left (for ancestors) or right (for descendants). For explanation purposes, all paths between Lee and 5601T are highlighted in blue. We see that 5601T appears four times in the plot, which is a unique and novel requirement provided by ggenealogy. Right: The paths that are highlighted in blue in the left plot from ggenealogy are shown here again, only now nodes cannot be repeated. We unsuccessfully try to constrain the horizontal position of the nodes by generation count, as was accomplished by the ggenealogy plot, without repeating nodes. The parenthetical number above each node represents the set of generation counts that node is away from the center node Lee; green ones indicate that the node could be successfully placed in one horizontal position, but red ones indicate that the node could not be successfully placed in only one horizontal position. Hence, without allowing nodes to repeat, this data information cannot be presented as it is in the ggenealogy graph on the left, and this is a current limitation in other currently-available graphical software that ggenealogy can now provide.}
    \label{fig:Lee}
\end{figure}

This generates the plot we see in the left side of Figure \ref{fig:Lee}. We notice that \code{Lee} has 3 generations of ancestors and 5 generations of descendants. However, we also notice that some varieties are repeated in the plot. For example, the variety \code{5601T} is represented four times - once as a third generation descendant of \code{Lee}, once as a fourth generation descendant of \code{Lee}, and twice as a fifth generation descendant of \code{Lee}.

This happens because there are various paths between \code{Lee} and \code{5601T} (see the right side of Figure \ref{fig:Lee}). Hence, if we are to present all ancestors and descendants of \code{Lee} - while constricting the horizontal axis to generational distance from \code{Lee} - then varieties like \code{5601T} must occur more than once in the plot.

One of the main motivations in developing the \code{plotAncDes()} function was an inability to find other similar software (\citealt{ape}) that could produce such a plot, where repeated nodes were permitted, see right side of Figure \ref{fig:Lee} for additional explanation. The \code{plotAncDes()} function generates plots that have increased readability, as it is relatively easy to compare how far varieties are from the center to obtain an idea of generational distance.

\section{Plotting Distance Matrix}

It may also be of interest to generate matrices where the colors indicates a variable (such as the degree of the shortest path) between all pairwise combinations of inputted varieties. The package \pkg{ggenealogy} also provides a function \code{plotDegMatrix()} for that purpose.

Here we generate a distance matrix for a set of 10 varieties, setting the x-label and y-label as ``Variety" and the legend label as ``Degree". Syntax from the \pkg{ggplot2} package (\citealt{ggplot2}) can be appended to the \code{plotDegMatrix()} function. In this example, we add \pkg{ggplot2} functionality to specify that pairs with small degrees are white, while those with large degrees are dark green, as well as to specify the label size of the legend title and text:

\begin{Code}
>R varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
  "Pella", "Tokyo", "Young", "Zane")
>R plotDegMatrix(varieties, ig, sbGeneal, "Variety", "Variety", "Degree") +
  ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
  legend.text = ggplot2::element_text(size = 15))
\end{Code}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{degMatrix}
    \caption{The shortest path degree matrix between ten varieties of interest.}
    \label{fig:degMatrix}
\end{figure}

This creates the plot in Figure \ref{fig:degMatrix}. We see that the degree of the shortest path between varieties \code{Bedford} and \code{Zane} is 10, which is consistent with what we saw earlier in Figure \ref{fig:pathTNZB}. However, we now also see that 10 may be a comparatively large degree in this same dataset.

\section{Conclusions}

The \pkg{ggenealogy} package offers various plotting tools that can assist those studying genealogical lineages in the data exploration phases, as well as in preparing publication-suitable images. As each plot comes with its pros and cons, we recommended for users to explore several visualization tools. If users are simultaneously using similar packages, we in particular recommend using the \code{plotAncDes()} function. This plot allows users to view generation counts of a variety of interest in a manner that is not as readily available in similar software packages (\citealt{ape}).

\section{Future Avenues}

Incorporation of the Shiny application (\citealt{shiny}) would allow users to examine \pkg{ggenealogy} tools in a more interactive way. The reactive programming would save them the time of using command-line for each change of input as well as the inefficiency of rerunning code. We also look forward to testing this package on additional genealogical data sets (\citealt{mgp}). Exploring several datasets with our software will allow us to fix remaining bugs, and provide us further insight into how to make our tools available for a flexible set of data inputs.

\section*{Acknowledgments} % Lindsay added

\bibliography{foo}

\end{document}

### 2016-03-05 JSS/jss-style/article.tex
MGP graphic

@@
 
-\section{Demonstration Dataset}
+\section{Soybean Genealogy Dataset}
 
-The available data used in the current demonstration of the software is a data frame structure that contains 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
-In total, there are 230 unique soybean varieties present.
+The \proglang{ggenealogy} package comes with two example datasets, one comprises a soybean genealogy and the other comprises an academic statistician genealogy.
+We will introduce both example datasets in this paper to demonstrate some of the tools available in the software.
+We start with the soybean genealogy, which is available in a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
+In total, there are 230 unique soybean varieties in the dataset.
 These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
@@
 
+\section{Academic Genealogy of Statisticians Dataset}
+
+In addition to the soybean genealogy, the \proglang{ggenealogy} package also comes with an academic genealogy of statisticians.
+To develop this later dataset, we contacted the Mathematics Genealogy Project, a web-based database for the genealogy of academic mathematicians.
+This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society.
+The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}, \citealt{sql}).
+
+Each entry in the database contained 26 variables pertaining to an individual who received a graduate-level academic degree in statistics.
+One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics").
+This process resulted in 8995 entries, which we further reduced by only retaining those entries that contained an advisor who was also in the field ``Statistics".
+For each of the remaining 3293 entries, we only maintained 6 of the variables.
+This process resulted in the second example dataset provided in the \proglang{ggenealogy} package
+
+This later example dataset is available in a data frame structure that contains 3293 rows, each which corresponds to a parent-child relationship, where the child is an individual who received a graduate-level academic degree in statistics.
+The parent is the advisor
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> data(statGeneal)
+R> dim(statGeneal)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 3293    6
+\end{CodeOutput}
+\begin{CodeInput}
+R> colnames(statGeneal)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "child"   "parent"  "year"    "country" "school"  "thesis"
+\end{CodeOutput}
+\end{CodeChunk}
+
+which is available in a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
+In total, there are 230 unique soybean varieties in the dataset.
+These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
+They also contain information on the developmental years, as well as the copy number variants, single nucleotide polymorphisms, protein content, and yield, of each of the soybeans.
+
 \section{Conclusions}
@@
 
+\begin{figure}%[h]
+    \centering
+    \includegraphics[width=\textwidth]{dCox.png}
+    \caption{plot7}
+    \label{fig:dCox}
+\end{figure}
+
 \section*{Acknowledgments} % Lindsay added
 
-\bibliography{foo}
+\bibliography{article}
 

### 2016-03-09 JSS/jss-style/article.tex
meeting update

@@
 
-We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{ig} using the function \code{dfToIG()}.
+We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function \code{dfToIG()}.
 
@@
 \begin{CodeInput}
-R> ig <- dfToIG(sbGeneal)
-R> ig
+R> sbIG <- dfToIG(sbGeneal)
+R> sbIG
 \end{CodeInput}
@@
 There are many statistics about the \code{sbGeneal} genealogical dataset that we may wish to know that cannot easily be obtained through images and tables.
-The package function \code{getBasicStatistics()} can be called, using the \code{ig} object as input.
+The package function \code{getBasicStatistics()} can be called, using the \code{sbIG} object as input.
 This will return a list of common graph theoretical measurements regarding the genealogical graph structure.
@@
 \begin{CodeInput}
-R> getBasicStatistics(ig)
+R> getBasicStatistics(sbIG)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> pathTN <- getPath("Tokyo", "Narow", ig, sbGeneal)
+R> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
 R> pathTN
@@
 \begin{Code}
-R> pathZB <- getPath("Zane", "Bedford", ig, sbGeneal)
-R> plotPath(pathZB)
+R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
+R> plotPath(pathBZ)
 \end{Code}
@@
 \begin{Code}
-R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, ig, binVector = 1:3,
+R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3,
   pathColor = "red")
-R> plotAllImage + ggplot2::theme(axis.text = element_text(size = 20),
-  axis.title = element_text(size = 20))
-R> plotAllImage <- plotPathOnAll(path = path, geneal = sbGeneal, ig = ig,
+R> plotAllImage ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+R> plotAllImage <- plotPathOnAll(path = path, geneal = sbGeneal, ig = sbIG
   binVector = sample(1:3), edgeCol = "gray84", pathEdgeCol = "blue", nodeSize
@@
     \centering
-    \includegraphics[width=\textwidth]{fig3}
+    \includegraphics[width=\textwidth]{plotTNBin3}
     \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
-    \label{fig:fig3}
+    \label{fig:plotTNBin3}
 \end{figure}
@@
 
-Hence, in the code above, the user selected a \code{binVector} value of three, and a \code{plotColor} of red, which produces the plot in Figure \ref{fig:fig3}.
+Hence, in the code above, the user selected a \code{binVector} value of three, and a \code{plotColor} of red, which produces the plot in Figure \ref{fig:plotTNBin3}.
 We see that edges not on the path of interest are thin and gray by default, whereas edges on the path of interest are bolded and red.
@@
 \begin{Code}
-R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, ig, binVector = 1:6,
+R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6,
   pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
@@
     \centering
-    \includegraphics[width=\textwidth]{fig4}
+    \includegraphics[width=\textwidth]{plotTNBin6}
     \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
-    \label{fig:fig4}
+    \label{fig:plotTNBin6}
 \end{figure}
 
-This similar code now outputs the plot shown in Figure \ref{fig:fig4}.
-We can immediately see that this plot more successfully mitigates text variety label overlap than the previous plot in \ref{fig:fig3}.
+This similar code now outputs the plot shown in Figure \ref{fig:plotTNBin6}.
+We can immediately see that this plot more successfully mitigates text variety label overlap than the previous plot in \ref{fig:plotTNBin3}.
 We can also confirm what we saw in the previous plot that indeed most varieties have development years between 1950 and 1970, and any textual overlap is confined to this range of years.
@@
 As such, we plan to add a feature to the \pkg{ggenealogy} package that allows users to manually fine-tune the plot they deem best after examining various bin sizes.
-For example, after comparing several bin sizes (1-12) on the current soy bean data set, we determined that the bin size of 6 produced minimal textual overlap, as seen in Figure 4.
+For example, after comparing several bin sizes (1-12) on the current soy bean data set, we determined that the bin size of 6 produced minimal textual overlap, as seen in \ref{fig:plotTNBin6}.
 
@@
   "Pella", "Tokyo", "Young", "Zane")
->R plotDegMatrix(varieties, ig, sbGeneal, "Variety", "Variety", "Degree") +
+>R plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") +
   ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{degMatrix}
+    \includegraphics[width=0.5\textwidth]{degMatrix}
     \caption{The shortest path degree matrix between ten varieties of interest.}
@@
 
-\section{Academic Genealogy of Statisticians Dataset}
+\section{Academic Genealogy of Statisticians}
 
@@
 This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society.
-The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}, \citealt{sql}).
+The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).
 
@@
 One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics").
-This process resulted in 8995 entries, which we further reduced by only retaining those entries that contained an advisor who was also in the field ``Statistics".
-For each of the remaining 3293 entries, we only maintained 6 of the variables.
-This process resulted in the second example dataset provided in the \proglang{ggenealogy} package
+Furthermore, we only retained those entries that, if they contained an advisor, then that advisor was also in the field ``Statistics".
+This process resulted in 8995 entries, which we reduced to 8166 entries by removing duplicate entries.
+With the final data frame of 8995 entries, we only maintained 6 of the variables.
 
-This later example dataset is available in a data frame structure that contains 3293 rows, each which corresponds to a parent-child relationship, where the child is an individual who received a graduate-level academic degree in statistics.
-The parent is the advisor
+This process resulted in the second example dataset provided in the \proglang{ggenealogy} package.
+This dataset is structured as a data frame that contains 8166 rows, each of which corresponds to a parent-child relationship, where both the child and the parent received postbaccalaureate degrees in statistics, and the parent was the academic advisor to the child.
+The columns of the dataset provide information about the year the child obtained the degree, the country and school from which the child obtained the degree, and the thesis title of the degree awarded to the child.
 
@@
 \begin{CodeOutput}
-[1] 3293    6
+[1] 8166    6
 \end{CodeOutput}
@@
 
-which is available in a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
-In total, there are 230 unique soybean varieties in the dataset.
-These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-They also contain information on the developmental years, as well as the copy number variants, single nucleotide polymorphisms, protein content, and yield, of each of the soybeans.
+%\section{Plotting Ancestors and Descendants by Generation}
+
+This second example dataset of academic genealogy is of a slightly different structure than the first example dataset of agronomical genealogy, because each child no longer requires exactly two parents.
+The ability to plot ancestors and descendants by generation was demonstrated on the agronomical genealogy in Figure \ref{fig:Lee}.
+As we believe this is the most novel plotting tool in the \proglang{ggenealogy} package, we can test it again here using the academic genealogy.
+
+We can determine which academic statistician in this \code{statGeneal} dataset has the largest number of descendants.
+To do so, we create a vector \code{indVec} that contains the names of all individuals (whether child or parent) in the dataset.
+We then use the \proglang{dplyr} package to apply the ggenealogy function \code{getDescendants} on each individual in the \code{indVec} vector (\citealt{dplyr}).
+We set the parameter \code{gen} to a conservatively large value of 100 as this dataset is unlikely to have any individuals with more than 100 generations of descendants.
+
+We then create a table to examine all values of descendant counts observed in the dataset, along with the number of individuals who had each of those values of descendant counts.
+Of the 8166 individuals in this dataset, 6252 of them have zero descendants, 322 of them have one descendant, and 145 of them have two descendants.
+There are only 17 individuals who have more than 30 descendants, and there is one individual who has the largest value of 159 descendants.
+We determine that this individual is the prominent British statistician Sir David Cox, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+>R library(dplyr)
+>R indVec <- unique(c(statGeneal$child, statGeneal$parent))
+>R indVec <- indVec[which(indVec != "", )]
+>R dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
+>R numDesc <- sapply(indVec, dFunc)
+>R table(numDesc)
+\end{CodeInput}
+\begin{CodeOutput}
+numDesc
+   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
+6252  322  145   88   58   36   31   22   23   14   17   13   14    9    9 
+  15   16   17   18   19   20   21   22   23   24   25   26   27   29   30 
+   6    4    3    2    5    7    5    3    3    2    2    6    1    1    3 
+  34   37   38   40   41   44   45   48   49   61   62   75   77   84  159 
+   2    1    1    1    1    1    1    1    2    1    1    1    1    1    1
+\end{CodeOutput}
+\begin{CodeInput}
+R> which(numDesc == 159)
+\end{CodeInput}
+\begin{CodeOutput}
+David Cox 
+     1980
+\end{CodeOutput}
+\end{CodeChunk}
+
+We can now visualize how these 159 descedants are related to Sir David Cox by plotting parent-child relationships, with generation count constricted on the horizontal axis, similar to what we performed in Figure \ref{fig:Lee}.
+We create this in Figure \ref{fig:dCox} using the code below.
+
+\begin{CodeInput}
+R> plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
+\end{CodeInput}
+
+\begin{figure}%[h]
+    \centering
+    \includegraphics[width=\textwidth]{dCox.png}
+    \caption{The 159 academic descendants of Sir David Cox.}
+    \label{fig:dCox}
+\end{figure}
+
+We see from Figure \ref{fig:dCox} that Sir David Cox had 42 children, many of them becoming notable statisticians, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
+The child of his who produced the most children of his own was Peter Bloomfield, who has 26 children and 49 descendants.
+In total, Sir David Cox has five generations of statistics students in this dataset.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> length(getChild("Peter Bloomfield", geneal))
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 26
+\end{CodeOutput}
+\begin{CodeInput}
+R> nrow(getDescendants("Peter Bloomfield", geneal, gen = 100))
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 49
+\end{CodeOutput}
+\end{CodeChunk}
+
+It would be insightful to examine a more detailed view of one of the longest strings of parent-child relationships between Sir David Cox and one of his fifth generation descendants.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> statIG <- dfToIG(statGeneal)
+R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected
+  = FALSE)
+R> plotPath(pathCB) + ggplot2::theme(axis.text = ggplot2::element_text(size
+  = 12), axis.title = ggplot2::element_text(size = 12)) +
+  ggplot2::scale_x_continuous(breaks = seq(1940, 2010, by = 10))
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{pathCB}
+    \caption{The shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional parent-child relationships that span about 55 years.
+We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than about 30 years in degree aquisition, whereas the three statisticians who earned their degrees later in this path only span less than about 10 years in degree acquisition.}
+    \label{fig:pathCB}
+\end{figure}
+
+\begin{figure}%[h]
+    \centering
+    \includegraphics[width=\textwidth]{plotCBText}
+    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
+    \label{fig:plotCBText}
+\end{figure}
+
+\begin{figure}%[h]
+    \centering
+    \includegraphics[width=\textwidth]{plotCBNoText}
+    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
+    \label{fig:plotCBNoText}
+\end{figure}
 
@@
 
-\begin{figure}%[h]
-    \centering
-    \includegraphics[width=\textwidth]{dCox.png}
-    \caption{plot7}
-    \label{fig:dCox}
-\end{figure}
-
 \section*{Acknowledgments} % Lindsay added

### 2016-03-09 JSS/jss-style/replicationScript.Rnw
meeting update

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission:\\
\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
\date{\today}
\begin{document}
\maketitle
\SweaveOpts{concordance=TRUE}
\tableofcontents
<<>>=
install.packages("ggenealogy")
library("ggenealogy")
@
%%%%%%%%%%%%%%%
%% Section 1 %%
%%%%%%%%%%%%%%%
\section{Generating a Graphical Object}

Examine structure of sbGeneal example dataset:

<<>>=
data(sbGeneal)
str(sbGeneal)
@

Convert sbGeneal into igraph object:

<<>>=
library("mstate")
tra2 <- !is.na(transMat(list(c(2), c(1))))
tra2
@
Use \texttt{mplskeleton} to construct a skeleton of the model parameter list \texttt{mpl}:
<<>>=
mpl <- mplskeleton(tmat = tra2)
str(mpl)
@
Parametrize the model and complete \texttt{mpl}:
<<>>=
mpl[[1]]$bhr[[2]] <- mpl[[2]]$bhr[[1]] <- function(t){return(0.5)}
## time-varying x2 and time-varying f(x2):
mpl[[1]]$eta[[2]] <- function(x.i, t){
  ifelse(t < 5, 
         return(1.0 * x.i[1] + 0.5 * x.i[2]), 
         return(1.0 * x.i[1] + 1.0 * x.i[3]))}
## time-varying x2 and time-varying f(x1):
mpl[[2]]$eta[[1]] <- function(x.i, t){
  ifelse(t < 5, 
         return(-0.5 * x.i[1] + 0.5 * x.i[2]), 
         return( 1.0 * x.i[1] + 0.5 * x.i[3]))}
@
Simulate event histories:
<<>>=
set.seed(123)
N <- 500
X <- matrix(nrow = N, ncol = 2, rnorm(2 * N))
X <- cbind(X, X[, 2] + runif(N)/10)
colnames(X) <- c("x1", "x2.t1", "x2.t2")
head(X)
Xstruc <- matrix(ncol = 2, nrow = 2, data = 0)
rownames(Xstruc) <- c("t1", "t2")
colnames(Xstruc) <- c("x1", "x2")
Xstruc[, 1] <- 1
Xstruc[, 2] <- c(2, 3)
Xstruc
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = c(5), Xstruc = Xstruc)
head(d$msm.basics)
head(d$ttsce[, 1:7])
head(d$tt.indicators)
@
Nelson-Aalen estimator of the cumulative hazard rate functions:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d = d$msm.basics, mpl = mpl, xlim = c(0, 15))
abline(a = 0, b = 0.5, col = "darkgrey")
@

Stratified partial likelihood modelling:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
m <- coxph(Surv(entry, exit, delta) ~ x1.12.t1 + x2.12.t1 + x1.12.t2 + 
             x2.12.t2 + x1.21.t1 + x2.21.t1 + x1.21.t2 + x2.21.t2 + 
             strata(trans), data = data.frame(d$msm.basics, d$ttsce))
round(summary(m)$coef[, c(1, 5)], 5)
plotcph(m)
points(c(c(1, 0.5, 1 - 1, 1 - 0.5, -0.5, 0.5, 1 - (-0.5), 0.5 - 0.5)), 
       pch = 16, col = "darkgrey")
@

Breslow estimate for cumulative hazard:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=be>>=
plotbe(m, mpl, xlim = c(0, 15), ylim = c(0, 10))
abline(a = 0, b = 0.5, col = "darkgrey")
@
%%%%%%%%%%%%%%%
%% Section 2 %%
%%%%%%%%%%%%%%%
\section{Illness-death Model with Recovery, a Non-linear Baseline Hazarad Rate Function and Non-linear Covariate Effects}
Prepare the simulation of event histories:
<<>>=
library("simMSM")
traIDM <- matrix(nrow = 3, ncol = 3, FALSE)
traIDM[1, 2] <- traIDM[1, 3] <- traIDM[2, 1] <- traIDM[2, 3] <- TRUE
## Or use mstate:
library("mstate")
traIDM <- !is.na(transMat(list(c(2, 3), c(1, 3), c())))
traIDM
## Build up mpl:
mpl <- mplskeleton(tmat = traIDM)
mpl[[1]]$bhr[[2]] <- function(t){0.5 + 0.4 * sin(t)}
mpl[[1]]$bhr[[3]] <- mpl[[2]]$bhr[[1]] <- 
  mpl[[2]]$bhr[[3]] <- function(t){0.5}
mpl[[1]]$eta[[2]] <- function(x.i, t){ 3 * (x.i^2)}
mpl[[1]]$eta[[3]] <- function(x.i, t){-3 * (x.i^2)}
mpl[[2]]$eta[[1]] <- mpl[[2]]$eta[[3]] <- 
  function(x.i, t){0.5 * x.i}
## simulate design matrix and event history data:
set.seed(123)
N <- 500
X <- matrix(ncol = 1, nrow = N, data = runif(N) - 0.5)
colnames(X) <- c("x")
Xstruc <- matrix(ncol = 1, nrow = 1, data = 1)
rownames(Xstruc) <- c("t1")
colnames(Xstruc) <- c("x")
Xstruc
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = NULL, Xstruc = Xstruc)
head(d$msm.basics)
@

Nelson-Aalen estimators:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d = d$msm.basics, mpl)
@

Convert to long data format and estimate a stratified Cox proportional hazards (partial likelihood) model:
<<>>=
ld <- tolongformat(d, mpl = mpl)
head(ld$msm.basics)
head(ld$tt.indicators)
m <- coxph(Surv(entry, exit, delta) ~ I(x.12^2) + I(x.13^2) + 
             x.21 + x.23 + strata(trans), 
           data = data.frame(ld$msm.basics, ld$ttsce))
round(summary(m)$coef[, c(1, 5)], 5)
@

Cox model coefficient illustration:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
plotcph(m)
points(c(3, -3, 0.5, 0.5), col = "darkgrey", pch = 16)
@

Breslow estimator:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotbe(m, mpl)
@

BayesX:
<<>>=
d <- data.frame(d$msm.basics, d$tt.indicators)
install.packages("BayesXsrc")
library("BayesXsrc")
p <- "delimiter = ;
dataset dat;
dat.infile using d.raw;
remlreg simmsm;
simmsm.mregress 
trans12=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
trans13=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
trans21=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
trans23=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2),
family=multistate lefttrunc=entry state=from maxit=1000 using dat;
delimiter = return;"
write.table(d, row.names = FALSE, quote = FALSE, file = "d.raw")
writeLines(p, con = "prg")
run.bayesx(prg = "prg", verbose = FALSE)
@

Prepare plotting the results:
<<>>=
setwd("./output/")
log.bhr <- beta.fixed <- beta.x <- vector("list", 4)
log.bhr[[1]] <- read.table("simmsm_f_exit_logbaseline.res", header=T)
log.bhr[[2]] <- read.table("simmsm_f_2_exit_logbaseline.res", header=T)
log.bhr[[3]] <- read.table("simmsm_f_3_exit_logbaseline.res", header=T)
log.bhr[[4]] <- read.table("simmsm_f_4_exit_logbaseline.res", header=T)
beta.fixed[[1]] <- read.table("simmsm_FixedEffects.res", header=T)
beta.fixed[[2]] <- read.table("simmsm_FixedEffects_2.res", header=T)
beta.fixed[[3]] <- read.table("simmsm_FixedEffects_3.res", header=T)
beta.fixed[[4]] <- read.table("simmsm_FixedEffects_4.res", header=T)
beta.x[[1]] <- read.table("simmsm_f_x_pspline.res", header=T)
beta.x[[2]] <- read.table("simmsm_f_2_x_pspline.res", header=T)
beta.x[[3]] <- read.table("simmsm_f_3_x_pspline.res", header=T)
beta.x[[4]] <- read.table("simmsm_f_4_x_pspline.res", header=T)
setwd("../")
plot.bhr <- function(log.bhr, beta.fixed, t.max){
  index.for.plot <- c("12,0", "13,0", "21,0", "23,0")
  par(mfrow = c(2, 2), mar = c(3.1, 3.1, 0.1, 0.1), tcl = -0.3)
  for(q in 1:4){
    plot(1, 1, ylim=c(0, 3), xlim=c(0, t.max), 
         type = "n", xaxt = "n", yaxt = "n", 
         xlab = "", ylab = "", bty = "n")
    axis(1, at = c(0, 4, 8, 12), label = FALSE)
    mtext(side = 1, text = c(0, 4, 8, 12), at = c(0, 4, 8, 12), 
          line = 0.5, cex = 0.5)
    axis(2, label = FALSE, at = 0:2)
    mtext(side = 2, text = 0:2, at = 0:2, line = 0.5, cex = 0.5)
    mtext("t", side = 1, line = 1.4, cex = 0.7)
    mtext(bquote(paste(lambda[.(index.for.plot[q])],(t))), side = 2, 
          line = 1.4, cex = 0.7)
    hilf <- beta.fixed[[q]]
    if(q == 1){
      ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <- hilf$pmode[1] - 0.255
    }
    if(q == 2){
      ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <- hilf$pmode[1] + 0.255
    }
    if(q %in% c(3, 4)){
      intercept <- hilf$pmode[1]
    }
    hilf <- log.bhr[[q]]
    polygon(c(hilf$exit, rev(hilf$exit)), 
            exp(intercept + c(hilf$ci95lower, rev(hilf$ci95upper))), 
            border = "NA", col = "lightgrey")
    lines(hilf$exit, exp(hilf$pmode + intercept), col = 1, lwd = 2)
    time <- seq(0, t.max, length = 100)
    if(q == 1){lines(time, (0.5 + 0.4 * sin(time)), lwd = 2, lty = 2)}
    if(q == 2){lines(time, (0.5 + 0*time), lwd = 2, lty = 2)}
    if(q == 3){lines(time, 0.5 + 0*time, lwd = 2, lty = 2)}
    if(q == 4){lines(time, 0.5 + 0*time, lwd = 2, lty = 2)}
  }
}
plot.spline <- function(beta.x){
  index.for.plot <- c("12,x", "13,x", "21,x", "23,x")
  par(mfrow = c(2, 2), mar = c(3.1, 3.1, 0.1, 0.1), tcl = -0.3)
  for(q in 1:4){
    plot(1, 1, ylim = c(-1.2, 1.2), xlim=c(-0.5, 0.5), type = "n", 
         xaxt = "n", yaxt = "n", xlab = "", ylab = "", bty = "n")
    axis(1, at = c(-0.5, 0, 0.5), label = FALSE)
    mtext(side = 1, text = c(-0.5, 0, 0.5), at = c(-0.5, 0, 0.5), 
          line = 0.5, cex = 0.5)
    axis(2, label = FALSE, at = -1:1)
    mtext(side = 2, text = -1:1, at = -1:1, line = 0.5, cex = 0.5)
    mtext("x", side = 1, line = 1.4, cex = 0.7)
    mtext(bquote(f[.(index.for.plot[q])](x)), side = 2, 
          line = 1.4, cex = 0.7)
    if(q == 1){
       ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <-  0.255
    }
    if(q == 2){
      ## -0.255 = mean(-3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <- -0.255
    }
    if(q %in% c(3, 4)){
      intercept <- 0
    }
    hilf <- beta.x[[q]]
    polygon(c(hilf$x, rev(hilf$x)), 
            intercept + c(hilf$ci95lower, rev(hilf$ci95upper)), 
            border = "NA", col = "lightgrey")
    lines(hilf$x, hilf$pmode + intercept, col = 1, lwd = 2)
    x <- seq(-0.5, 0.5, length = 100)
    if(q == 1){lines(x, 3*(x^2), lwd = 2, lty = 2)}
    if(q == 2){lines(x,-3*(x^2), lwd = 2, lty = 2)}
    if(q == 3){lines(x, 0.5 * x, lwd = 2, lty = 2)}
    if(q == 4){lines(x, 0.5 * x, lwd = 2, lty = 2)}
  }
}
@

BayesX baseline hazard rate function estimates:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=bayesX_bhr>>=
plot.bhr(log.bhr, beta.fixed, t.max = print(max(d$exit)))
@

BayesX covariate effect function estimates:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=bayesX_x>>=
plot.spline(beta.x)
@
%%%%%%%%%%%%%%%
%% Section 3 %%
%%%%%%%%%%%%%%%
\section{Illness-death Model with Partial Markov Influences}
Model build up in a general way:
<<>>=
traIDM <- matrix(nrow = 3, ncol = 3, FALSE)
traIDM[1, 2] <- traIDM[1, 3] <- traIDM[2, 1] <- traIDM[2, 3] <- TRUE
mpl <- mplskeleton(tmat = traIDM)
mpl[[1]]$bhr[[2]] <- mpl[[1]]$bhr[[3]] <- mpl[[2]]$bhr[[1]] <- 
                     mpl[[2]]$bhr[[3]] <- function(t){0.25}
mpl[[1]]$eta[[2]] <- mpl[[1]]$eta[[3]] <- mpl[[2]]$eta[[1]] <- 
  mpl[[2]]$eta[[3]] <- function(x.i, t){
    ifelse(t < 5, 
      return(0.5 * x.i[1]), 
      return(0.5 * x.i[2]))}
set.seed(123)
N <- 500
X <- matrix(nrow = N, ncol = 1, rnorm(N))
X <- cbind(X, X[, 1] + rnorm(N)/10)
colnames(X) <- c("x1.t1", "x1.t2")
Xstruc <- matrix(ncol = 1, nrow = 2, data = 0)
rownames(Xstruc) <- c("t1", "t2")
colnames(Xstruc) <- c("x1")
Xstruc[, 1] <- c(1, 2)
Xstruc
@
Now set-up the partial Markov influences. 
Function \texttt{partial.markov.x} has to take 5 input arguments representig vectors of past history information. 
They have to take names \texttt{entry}, \texttt{exit}, \texttt{from}, \texttt{to}, and \texttt{delta}:
<<>>=
partial.markov.x <- function(entry, exit, from, to, delta){
  count.12 <- sum(as.numeric((from == 1) & (to == 2) & (delta == 1)))
  count.21 <- sum(as.numeric((from == 2) & (to == 1) & (delta == 1)))
  return(c(count.12, count.21))}
@
The object \texttt{partial.markov.eta} is a list of lists in analogy to \texttt{mpl}:
<<>>=
partial.markov.eta <- pmeskeleton(traIDM)
str(partial.markov.eta)
partial.markov.eta[[1]][[2]] <- function(x){return( 0.25 * x[1])}
partial.markov.eta[[1]][[3]] <- function(x){return( 0.5 * x[1])}
partial.markov.eta[[2]][[1]] <- function(x){return(-0.5 * x[1] + 0.25 * x[2])}
partial.markov.eta[[2]][[3]] <- function(x){return(0)}
@
Event history simulation:
<<>>=
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = c(5), Xstruc = Xstruc, 
                       partial.markov.x = partial.markov.x, 
                       partial.markov.eta = partial.markov.eta)
@
Nelson-Aalen estimator of the cumulative hazard rate functions:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d$msm.basics, mpl = mpl)
abline(a = 0, b = 0.25, col = "darkgrey")
@

Stratified partial likelihood modelling:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
ld <- tolongformat(d = d, mpl = mpl)
head(ld$msm.basics, n = 10)
head(ld$ttsce[, 1:6], n = 10)
m <- coxph(Surv(entry, exit, delta) ~ x1 + 
             pm.x1.12 + pm.x1.13 + pm.x1.21 + pm.x2.21 + 
             strata(trans), 
           data = data.frame(ld$msm.basics, ld$ttsce))
plotcph(m)
points(c(0.5, 0.25, 0.5, -0.5, 0.25), pch = 16, col = "darkgrey")
@

Breslow estimate for cumulative hazard:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=be>>=
plotbe(m, mpl = mpl)
abline(a = 0, b = 0.25, col = "darkgrey")
@
\end{document}

### 2016-03-09 JSS/jss-style/replicationScript.tex
meeting update

--- 2016-03-09 JSS/jss-style/replicationScript.Rnw
@@
 \parindent 0pt
-\title{Replication Script to JSS Submission:\\
-\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
-\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
+\title{Replication Script to JSS Submission 1518:\\
+\textsl{Flexible Simulation of Event Histories: A~Guide for the \textsf{R}~Add-on~Package~\texttt{simMSM}}}
+\author{Holger Reulen\\ University of G\"ottingen}
 \date{\today}
@@
 <<>>=
-install.packages("ggenealogy")
-library("ggenealogy")
+options(prompt = " ", continue = "  ")
+install.packages("simMSM")
+library("simMSM")
 @
@@
 %%%%%%%%%%%%%%%
-\section{Generating a Graphical Object}
-
-Examine structure of sbGeneal example dataset:
-
+\section{Two State-type Model with Time-varying Effects and Time-change Covariates}
+Build transition-type definition matrix on your own:
 <<>>=
-data(sbGeneal)
-str(sbGeneal)
+tra2 <- matrix(ncol = 2, nrow = 2, data = FALSE)
+tra2[1, 2] <- tra2[2, 1] <- TRUE
+tra2
 @
-
-Convert sbGeneal into igraph object:
-
+Or use function \texttt{transMat} from R add-on package \texttt{mstate}:
 <<>>=

### 2016-03-09 JSS/jss-style/article.tex
acknowledgements and authors

--- 2016-03-05 JSS/jss-style/article.tex
@@
         Susan VanderPlas\\Iowa State University \And
-        Di Cook\\Monash University}
+        Di Cook\\Monash University\And
+        Michelle Graham\\Iowa State University}
 \title{\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data}
@@
   E-mail: \email{dicook@monash.edu}\\
-  URL: \url{https://github.com/dicook/}
+  URL: \url{https://github.com/dicook/}\\
+
+  Michelle Graham\\
+  USDA-Agriculture Research Service, Corn, Insects and Crop Genetics Research Unit\\
+  Department of Agronomy, Iowa State University\\
+  1575 Agronomy Building\\
+  Ames, IA, 50011, United States of America\\
+  E-mail: \email{magraham@iastate.edu}
 }
@@
 
+The authors are grateful for the financial support from the United Soybean Board, North Central Soybean Research Program, the Iowa Soybean Association, the NSF Plant Genome Research Program (award number 0820642), and the USDA-Agricultural Research Service.
+
 \bibliography{article}

### 2016-03-09 JSS/jss-style/article.tex
my full name

@@
         Susan VanderPlas\\Iowa State University \And
-        Di Cook\\Monash University\And
+        Dianne Cook\\Monash University\And
         Michelle Graham\\Iowa State University}
@@
 
-  Di Cook\\
+  Dianne Cook\\
   Department of Econometrics and Business Statistics\\

### 2016-03-10 JSS/jss-style/article.tex
Generalize intro outside of biological context

@@
 \Abstract{
-  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, and displaying the results.
-It is possible to draw the genealogy in relation to variables related to the nodes, and to determine and display the shortest path distances between genetic lines.
+  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} software package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, and displaying the results.
+It is possible to draw the genealogy in relation to variables related to the nodes, and to determine and display the shortest path distances between the nodes.
 Production of pairwise distance matrices and genealogical diagrams constrained on generation are also available in the visualization toolkit.
-The software is being tested on datasets with milestone cultivars of soybean (\citealt{soybean}) and barley varieties [2], as well as on data from the \citealt{mgp}, a database of the academic genealogy of mathematicians.
-It is currently available on the Comprehensive R Archive Network.}
+The tools are being tested on a dataset with milestone cultivars of soybean varieties (\citealt{soybean}) as well as on a web-based database of the academic genealogy of mathematicians (\citealt{mgp}).
+The software package is currently available on the Comprehensive R Archive Network.}
 
@@
 
-Genealogy is the study of the parent-child relationships.
-By tracing through lineages of groups of individuals, the history of features that have been modified over time can be studied.
-Comparative geneticists, computational biologists, and bioinformaticians commonly use these tools to better understand the historical changes that caused novel and desirable traits to arise in lineages.
-For example, in crops, desirable modifications could include an increase in protein yield or an increase in disease resistance.
-However, there are also times when lineages of detrimental traits are analyzed, such as to determine the origin of hazardous traits in rapidly-evolving viruses.
+Genealogy is the study of parent-child relationships.
+By tracing through parent-child lineages, genealogists can study the histories of features that have been modified over time.
+Comparative geneticists, computational biologists, and bioinformaticians commonly use genealogical tools to better understand the histories of novel traits arising across biological lineages.
+For example, desirable modifications in crops could include an increase in protein yield or an increase in disease resistance, and genealogical structures could be used to assess how these desirable traits developed.
+At the same time, genealogical lineages can also be used to assess detrimental features, such as to determine the origin of hazardous traits in rapidly-evolving viruses.
 
-Furthermore, genealogical relationships can be applied outside of a strict biological sense, an example being the \citealt{mgp}, which maintains the family lineage of academic mathematicians, with documentation of doctoral advisors, doctoral students, and graduation years of all mathematicians in academia.
-Such family lineages allow us to understand the position of one member in the larger historical picture, and to accurately preserve past relationships for the knowledge of future generations.
+Genealogical structures can also serve as informative tools outside of a strict biological sense.
+For instance, we can trace mentoring relationships between students and disseration supervisors with the use of academic genealogies.
+This can allow us to understand the position of one member in the larger historical picture, and to accurately preserve past relationships for the knowledge of future generations.
+Similarly, linguistic genealogies can be used to decipher the historical changes of vocabulary and grammatical features across related languages.
+In short, there is a diverse array of disciplines that can elicit useful information about features of interest by using genealogical data.
 
-In all these examples, the data structures containing the genealogical relationships can be represented visually.
-Access to various types of visual plots and diagrams of the lineage can allow scientists and others to more efficiently and accurately explore an otherwise complicated data structure.
-We introduce here a developing visualization toolkit that is intended to assist users in their exploration and analysis of genealogical relationships.
-In this paper, we will demonstrate some of the main features of this software package, and summarize any novelty that such features may provide users, using an example lineage dataset of soybean cultivars (\citealt{soybean}).
+In all these examples, the genealogical relationships can be represented visually.
+Access to various types of plotting tools can allow scientists and others to more efficiently and accurately explore features of interest across the genealogy.
+We introduce here a developing visualization toolkit that is intended to assist users in their exploration and analysis of genealogical structures.
+In this paper, we will demonstrate the main tools of this software package using two example genealogical datasets, one of soybean cultivars (\citealt{soybean}) and the other of academic mathematicians (\citealt{mgp}).
 

### 2016-03-10 JSS/jss-style/article.tex
Revamped intro

@@
 
-The \proglang{R} package \proglang{pedigree} is named after the standardized chart used by genealogists to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}).
-The package uses a similar data input type as \proglang{ggenealogy}, namely a data frame for parent and child labels.
-While it does provide tools to perform methods on the data set, such as rapidly determining the generation count for each member in the pedigree, it does not provide any visualization tools.
+The \proglang{R} package \proglang{pedigree} is named after the standardized chart used to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}).
+This package does provide tools that perform methods on parent-child datasets, such as rapidly determining the generation count for each member in the pedigree.
+However, it does not provide any visualization tools.
 
-Another \proglang{R} package called \proglang{kinship2} does produce basic pedigree charts, one from the package vignette of which is provided as an example in Figure \ref{fig:kinshipFig} (\citealt{kin}).
-This pedigree chart adheres to the standard notations used for visualizing genealogical structures.
-Even though doing so creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational cross-breeding occurs, as is often the case in agronomical genealogical lineages.
+Another \proglang{R} package called \proglang{kinship2} does produce basic pedigree charts (\citealt{kin}).
+In Figure \ref{fig:kinshipFig}, we provide an example pedigree chart from the \proglang{kinship2} package vignette.
+This pedigree chart adheres to the standard set of symbols used for visualizing genealogical structures: Males are represented with squares and females with circles.
+Parents are connected to each other by horizontal lines, and to their children by vertical lines.
+Siblings are connected by horizontal sibship lines.
+Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational cross-breeding occurs, as is often the case in agronomical genealogical lineages.
+
 We demonstrate how the standardized pedigree charts in the \proglang{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational cross-breeding case in Figure \ref{fig:kinshipFig}.
-We then later illustrate how new plotting tools in the \proglang{ggenealogy} package resolve this previously unaddressed area in Figure \ref{fig:Lee}.
+In that figure, each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202.
+We superimposed green-highlighted individual 215 onto the pedigree chart for explanatory purposes.
+Its parents are individuals 201 and 206, which are from generations one and two, respectively, and have a parent-child relationship between themselves.
+As an offspring of a parent-child relationship, individual 215 is both a second generation and third generation individual.
+Hence, individual 215 should be displayed in both second and third generational positions on the vertical axis.
+However, standard pedigree tools only allow for an individual to be displayed once.
+As a result, in special cases where inter-generational cross-breading occurs, such as in agronomical applications, standardized tools for visualizing genealogical information cannot non-ambiguously portray the genealogical dataset.
 
-In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures, as is shown in Figure \ref{fig:Graph} (\citealt{graphvizCit}, \citealt{cytoscapeCit}).
-These programs allow for visualizations of graphs, which are defined as objects with sets of nodes and edges, where nodes cannot be repeated.
-However, as is described in Figure \ref{fig:Lee}, for certain genealogical datasets to be non-ambiguously displayed (as is described in ), Hence, as is the case with the aforementioned \proglang{R} packages, these popular software also cannot
+In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}).
+Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated.
+In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational cross-breeding.
+We again illustrate this problem by superimposing an individual whose parents have a parent-child relationship in Figure \ref{fig:Graph}.
 
@@
     \includegraphics[width=0.65\textwidth]{kinshipFig}
-    \caption{Example pedigree chart output from the \proglang{kinship2} package.
-A standardized set of symbols are used to represent males with squares and females with circles.
-Parents are connected to each other by horizontal lines, and to their children by vertical lines.
-Siblings are connected by horizontal sibship lines.
-Each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202.
-We superimposed green-highlighted individual 215 and its corresponding connections onto the pedigree chart for explanatory purposes.
-Its parents are individuals 201 and 206, which are from generations one and two, respectively, and have a parent-child relationship themselves.
-As an offspring of a parent-child relationship, individual 215 is both a second generation and third generation individual.
-Hence, individual 215 should be displayed in both generational positions on the vertical axis.
-However, standard pedigree tools only allow for an individual to be displayed once.
-As a result, in cases where inter-generational cross-breading occurs, such as in agronomical applications, standardized tools for visualizing genealogical information cannot non-ambiguously portray the genealogical dataset.}
+    \caption{Example pedigree chart from the \proglang{kinship2} package, where the vertical axis denotes generation count.
+We superimposed green-highlighted individual 215 for explanatory purposes.
+As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual.
+Hence, it should be displayed twice on the vertical axis, once for each of its generation counts.
+However, standard pedigree tools only allow for an individual to be displayed once, leading to ambiguous portrayal of the genealogical dataset.}
     \label{fig:kinshipFig}
@@
     \includegraphics[width=0.6\textwidth]{Graph}
-    \caption{Example genealogical display using popular software like \proglang{GraphViz} and \proglang{Cytoscape}.
-The generation count of an individual node is constrained by its position on the vertical axis.
+    \caption{Example genealogical display using popular graph software like \proglang{GraphViz} and \proglang{Cytoscape}, with generation count denoted by the vertical axis.
 As was shown in Figure \ref{fig:kinshipFig}, here too we superimpose a green node that has parents from the first and second generations.
-As an offspring of parents from two different generations, this green node is both a second generation and third generation individual of the original parent.
-Hence, it should be displayed in both corresponding generation positions on the vertical axis.
-However, standard graph visualization tools only allow for a given ode to be displayed once.
+This green node is both a second and third generation individual of the original parent, and should be displayed in both corresponding generation positions on the vertical axis.
+However, standard graph visualization tools only allow for a given node to be displayed once.
 As a result, this green node must be ambiguously positioned as either a second or third generation individual; here, it is positioned as a third generation individual.}
@@
 
-\section{Soybean Genealogy Dataset}
+\section{Example Datasets}
 

### 2016-03-10 JSS/jss-style/article.tex
rownames misaligned in sbGeneal

@@
 We will introduce both example datasets in this paper to demonstrate some of the tools available in the software.
-We start with the soybean genealogy, which is available in a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
+
+\subsection{Soybean Genealogy}
+
+We start with the soybean genealogy, which is available as a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
 In total, there are 230 unique soybean varieties in the dataset.
 These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-They also contain information on the developmental years, as well as the copy number variants, single nucleotide polymorphisms, protein content, and yield, of each of the soybeans.
+They also contain information on the developmental years, copy number variants, single nucleotide polymorphisms, and protein content and yield for each of the soybeans.
 
-In this context, the software could ideally be used by bioinformaticians, geneticists, and agronomists who wish to study how soybean varieties are related.
-By referencing the visualization of the genealogical tree, these scientists may better understand genetic testing results - in this particular dataset, in terms of copy number variants, single nucleotide polymorphisms, protein content, and yield - and use that knowledge in future breeding sessions.
-
-\section{Generating a Graphical Object}
-
-Most functions in the software package require an input parameter of a graph structure.
-Therefore, as a preprocessing step, we must first convert our original data frame structure into a graph structure.
-Below, we read in the \proglang{R} data file \code{sbGeneal} that is included in the package as a sample data set of soybean genealogy.
+In this context, the software could ideally be used by agronomists who wish to study how soybean varieties are related.
+By referencing the visualization of the genealogical structure, these scientists may better understand genetic testing results - in this particular dataset, in terms of copy number variants, single nucleotide polymorphisms, and protein content and yield - and use that knowledge in future breeding sessions.
 
@@
 \end{CodeChunk}
+
+\subsection{Academic Genealogy of Statisticians}
+
+In addition to the soybean genealogy, the \proglang{ggenealogy} package also comes with an academic genealogy of statisticians.
+To develop this later dataset, we contacted the Mathematics Genealogy Project, a web-based database for the genealogy of academic mathematicians.
+This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society.
+The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).
+
+Each entry in the database contained 26 variables pertaining to an individual who received a graduate-level academic degree in mathematics.
+One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics").
+Furthermore, we only retained entries that had an advisor if that advisor was also in the field of ``Statistics".
+This process resulted in 8995 entries, which we reduced to 8166 entries by removing duplicate entries.
+With the final data frame of 8995 entries, we only maintained 6 of the original 26 variables.
+
+This process resulted in the second example dataset provided in the \proglang{ggenealogy} package.
+This dataset is structured as a data frame that contains 8166 rows, each of which corresponds to a parent-child relationship, where both the child and the parent received postbaccalaureate degrees in statistics, and the parent was the academic advisor to the child.
+The columns of the dataset provide information about the year the child obtained the degree, the country and school from which the child obtained the degree, and the thesis title of the degree awarded to the child.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> data(statGeneal)
+R> dim(statGeneal)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 8166    6
+\end{CodeOutput}
+\begin{CodeInput}
+R> colnames(statGeneal)
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "child"   "parent"  "year"    "country" "school"  "thesis"
+\end{CodeOutput}
+\end{CodeChunk}
+
+\section{Generating a Graphical Object}
+
+Most functions in the software package require an input parameter of a graph structure.
+Therefore, as a preprocessing step, we must first convert our original data frame structure into a graph structure.
+Below, we read in the \proglang{R} data file \code{sbGeneal} that is included in the package as a sample data set of soybean genealogy.
 
@@
 
-In addition to the soybean genealogy, the \proglang{ggenealogy} package also comes with an academic genealogy of statisticians.
-To develop this later dataset, we contacted the Mathematics Genealogy Project, a web-based database for the genealogy of academic mathematicians.
-This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society.
-The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).
-
-Each entry in the database contained 26 variables pertaining to an individual who received a graduate-level academic degree in statistics.
-One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics").
-Furthermore, we only retained those entries that, if they contained an advisor, then that advisor was also in the field ``Statistics".
-This process resulted in 8995 entries, which we reduced to 8166 entries by removing duplicate entries.
-With the final data frame of 8995 entries, we only maintained 6 of the variables.
-
-This process resulted in the second example dataset provided in the \proglang{ggenealogy} package.
-This dataset is structured as a data frame that contains 8166 rows, each of which corresponds to a parent-child relationship, where both the child and the parent received postbaccalaureate degrees in statistics, and the parent was the academic advisor to the child.
-The columns of the dataset provide information about the year the child obtained the degree, the country and school from which the child obtained the degree, and the thesis title of the degree awarded to the child.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> data(statGeneal)
-R> dim(statGeneal)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] 8166    6
-\end{CodeOutput}
-\begin{CodeInput}
-R> colnames(statGeneal)
-\end{CodeInput}
-\begin{CodeOutput}
-[1] "child"   "parent"  "year"    "country" "school"  "thesis"
-\end{CodeOutput}
-\end{CodeChunk}
-
 %\section{Plotting Ancestors and Descendants by Generation}

### 2016-03-14 JSS/jss-style/article.tex
Latest edits

@@
 \documentclass[article,shortnames]{jss}
+\usepackage{float}
 
@@
 Genealogical structures can also serve as informative tools outside of a strict biological sense.
-For instance, we can trace mentoring relationships between students and disseration supervisors with the use of academic genealogies.
-This can allow us to understand the position of one member in the larger historical picture, and to accurately preserve past relationships for the knowledge of future generations.
+For instance, we can trace mentoring relationships between students and dissertation supervisors with the use of academic genealogies.
+This can allow us to understand the position of one member in the larger historical picture of academia, and to accurately preserve past relationships for the knowledge of future generations.
 Similarly, linguistic genealogies can be used to decipher the historical changes of vocabulary and grammatical features across related languages.
@@
 We introduce here a developing visualization toolkit that is intended to assist users in their exploration and analysis of genealogical structures.
-In this paper, we will demonstrate the main tools of this software package using two example genealogical datasets, one of soybean cultivars (\citealt{soybean}) and the other of academic mathematicians (\citealt{mgp}).
+In this paper, we will demonstrate the main tools of this software package \pkg{ggenealogy} using two example genealogical datasets, one of soybean cultivars (\citealt{soybean}) and the other of academic mathematicians (\citealt{mgp}).
 
@@
 There are several useful \proglang{R} packages that offer tools for analyzing and visualizing genealogical datasets.
-Here, we introduce these packages, and emphasize the new features that \proglang{ggenealogy} brings to this collection of work.
+Here, we introduce these packages, and emphasize the new features that \pkg{ggenealogy} brings to this collection of work.
 
@@
 Its parents are individuals 201 and 206, which are from generations one and two, respectively, and have a parent-child relationship between themselves.
-As an offspring of a parent-child relationship, individual 215 is both a second generation and third generation individual.
+As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual.
 Hence, individual 215 should be displayed in both second and third generational positions on the vertical axis.
@@
 
-In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}).
-Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated.
-In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational cross-breeding.
-We again illustrate this problem by superimposing an individual whose parents have a parent-child relationship in Figure \ref{fig:Graph}.
-
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
@@
 
-\begin{figure}[h]
+In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}).
+Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated.
+In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational cross-breeding.
+
+We again illustrate this problem in Figure \ref{fig:Graph} with an example genealogy using popular graph drawing software like \proglang{GraphViz} and \proglang{Cytoscape}.
+Here, generation count is denoted by the vertical axis.
+As was shown in Figure \ref{fig:kinshipFig}, here too we superimpose a green node that has parents from two different generations.
+This green node is both a second and third generation individual, and should be displayed in both corresponding generation positions on the vertical axis.
+However, standard graph visualization tools only allow for a given node to be displayed once.
+As a result, this green node must be ambiguously positioned in either the second or third generation position; in the figure, it is denoted as a third generation individual.
+In Section \ref{remedy}, we will demonstrate \pkg{ggenealogy} plots that can remedy these problems.
+
+\begin{figure}[H]
     \centering
@@
     \caption{Example genealogical display using popular graph software like \proglang{GraphViz} and \proglang{Cytoscape}, with generation count denoted by the vertical axis.
-As was shown in Figure \ref{fig:kinshipFig}, here too we superimpose a green node that has parents from the first and second generations.
-This green node is both a second and third generation individual of the original parent, and should be displayed in both corresponding generation positions on the vertical axis.
-However, standard graph visualization tools only allow for a given node to be displayed once.
-As a result, this green node must be ambiguously positioned as either a second or third generation individual; here, it is positioned as a third generation individual.}
+As was shown in Figure \ref{fig:kinshipFig}, the green node has parents from two different generations, and hence must be ambiguously positioned as one of two generation counts.}
     \label{fig:Graph}
@@
 \section{Example Datasets}
+\label{exData}
 
-The \proglang{ggenealogy} package comes with two example datasets, one comprises a soybean genealogy and the other comprises an academic statistician genealogy.
+The \pkg{ggenealogy} package comes with two example datasets, one comprises a soybean genealogy and the other comprises an academic statistician genealogy.
 We will introduce both example datasets in this paper to demonstrate some of the tools available in the software.
@@
 
-We start with the soybean genealogy, which is available as a data frame structure with 412 rows, each representing a direct parent-child relationship between a pair of soybean varieties.
-In total, there are 230 unique soybean varieties in the dataset.
+We start with the soybean genealogy, which is available as a data frame structure with 390 rows and five columns.
+Each row contains information about a particular child soybean variety, including the name of the child, its protein yield, the year it was developed, whether or not its developmental year was imputed, and the name of its parent.
+Excepting the column that indicates the name of the child, the other four columns can have missing values (NA).
+As a result, while each row does contain information about a particular child soybean variety, whether or not a given row also contains information about a parent-child relationship between a pair of soybeans depends on whether or not the parent column has a missing value.
+
+In total, there are 230 soybean varieties in the dataset, 206 of which are children and 165 of which are parents.
+There are soybeans that are both children and parents.
+Of the children, 156 have two parents, 28 have one parent, and 22 have zero parents.
+There are 340 parent-child relationships in the dataset.
+
 These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-They also contain information on the developmental years, copy number variants, single nucleotide polymorphisms, and protein content and yield for each of the soybeans.
-
+They also contain information on the copy number variants, single nucleotide polymorphisms, and protein content for each of the soybeans, although we removed that information for a succinct example dataset.
 In this context, the software could ideally be used by agronomists who wish to study how soybean varieties are related.
 By referencing the visualization of the genealogical structure, these scientists may better understand genetic testing results - in this particular dataset, in terms of copy number variants, single nucleotide polymorphisms, and protein content and yield - and use that knowledge in future breeding sessions.
+
+We can load the example dataset of soybean genealogy (\code{sbGeneal}) and examine its structure.
 
@@
 \begin{CodeOutput}
-'data.frame':	412 obs. of  5 variables:
+'data.frame':	390 obs. of  5 variables:
  $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
@@
 
-In addition to the soybean genealogy, the \proglang{ggenealogy} package also comes with an academic genealogy of statisticians.
-To develop this later dataset, we contacted the Mathematics Genealogy Project, a web-based database for the genealogy of academic mathematicians.
+The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns.
+Each row contains information about a particular child who received a graduate-level academic degree in statistics, including the name of the child, the year the child obtained the degree, the country and school from which the child obtained the degree, the thesis title of the degree awarded to the child, and the name of its parent.
+There are no missing values for the country and school from which the child received its degree or the name of the child; however, some of the years contain missing values (NA), and some of the parent and thesis names contain empty strings (``").
+As a result, while each row does contain information about a particular child, whether or not a row also contains information about a parent-child relationship between a pair of academic statisticians depends on whether or not the parent column has an empty string.
+
+In total, there are 7122 individuals in the dataset, 7122 of which are children and 872 of which are parents.
+Every parent is also a child, but not every child is also a parent.
+Of the children, two have four parents, ten have three parents, 226 have two parents, 2801 have one parent, and 4083 have no parents.
+There are 3291 parent-child relationships in the dataset.
+
+To develop this later dataset, we contacted the \citealt{mgp}, a web-based database for the genealogy of academic mathematicians.
 This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society.
@@
 One of these variables was called ``msc" (Mathematics Subject Classification), and we selected only those entries that contained a value of 62 for this variable (coded as ``Statistics").
-Furthermore, we only retained entries that had an advisor if that advisor was also in the field of ``Statistics".
-This process resulted in 8995 entries, which we reduced to 8166 entries by removing duplicate entries.
-With the final data frame of 8995 entries, we only maintained 6 of the original 26 variables.
+Furthermore, we only retained entries that had a parent if that parent was also in the field of ``Statistics".
+Hence, in our parent-child relationships, both the child and the parent received postbaccalaureate degrees in statistics, and the parent was the academic advisor to the child.
+This process resulted in 8995 entries, which we reduced to 8165 entries by removing duplicate entries.
+With the final data frame of 8165 entries, we only maintained 6 of the original 26 variables.
 
-This process resulted in the second example dataset provided in the \proglang{ggenealogy} package.
-This dataset is structured as a data frame that contains 8166 rows, each of which corresponds to a parent-child relationship, where both the child and the parent received postbaccalaureate degrees in statistics, and the parent was the academic advisor to the child.
-The columns of the dataset provide information about the year the child obtained the degree, the country and school from which the child obtained the degree, and the thesis title of the degree awarded to the child.
+We can load the example dataset of academic genealogy of statisticians (\code{statGeneal}) and examine its structure.
 
@@
 \begin{CodeOutput}
-[1] 8166    6
+[1] 8165    6
 \end{CodeOutput}
@@
 
+\section{Genealogical Input Format}
+
+As is the case with both example data files introduced above, \pkg{ggenealogy} requires that the genealogy input file is a data frame structure with at least two columns.
+One column must be labeled ``child", and each case in that column must be of type character.
+The other column must be labeled ``parent," and each case in that column must either be of type character, type NA, or type ``".
+At this point, any \pkg{ggenealogy} plot that only requires information about parent-child relationships can be used.
+
+However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy.
+For these plots, the input data frame should also contain a third column.
+In both example data files, this column is labeled ``year," and each case in that column can either be of type numeric, type NA, or type ``".
+At this point, any \pkg{ggenealogy} plot can be used.
+
 \section{Generating a Graphical Object}
 
-Most functions in the software package require an input parameter of a graph structure.
+Most functions in the \pkg{ggenealogy} software package require an input parameter of a graph structure.
 Therefore, as a preprocessing step, we must first convert our original data frame structure into a graph structure.
@@
 
-There are many statistics about the \code{sbGeneal} genealogical dataset that we may wish to know that cannot easily be obtained through images and tables.
+There are many statistics about the \code{sbGeneal} dataset that we may wish to know that cannot easily be obtained through images and tables.
 The package function \code{getBasicStatistics()} can be called, using the \code{sbIG} object as input.
-This will return a list of common graph theoretical measurements regarding the genealogical graph structure.
+This will return a list of common graph theoretical measurements regarding the genealogical structure.
 For instance, is the whole structure connected?
@@
 
-As this data set deals with soy bean lineages, it may be useful for agronomists to track how two varieties are related to each other via parent-child relationships.
-Then, any dramatic changes in protein yield, SNP varieties, and other measures of interest between the two varieties can be traced across their genetic timeline, and pinpointed to certain varieties within their historical lineage.
-The \pkg{ggenealogy} software allows users to select two varieties of interest, and determine the shortest pathway of parent-child relationships between them, using the \code{getPath()} function.
-This will return a list path that contains the variety names and their years in the path.
+With soybean lineages, it may be useful for agronomists to track how two varieties are related to each other via parent-child relationships.
+Then, any dramatic changes in protein yield and other measures of interest between the two varieties can be traced across their genetic timeline.
+The \pkg{ggenealogy} package allows users to select two varieties of interest, and determine the shortest pathway of parent-child relationships between them, using the \code{getPath()} function.
+This will return a list that contains the variety names and their years in the path.
 
@@
 
-The returned path object can then be plotted using the \code{plotPath()} function, as demonstrated below.
+The returned path object can then be plotted using the \code{plotPath()} function.
 
@@
     \caption{Left: The shortest path between varieties Tokyo and Narow is strictly composed of a unidirectional sequence of parent-child relationships.
-Right: The shortest path between varieties Zane and Bedford is not strictly composed of unidirectional parent-child relationships, but have a cousin-like relationship.}
+Right: The shortest path between varieties Zane and Bedford is not strictly composed of unidirectional parent-child relationships; they instead have a cousin-like relationship.}
     \label{fig:pathTNZB}
@@
 
-This produces a visual that informs users of all the varieties involved in the shortest path between the two varieties of interest, see left half of Figure \ref{fig:pathTNZB}.
+This produces a visual that informs users of all the varieties involved in the shortest path between the two varieties of interest (see left half of Figure \ref{fig:pathTNZB}).
 In this plot, the years of all varieties involved in the path are indicated on the horizontal axis, while the vertical axis has no meaning other than to simply to display the labels evenly spaced vertically.
@@
 The resulting plot (right half of Figure \ref{fig:pathTNZB}) allows us to quickly determine that \code{Bedford} is not a parent, grandparent, or any great grandparent of \code{Zane}.
-Instead, we see that these two varieties are not related through a unidirectional parent-child lineage, but have a cousin-like relationship.
+Instead, we see that these two varieties are not related through a unidirectional parent-child lineage, but instead have a cousin-like relationship.
 The oldest common ancestor between \code{Zane} and \code{Bedford} is the variety \code{D55-4090}, which was developed in the mid 1940s.
@@
 Furthermore, as determined by the figure, for both \code{Zane} and \code{Bedford}, there are four varieties of unidirectional parent-child relationships between each of them and their common ancestor \code{D55-4090}.
-Hence, any parameter of interest that differentiates \code{Zane} and \code{Bedford} (protein yield, disease resistance, etc.) can also be examined across these two separate lineage histories.
+Hence, any feature of interest that differentiates \code{Zane} and \code{Bedford} (protein yield, disease resistance, etc.) can also be examined across these two separate lineage histories.
 
@@
 
-Now that we can create path objects, we may wish to know how those paths are positioned in comparison to the genealogical lineage of the entire data structure.
+Now that we can create path objects, we may wish to know how those paths are positioned compared to the entire genealogical lineage.
 For instance, of the documented soybean cultivar lineage varieties, where does the shortest path between two varieties of interest exist?
-Are these two varieties comparatively older compared to the overall data structure?
+Are these two varieties older compared to the overall data structure?
 Are they newer?
@@
 There is a function available in the \pkg{ggenealogy} package \code{plotPathOnAll()} that can allow users to quickly visualize their path of interest superimposed over all varieties and edges present in the whole data structure.
-Here we will produce a plot of the previously-determined shortest path between varieties \code{Tokyo} and \code{Narow} across the entire dataset:
+Here we will produce a plot of the shortest path between varieties \code{Tokyo} and \code{Narow} across the entire dataset, as is displayed in Figure \ref{fig:plotTNBin3}.
 
 \begin{Code}
-R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3,
-  pathColor = "red")
-R> plotAllImage ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
-R> plotAllImage <- plotPathOnAll(path = path, geneal = sbGeneal, ig = sbIG
-  binVector = sample(1:3), edgeCol = "gray84", pathEdgeCol = "blue", nodeSize
-  = 1, pathNodeSize = 3)
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red",
+  nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
+  element_text(size = 12), axis.title =ggplot2::element_text(size = 12))
 \end{Code}
@@
 
-While the first three explicit parameters to the function \code{plotPathOnAll()} have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation.
-The motivation of the \code{plotPathOnAll()} is to write variety text labels on a plot, with the center of each variety label constricted on the horizontal axis to its developmental year.
-As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the text labels.
-Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have labels with large numbers of characters (width of x coordinate).
+In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions (\citealt{ggplot2}).
+While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation.
+The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, developmental year).
+As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels.
+Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have long text labels (width of x coordinate).
 
@@
 However, for each variety, the y coordinate is arbitrary.
-Hence, in an attempt to mitigate text overlapping, the \code{plotPathOnAll()} does not randomly assign the y coordinate.
+Hence, in an attempt to mitigate text overlapping, the \code{plotPathOnAll()} function does not randomly assign the y coordinate.
 Instead, it allows users to partially control the y coordinates with a user-determined number of bins (\code{binVector}).
 
-If the user determines to produce a plot using three bins, as in the example code above, then the varieties are all grouped into three bins based on their years of development.
-In other words, there will be bin 1 (the ``oldest bin") which includes one-third of the total number of varieties all with the oldest developmental years, bin 2 (the ``middle bin"), and bin 3 (the ``youngest bin").
+If the user decides to produce a plot using three bins, as in the example code above, then the varieties are all grouped into three bins based on their year values.
+In other words, there will be bin 1 (the ``oldest bin") which includes the one-third of varieties with the oldest developmental years, bin 2 (the ``middle bin"), and bin 3 (the ``youngest bin").
+Then, in order to decrease text overlap, the consecutively increasing y-axis coordinates are alternatively assigned to the three bins (For example: bin 1, bin 2, bin 3, bin 1, bin 2, bin 3, ...) repeatedly until all varieties are addressed.
+This algorithm means that for any pair of varieties within a given bin, there are exactly two other varieties vertically separating them.
 
-Then, in order to decrease text overlap, the consecutively increasing y-axis coordinates are alternatively assigned to the three bins (For example: bin 1, then bin 2, then bin 3, then bin 1, then bin 2, then bin 3, ...) repeatedly until all varieties are accounted for.
-This algorithm means that for any pair of varieties within a given bin constrained to those years on the horizontal axis, there are exactly two other varieties placed between them vertically on the y-axis that come from the two other bins 1 constrained to a different set of year values on the horizontal axis.
+In the code above, \code{binVector} was assigned a value of 3, and \code{pathEdgeCol} was assigned a value of ``red".
+Additionally, we specified a size of 2.5 for the non-path node test using the \code{nodeSize} parameter, and a size of 4 for the path node text using the \code{pathNodeSize} parameter.
+There are several other parameters in the \code{plotPathOnAll()} function, which can be read in more detail using the help command.
 
-Hence, in the code above, the user selected a \code{binVector} value of three, and a \code{plotColor} of red, which produces the plot in Figure \ref{fig:plotTNBin3}.
-We see that edges not on the path of interest are thin and gray by default, whereas edges on the path of interest are bolded and red.
-We also see that varieties in the path of interest are boldfaced by default.
+This code resulted in Figure \ref{fig:plotTNBin3}, where we see that edges not on the path of interest are thin and gray by default, whereas edges on the path of interest are bolded by default.
+We also see that variety labels in the path of interest are boldfaced by default.
+Figure \ref{fig:plotTNBin3} presents useful information: We immediately gather that the path of interest does span most of the years of the data structure.
+In fact, \code{Tokyo} appears to be the oldest variety in the dataset, and \code{Narow} appears to be one of the youngest varieties.
+We can also determine that the majority of varieties appear to have development years between 1950 and 1970.
 
-The plot presents useful information: We immediately gather that the path of interest between does span most of the years of the data structure.
-In fact, \code{Tokyo} appears to be the oldest variety present in the dataset, and \code{Narow} appears to be one of the youngest varieties.
-We can also determine that the vast majority of varieties appear to have development years between 1950 and 1970.
-
-However, this plot has significant empty spaces between the noticeably distinct bins, whereas almost all text labels are overlapping, thereby decreasing their readability.
-To force some variety text labels into these spaces, the user may consider using a larger number of bins.
-Hence, we next examine a bin size of 6:
+However, Figure \ref{fig:plotTNBin3} has significant empty spaces between the noticeably distinct bins, whereas almost all text labels are overlapping, thereby decreasing their readability.
+To force text labels into these spaces, the user may consider using a larger number of bins.
+Hence, we next examine a bin size of 6 to create Figure \ref{fig:plotTNBin6}.
 
 \begin{Code}
-R> plotAllImage <- plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6,
-  pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol =
+  "seagreen2", nodeSize = 1, pathNodeSize = 3)
 \end{Code}
 
-\begin{figure}%[h]
+\begin{figure}[H]
     \centering
@@
 
-This similar code now outputs the plot shown in Figure \ref{fig:plotTNBin6}.
-We can immediately see that this plot more successfully mitigates text variety label overlap than the previous plot in \ref{fig:plotTNBin3}.
+We can immediately see that Figure \ref{fig:plotTNBin6} more successfully mitigates text overlap compared to Figure \ref{fig:plotTNBin3}.
 We can also confirm what we saw in the previous plot that indeed most varieties have development years between 1950 and 1970, and any textual overlap is confined to this range of years.
 
-\section{Future Directions: Fine-tuning Superimposition}
+\section{Plotting Ancestors and Descendants by Generation}
+\label{remedy}
 
-This visualization tool is suitable as part of a data exploration phase, but not for any users seeking publication quality plots due to the remaining textual overlap.
-We continue to work towards further reducing this textual overlap, although it is impossible to guarantee no textual overlap, especially with larger datasets with dense subgroups of varieties having similar years.
+The most novel visual function in \pkg{ggenealogy}, \code{plotAncDes()} allows users to view the ancestors and descendants of a given variety.
+The inputted variety is highlighted in the center of the plot, ancestors are displayed to the left of the center, and descendants are displayed to the right of the center.
+The further from the center that a variety is located, the more generations that variety is distanced from the centered variety of interest.
 
-As such, we plan to add a feature to the \pkg{ggenealogy} package that allows users to manually fine-tune the plot they deem best after examining various bin sizes.
-For example, after comparing several bin sizes (1-12) on the current soy bean data set, we determined that the bin size of 6 produced minimal textual overlap, as seen in \ref{fig:plotTNBin6}.
+This particular \pkg{ggenealogy} tool is unique because most available genealogy and graph visualization software do not allow for repeated labels.
+It is a useful tool because, as was demonstrated in Figures \ref{fig:kinshipFig} and \ref{fig:Graph}, some genealogical datasets require repeated node labels if they are to be visualized by generation counts.
+Indeed, our example soybean genealogy is one such dataset.
 
-However, there still remained one dozen cases of partial textual overlap.
-As an example, the labels of \code{Crawford} (1974) and \code{Swift} (1973) overlap at just below the vertical midpoint of the plot.
-The proposed function would allow users to hardcode the label of either variety and manually assign it to a new vertical coordinate.
-For instance, a user might select \code{Swift} and slightly decrease its vertical coordinate so that it is drawn halfway between \code{Crawford} and \code{Bradley}.
-
-If the user sequentially fine-tuned the vertical positions of overlapping text labels for the small fraction of labels that remained overlapped after the automated function, and if they monitored the progress by visually inspecting updated plots until there were no more overlaps, then the plot could be used in presentations and publications.
-
-\section{Plotting Ancestors and Descendants by Generation}
-
-The most novel visual tool in \pkg{ggenealogy}, \code{plotAncDes()}, allows users to view the ancestors and descendants of a given variety.
-The inputted variety is highlighted in the center of the plot, ancestors are displayed to the left of the center, and descendants to the right of the center.
-The further from the center, the larger the number of generations that particular ancestor/descendant is from the centered variety of interest.
-
-This particular \pkg{ggenealogy} tool is unique because most available graphical software only consider simple graphs, with no repeated nodes.
-However, in some genealogical lineage datasets, some varieties must be repeated if they are to be visualized by generation counts.
-This was found to be the case in the soy bean dataset.
-
-As an example, we will create a plot of the ancestors and descendants of the variety \code{Lee}.
+To demonstrate this tool, we will create a plot of the ancestors and descendants of the variety \code{Lee}.
 We specify that the maximum number of ancestor and descendant generations are both 6, and that the text of the variety of interest is highlighted in blue:
@@
 
-\begin{figure}%[h]
+This generates the plot in the left side of Figure \ref{fig:Lee}.
+We notice that \code{Lee} has 3 generations of ancestors and 5 generations of descendants.
+We also notice that some varieties are repeated in the plot, which is a unique feature provided by \pkg{ggenealogy}.
+For example, the variety \code{5601T} is represented four times - once as a third generation descendant of \code{Lee}, once as a fourth generation descendant of \code{Lee}, and twice as a fifth generation descendant of \code{Lee}.
+The variety \code{5601T} was repeated multiple times because there are multiple paths between \code{Lee} and \code{5601T}.
+For explanation purposes, all paths between \code{Lee} and \code{5601T} were manually highlighted in blue.
+
+The right side of Figure \ref{fig:Lee} is not an output plot of \pkg{ggenealogy}.
+Instead, it was simply created for didactic purposes.
+Here, the paths that were manually highlighted in blue in the left plot produced by \pkg{ggenealogy} are shown again, only now nodes cannot be repeated.
+The parenthetical number above each node represents the set of generation counts distancing that node from the center node \code{Lee}; green parentheses indicate that the node could be successfully placed in one horizontal position, but red parentheses indicate that the node could not be successfully placed in one horizontal position.
+We see that node \code{TN89-39} cannot simultaneously be represented as both a third and fourth descendent of node \code{Lee}, and node \code{5601T} cannot simultaneously be represented as a third, fourth, and fifth descendent of node \code{Lee}.
+Hence, without allowing nodes to repeat, this dataset cannot be presented in the graph on the right as it can be in the \pkg{ggenealogy} graph on the left.
+This is a current limitation in other genealogy and graphical software that \pkg{ggenealogy} can now provide.
+
+\begin{figure}[H]
     \centering
     \includegraphics[width=\textwidth]{Lee}
-    \caption{Left: All varieties within three generations of ancestors and five generations of descendants of the variety Lee are shown.
-The generation distance from the center variety is neatly displayed in this graph; the larger the generation distance from the center variety, the further left (for ancestors) or right (for descendants).
-For explanation purposes, all paths between Lee and 5601T are highlighted in blue.
-We see that 5601T appears four times in the plot, which is a unique and novel requirement provided by ggenealogy.
-Right: The paths that are highlighted in blue in the left plot from ggenealogy are shown here again, only now nodes cannot be repeated.
-We unsuccessfully try to constrain the horizontal position of the nodes by generation count, as was accomplished by the ggenealogy plot, without repeating nodes.
-The parenthetical number above each node represents the set of generation counts that node is away from the center node Lee; green ones indicate that the node could be successfully placed in one horizontal position, but red ones indicate that the node could not be successfully placed in only one horizontal position.
-Hence, without allowing nodes to repeat, this data information cannot be presented as it is in the ggenealogy graph on the left, and this is a current limitation in other currently-available graphical software that ggenealogy can now provide.}
+    \caption{Left: All ancestors and descendants of the variety Lee are shown in this ggenealogy plot.
+Right: We now attempt to mimic the blue paths in the ggenealogy plot on the left, only now nodes cannot be repeated.
+The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee.
+The presence of red parentheses indicate that the plot on the right cannot unambiguously display the example soybean genealogy in the way that the ggenealogy plot on the left can accomplish.}
     \label{fig:Lee}
 \end{figure}
-
-This generates the plot we see in the left side of Figure \ref{fig:Lee}.
-We notice that \code{Lee} has 3 generations of ancestors and 5 generations of descendants.
-However, we also notice that some varieties are repeated in the plot.
-For example, the variety \code{5601T} is represented four times - once as a third generation descendant of \code{Lee}, once as a fourth generation descendant of \code{Lee}, and twice as a fifth generation descendant of \code{Lee}.
-
-This happens because there are various paths between \code{Lee} and \code{5601T} (see the right side of Figure \ref{fig:Lee}).
-Hence, if we are to present all ancestors and descendants of \code{Lee} - while constricting the horizontal axis to generational distance from \code{Lee} - then varieties like \code{5601T} must occur more than once in the plot.
-
-One of the main motivations in developing the \code{plotAncDes()} function was an inability to find other similar software (\citealt{ape}) that could produce such a plot, where repeated nodes were permitted, see right side of Figure \ref{fig:Lee} for additional explanation.
-The \code{plotAncDes()} function generates plots that have increased readability, as it is relatively easy to compare how far varieties are from the center to obtain an idea of generational distance.
 
@@
 Here we generate a distance matrix for a set of 10 varieties, setting the x-label and y-label as ``Variety" and the legend label as ``Degree".
-Syntax from the \pkg{ggplot2} package (\citealt{ggplot2}) can be appended to the \code{plotDegMatrix()} function.
-In this example, we add \pkg{ggplot2} functionality to specify that pairs with small degrees are white, while those with large degrees are dark green, as well as to specify the label size of the legend title and text:
+In this example, we add \pkg{ggplot2} functionality to specify that pairs with small degrees are white, while those with large degrees are dark green, as well as to specify the text size of the legend title and label.
 
@@
   ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
-  ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
-  legend.text = ggplot2::element_text(size = 15))
+  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text
+  = ggplot2::element_text(size = 15))
 \end{Code}
+
+This creates the plot in Figure \ref{fig:degMatrix}.
+We see that the degree of the shortest path between varieties \code{Bedford} and \code{Zane} is 10, which is consistent with what we saw earlier in Figure \ref{fig:pathTNZB}.
+However, we now also see that a degree of 10 may be considered a large degree for this dataset.
 
@@
 
-This creates the plot in Figure \ref{fig:degMatrix}.
-We see that the degree of the shortest path between varieties \code{Bedford} and \code{Zane} is 10, which is consistent with what we saw earlier in Figure \ref{fig:pathTNZB}.
-However, we now also see that 10 may be a comparatively large degree in this same dataset.
-
 \section{Academic Genealogy of Statisticians}
 
-%\section{Plotting Ancestors and Descendants by Generation}
+The \pkg{ggenealogy} package comes with two example datasets, and while we have introduced the agronomical genealogy, we have yet to introduce the academic genealogy.
+As was demonstrated in Section \ref{exData}, every parent in the academic genealogy is also a child, and some children in the academic genealogy have more than two parents.
+Neither of these features was the case in the agronomical genealogy.
+Additionally, the academic genealogy is much larger than the agronomical genealogy.
+Some of these differences may affect how one would approach \pkg{ggenealogy} plotting tools.
+For this reason, we will now demonstrate some of the \pkg{ggenealogy} plotting tools we already introduced, only now applied to the academic genealogy.
 
-This second example dataset of academic genealogy is of a slightly different structure than the first example dataset of agronomical genealogy, because each child no longer requires exactly two parents.
-The ability to plot ancestors and descendants by generation was demonstrated on the agronomical genealogy in Figure \ref{fig:Lee}.
-As we believe this is the most novel plotting tool in the \proglang{ggenealogy} package, we can test it again here using the academic genealogy.
+The ability to plot ancestors and descendants by generation was demonstrated using the agronomical genealogy in Figure \ref{fig:Lee}.
+As we believe this is the most novel plotting tool in the \pkg{ggenealogy} package, we will test it again here using the academic genealogy.
 
-We can determine which academic statistician in this \code{statGeneal} dataset has the largest number of descendants.
-To do so, we create a vector \code{indVec} that contains the names of all individuals (whether child or parent) in the dataset.
-We then use the \proglang{dplyr} package to apply the ggenealogy function \code{getDescendants} on each individual in the \code{indVec} vector (\citealt{dplyr}).
+We need to choose a central individual of interest in order to create this plot.
+Perhaps we can use the academic statistician in the dataset that has the largest number of descendants.
+To determine the name of this individual, below we use the \pkg{ggenealogy} function \code{getNode()} to create a vector \code{indVec} that contains the names of all individuals in the dataset.
+We then use the \pkg{dplyr} package to apply the \pkg{ggenealogy} function \code{getDescendants()} on each individual in the \code{indVec} vector (\citealt{dplyr}).
 We set the parameter \code{gen} to a conservatively large value of 100 as this dataset is unlikely to have any individuals with more than 100 generations of descendants.
 
-We then create a table to examine all values of descendant counts observed in the dataset, along with the number of individuals who had each of those values of descendant counts.
-Of the 8166 individuals in this dataset, 6252 of them have zero descendants, 322 of them have one descendant, and 145 of them have two descendants.
+After that, we can generate a table to examine all values of descendant counts in the dataset, along with the number of individuals who have each of those values of descendant counts.
+Of the 8165 individuals in this dataset, 6252 of them have zero descendants, 322 of them have one descendant, and 145 of them have two descendants.
 There are only 17 individuals who have more than 30 descendants, and there is one individual who has the largest value of 159 descendants.
@@
 >R library(dplyr)
->R indVec <- unique(c(statGeneal$child, statGeneal$parent))
+>R indVec <- getNodes(statGeneal)
 >R indVec <- indVec[which(indVec != "", )]
@@
 
-We can now visualize how these 159 descedants are related to Sir David Cox by plotting parent-child relationships, with generation count constricted on the horizontal axis, similar to what we performed in Figure \ref{fig:Lee}.
-We create this in Figure \ref{fig:dCox} using the code below.
+We can now visualize how these 159 descendants are related to Sir David Cox by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}.
+As such, we create Figure \ref{fig:dCox} using the code below.
 
@@
     \includegraphics[width=\textwidth]{dCox.png}
-    \caption{The 159 academic descendants of Sir David Cox.}
+    \caption{The 159 academic statistician descendants of Sir David Cox.}
     \label{fig:dCox}
@@
 
-We see from Figure \ref{fig:dCox} that Sir David Cox had 42 children, many of them becoming notable statisticians, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
-The child of his who produced the most children of his own was Peter Bloomfield, who has 26 children and 49 descendants.
-In total, Sir David Cox has five generations of statistics students in this dataset.
+We see from Figure \ref{fig:dCox} that Sir David Cox had 42 children, many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
+Of his children, the one who produced the most children of their own was Peter Bloomfield, who has 26 children and 49 descendants.
+In total, Sir David Cox had five generations of academic statistics mentees in this dataset.
 
@@
 
-It would be insightful to examine a more detailed view of one of the longest strings of parent-child relationships between Sir David Cox and one of his fifth generation descendants.
+At this point, it would be insightful to examine a more detailed view of one of the longest strings of parent-child relationships between Sir David Cox and one of the two individuals who are his fifth generation descendants.
+We do so with the code below, choosing his fifth generation descendant to be Petra Buzkova.
 
@@
 R> statIG <- dfToIG(statGeneal)
-R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected
-  = FALSE)
+R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
+  isDirected = FALSE)
 R> plotPath(pathCB) + ggplot2::theme(axis.text = ggplot2::element_text(size
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
     \includegraphics[width=\textwidth]{pathCB}
-    \caption{The shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional parent-child relationships that span about 55 years.
-We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than about 30 years in degree aquisition, whereas the three statisticians who earned their degrees later in this path only span less than about 10 years in degree acquisition.}
+    \caption{The shortest path between Sir David Cox and one of his fifth generation descendants, Petra Buzkova.}
     \label{fig:pathCB}
@@
 
-\begin{figure}%[h]
+This code results in Figure \ref{fig:pathCB}.
+We see that the shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional parent-child relationships that span about 55 years.
+We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than about 30 years in degree acquisition, whereas the three statisticians who earned their degrees later in this path only span less than about ten years in degree acquisition.
+
+We also notice in Figure \ref{fig:pathCB} that Sir David Cox received his statistics degree in about 1950, and Petra Buzkova received her statistics degree in about 2005.
+This genealogy only contains historical information about obtained degrees, and does not project into the future.
+Hence, we can be assured that Petra Buzkova is one of the younger individuals in the dataset, at least in the sense that the youngest individual could only have received his or her degree ten years after Petra Buzkova.
+However, we cannot be assured that Sir David Cox is one of the oldest individuals in the dataset.
+As such, it would be informative to superimpose this path of interest onto the entire dataset, using the \code{plotPathOnAll()} function of the \pkg{ggenealogy} package, as we did for the soybean genealogy in Figures \ref{fig:plotTNBin3} and \ref{fig:plotTNBin6}.
+
+We can achieve this using the below code.
+After trial and error, we use a \code{binVector} of size 200, and append \pkg{ggplot2} syntax to define suitable x-axis limits.
+The output of this process is illustrated in Figure \ref{fig:plotCBText}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::
+  scale_x_continuous(limits = c(1850, 2030)) + ggplot2::theme(axis.text =
+  ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size
+  = 12))
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{figure}[H]
     \centering
     \includegraphics[width=\textwidth]{plotCBText}
-    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
+    \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.}
     \label{fig:plotCBText}
@@
 
-\begin{figure}%[h]
+We see from the resulting Figure \ref{fig:plotCBText} that almost all text labels for individuals who received their graduate-level statistics degrees between 1950 and 2015 are undecipherable.
+We also see that the year Sir David Cox acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for statistics degree acquisitions in this dataset occur around 1860.
+However, the number of individuals who are documented as receiving their statistics degrees between 1860 and 1950 are few enough so that their text labels are somewhat readable.
+
+The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{binVector} will not solve the text overlapping problem.
+Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying a value of \code{TRUE} for the input parameter \code{nodeLabel}.
+As is further detailed in the help manual of the function, doing so will generate a plot where only text labels for individuals on the path of interest are labeled; all other individual labels are represented as dots.
+We provide the code for this below, which results in Figure \ref{fig:plotCBNoText}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = 4,
+  pathNodeSize = 4, nodeLabel=FALSE) + ggplot2::theme(axis.text = ggplot2::
+  element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{figure}[H]
     \centering
     \includegraphics[width=\textwidth]{plotCBNoText}
-    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
+    \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.
+Individuals on the shortest path have text labels; all other individuals have dot labels.}
     \label{fig:plotCBNoText}
 \end{figure}
+
+In Figure \ref{fig:plotCBNoText}, we can now see each individual on the path of interest, and how their values for the variable year are overlaid on the entire genealogy structure.
+We can also more clearly see that, even though only ten years span between the youngest individual in the genealogy and Petra Buzkova, there are many individuals in that last decade.
+Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of statistics degree acquisitions.
+
+\section{Future Avenues}
+
+Incorporation of the Shiny application would allow users to examine \pkg{ggenealogy} tools in a more interactive way (\citealt{shiny}).
+The reactive programming would save them the time of using command-line for each change of input as well as the inefficiency of rerunning code.
+We also aim to incorporate plotting tools that can examine not only quantitative variables (such as our example variable of ``year"), but also categorical variables associated with individuals in datasets.
+Moreover, we look forward to testing the \pkg{ggenealogy} package on additional genealogical data sets.
+Exploring several datasets with the software will allow us to fix remaining bugs, and provide us further insight into how to make our tools available for a wide range of data input formats.
+
+The \pkg{ggenealogy} visualization tool \code{plotPathOnAll()} is suitable as a data exploration tool, but not always as a publication tool.
+This is because we still see textual overlap in small-enough datasets (see Figure \ref{fig:plotTNBin6}).
+As such, we plan to add a feature to the package that allows users to manually fine-tune automated plots.
+For example, after comparing several bin sizes on the soybean genealogy, we determined that the bin size of 6 produced the minimal textual overlap, as is seen in Figure \ref{fig:plotTNBin6}.
+If we could subsequently fine-tune the vertical positions of the small fraction of text labels that remained overlapped after the automated \pkg{gggenealogy} function, then we could potentially remove all overlaps, and the plot could be used in presentations and publications.
+Of course, it is impossible to eliminate textual overlap in larger datasets (see Figure \ref{fig:plotCBText}).
+In such cases, we can remedy this problem by representing individuals who are not on the path of interest with dots instead of text (see Figure \ref{fig:plotCBNoText}).
 
@@
 If users are simultaneously using similar packages, we in particular recommend using the \code{plotAncDes()} function.
-This plot allows users to view generation counts of a variety of interest in a manner that is not as readily available in similar software packages (\citealt{ape}).
+This plot allows users to view generation counts of a variety of interest in a manner that is not as readily available in similar software packages.
 
-\section{Future Avenues}
-
-Incorporation of the Shiny application (\citealt{shiny}) would allow users to examine \pkg{ggenealogy} tools in a more interactive way.
-The reactive programming would save them the time of using command-line for each change of input as well as the inefficiency of rerunning code.
-We also look forward to testing this package on additional genealogical data sets (\citealt{mgp}).
-Exploring several datasets with our software will allow us to fix remaining bugs, and provide us further insight into how to make our tools available for a flexible set of data inputs.
-
-\section*{Acknowledgments} % Lindsay added
+\section*{Acknowledgments}
 

### 2016-04-18 JSS/jss-style/article.tex
plotly ref

@@
 The tools are being tested on a dataset with milestone cultivars of soybean varieties (\citealt{soybean}) as well as on a web-based database of the academic genealogy of mathematicians (\citealt{mgp}).
-The software package is currently available on the Comprehensive R Archive Network.}
+The software package is currently available on the Comprehensive \proglang{R} Archive Network.}
 
@@
 \begin{CodeInput}
+R> install.packages("ggenealogy")
 R> library("ggenealogy")
@@
 \begin{CodeInput}
-R> length(getChild("Peter Bloomfield", geneal))
+R> length(getChild("Peter Bloomfield", statGeneal))
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> nrow(getDescendants("Peter Bloomfield", geneal, gen = 100))
+R> nrow(getDescendants("Peter Bloomfield", statGeneal, gen = 100))
 \end{CodeInput}
@@
   pathNodeSize = 4, nodeLabel=FALSE) + ggplot2::theme(axis.text = ggplot2::
-  element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+  element_text(size = 12), axis.title = ggplot2::element_text(size = 12)) + 
+  ggplot2::xlab(1850, 2025)
 \end{CodeInput}
@@
 Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of statistics degree acquisitions.
+This figure is interactive with the use of Plotly (\citealt{plotly}).
 

### 2016-04-20 JSS/jss-style/article.tex
MG revisions

@@
   Michelle Graham\\
-  USDA-Agriculture Research Service, Corn, Insects and Crop Genetics Research Unit\\
+  USDA-Agriculture Research Service, Corn Insects and Crop Genetics Research Unit\\
   Department of Agronomy, Iowa State University\\
-  1575 Agronomy Building\\
+  1565 Agronomy Building\\
   Ames, IA, 50011, United States of America\\
-  E-mail: \email{magraham@iastate.edu}
+  E-mail: \email{michelle.graham@ars.usda.gov}
 }
@@
 We introduce here a developing visualization toolkit that is intended to assist users in their exploration and analysis of genealogical structures.
-In this paper, we will demonstrate the main tools of this software package \pkg{ggenealogy} using two example genealogical datasets, one of soybean cultivars (\citealt{soybean}) and the other of academic mathematicians (\citealt{mgp}).
+In this paper, we demonstrate the main tools of the software package \pkg{ggenealogy} using two example genealogical datasets, one of soybean cultivars (\citealt{soybean}) and the other of academic mathematicians (\citealt{mgp}).
 
@@
 Siblings are connected by horizontal sibship lines.
-Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational cross-breeding occurs, as is often the case in agronomical genealogical lineages.
+Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational breeding occurs, as is often the case in agronomic genealogical lineages.
 
-We demonstrate how the standardized pedigree charts in the \proglang{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational cross-breeding case in Figure \ref{fig:kinshipFig}.
+We demonstrate how the standardized pedigree charts in the \proglang{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational breeding case in Figure \ref{fig:kinshipFig}.
 In that figure, each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202.
@@
 However, standard pedigree tools only allow for an individual to be displayed once.
-As a result, in special cases where inter-generational cross-breading occurs, such as in agronomical applications, standardized tools for visualizing genealogical information cannot non-ambiguously portray the genealogical dataset.
+As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset.
 
@@
 Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated.
-In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational cross-breeding.
+In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational breeding.
 
@@
 We start with the soybean genealogy, which is available as a data frame structure with 390 rows and five columns.
-Each row contains information about a particular child soybean variety, including the name of the child, its protein yield, the year it was developed, whether or not its developmental year was imputed, and the name of its parent.
-Excepting the column that indicates the name of the child, the other four columns can have missing values (NA).
+These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
+They contain information on the copy number variants, single nucleotide polymorphisms, and protein content for each of the varieties, although we removed that information for a succinct example dataset.
+In this context, the software could ideally be used by agronomists who wish to study how soybean varieties are related.
+By referencing the visualization of the genealogical structure, these scientists may better understand genetic testing results - in this particular dataset, in terms of copy number variants, single nucleotide polymorphisms, protein content, and yield - and use that knowledge in future breeding decisions.
+
+Each row contains information about a particular child soybean variety, including the name of the child, its yield, the year it was released, whether or not its release year was imputed, and the name of its parent.
+While the name of the child is required, the other four columns can have missing values (which are represented in \proglang{R} with the symbol NA for ``not available").
 As a result, while each row does contain information about a particular child soybean variety, whether or not a given row also contains information about a parent-child relationship between a pair of soybeans depends on whether or not the parent column has a missing value.
@@
 There are 340 parent-child relationships in the dataset.
-
-These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-They also contain information on the copy number variants, single nucleotide polymorphisms, and protein content for each of the soybeans, although we removed that information for a succinct example dataset.
-In this context, the software could ideally be used by agronomists who wish to study how soybean varieties are related.
-By referencing the visualization of the genealogical structure, these scientists may better understand genetic testing results - in this particular dataset, in terms of copy number variants, single nucleotide polymorphisms, and protein content and yield - and use that knowledge in future breeding sessions.
 
@@
 The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns.
-Each row contains information about a particular child who received a graduate-level academic degree in statistics, including the name of the child, the year the child obtained the degree, the country and school from which the child obtained the degree, the thesis title of the degree awarded to the child, and the name of its parent.
-There are no missing values for the country and school from which the child received its degree or the name of the child; however, some of the years contain missing values (NA), and some of the parent and thesis names contain empty strings (``").
-As a result, while each row does contain information about a particular child, whether or not a row also contains information about a parent-child relationship between a pair of academic statisticians depends on whether or not the parent column has an empty string.
-
-In total, there are 7122 individuals in the dataset, 7122 of which are children and 872 of which are parents.
-Every parent is also a child, but not every child is also a parent.
-Of the children, two have four parents, ten have three parents, 226 have two parents, 2801 have one parent, and 4083 have no parents.
-There are 3291 parent-child relationships in the dataset.
-
 To develop this later dataset, we contacted the \citealt{mgp}, a web-based database for the genealogy of academic mathematicians.
@@
 This process resulted in 8995 entries, which we reduced to 8165 entries by removing duplicate entries.
-With the final data frame of 8165 entries, we only maintained 6 of the original 26 variables.
+With the final data frame of 8165 entries, we only maintained six of the original 26 variables.
+
+Each row of the final data frame contains information about a particular child who received a graduate-level academic degree in statistics, including the name of the child, the year the child obtained the degree, the country and school from which the child obtained the degree, the thesis title of the degree awarded to the child, and the name of its parent.
+There are no missing values for the country and school from which the child received its degree or the name of the child; however, some of the years contain missing values (NA), and some of the parent and thesis names contain empty strings (``").
+As a result, while each row does contain information about a particular child, whether or not a row also contains information about a parent-child relationship between a pair of academic statisticians depends on whether or not the parent column has an empty string.
+
+In total, there are 7122 individuals in the dataset, 7122 of which are children and 872 of which are parents.
+Every parent is also a child, but not every child is also a parent.
+Of the children, two have four parents, ten have three parents, 226 have two parents, 2801 have one parent, and 4083 have no parents.
+There are 3291 parent-child relationships in the dataset.
 
@@
 
-With soybean lineages, it may be useful for agronomists to track how two varieties are related to each other via parent-child relationships.
-Then, any dramatic changes in protein yield and other measures of interest between the two varieties can be traced across their genetic timeline.
+With soybean lineages, it may be useful for soybean breeders to track how two varieties are related to each other via parent-child relationships.
+Then, any dramatic changes in yield and other measures of interest between the two varieties can be traced across their genetic timeline.
 The \pkg{ggenealogy} package allows users to select two varieties of interest, and determine the shortest pathway of parent-child relationships between them, using the \code{getPath()} function.
@@
 This produces a visual that informs users of all the varieties involved in the shortest path between the two varieties of interest (see left half of Figure \ref{fig:pathTNZB}).
-In this plot, the years of all varieties involved in the path are indicated on the horizontal axis, while the vertical axis has no meaning other than to simply to display the labels evenly spaced vertically.
+In this plot, the release year of all varieties involved in the path are indicated on the horizontal axis, while the vertical axis has no meaning other than to simply to display the labels evenly spaced vertically.
 The shortest path between varieties \code{Tokyo} and \code{Narow} is composed of a unidirectional series of parent-child relationships, with \code{Tokyo} as the starting ancestor in the early 1900s, \code{Narow} as the most recent descendent in the mid 1980s, and three varieties in between.
@@
 Next, we can run the same set of functions on a different pair of varieties.
-First, a call to the \pkg{ggenealogy} function \code{getYear()} indicates that variety \code{Bedford} was developed in 1978 and variety \code{Zane} in 1985.
+First, a call to the \pkg{ggenealogy} function \code{getYear()} indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.
 
@@
 Instead, we see that these two varieties are not related through a unidirectional parent-child lineage, but instead have a cousin-like relationship.
-The oldest common ancestor between \code{Zane} and \code{Bedford} is the variety \code{D55-4090}, which was developed in the mid 1940s.
+The oldest common ancestor between \code{Zane} and \code{Bedford} is the variety \code{D55-4090}, which was released in the mid 1940s.
 
 Furthermore, as determined by the figure, for both \code{Zane} and \code{Bedford}, there are four varieties of unidirectional parent-child relationships between each of them and their common ancestor \code{D55-4090}.
-Hence, any feature of interest that differentiates \code{Zane} and \code{Bedford} (protein yield, disease resistance, etc.) can also be examined across these two separate lineage histories.
+Hence, any feature of interest that differentiates \code{Zane} and \code{Bedford} (protein content, yield, disease resistance, etc.) can also be examined across these two separate lineage histories.
 
@@
 While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation.
-The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, developmental year).
+The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, year of release).
 As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels.
@@
 If the user decides to produce a plot using three bins, as in the example code above, then the varieties are all grouped into three bins based on their year values.
-In other words, there will be bin 1 (the ``oldest bin") which includes the one-third of varieties with the oldest developmental years, bin 2 (the ``middle bin"), and bin 3 (the ``youngest bin").
+In other words, there will be bin 1 (the ``oldest bin") which includes the one-third of varieties with the oldest years of release, bin 2 (the ``middle bin"), and bin 3 (the ``youngest bin").
 Then, in order to decrease text overlap, the consecutively increasing y-axis coordinates are alternatively assigned to the three bins (For example: bin 1, bin 2, bin 3, bin 1, bin 2, bin 3, ...) repeatedly until all varieties are addressed.
@@
 In fact, \code{Tokyo} appears to be the oldest variety in the dataset, and \code{Narow} appears to be one of the youngest varieties.
-We can also determine that the majority of varieties appear to have development years between 1950 and 1970.
+We can also determine that the majority of varieties were released between 1950 and 1970.
 
@@
 We can immediately see that Figure \ref{fig:plotTNBin6} more successfully mitigates text overlap compared to Figure \ref{fig:plotTNBin3}.
-We can also confirm what we saw in the previous plot that indeed most varieties have development years between 1950 and 1970, and any textual overlap is confined to this range of years.
+We can also confirm what we saw in the previous plot that indeed most varieties were released between 1950 and 1970, and any textual overlap is confined to this range of years.
 
@@
 The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee.
-The presence of red parentheses indicate that the plot on the right cannot unambiguously display the example soybean genealogy in the way that the ggenealogy plot on the left can accomplish.}
+The presence of red parentheses indicate that the plot on the right ambiguously display the example soybean genealogy in the way that the ggenealogy plot on the left can accomplish.}
     \label{fig:Lee}
@@
 
-The \pkg{ggenealogy} package comes with two example datasets, and while we have introduced the agronomical genealogy, we have yet to introduce the academic genealogy.
+The \pkg{ggenealogy} package comes with two example datasets, and while we have introduced the plant breeding genealogy, we have yet to introduce the academic genealogy.
 As was demonstrated in Section \ref{exData}, every parent in the academic genealogy is also a child, and some children in the academic genealogy have more than two parents.
-Neither of these features was the case in the agronomical genealogy.
-Additionally, the academic genealogy is much larger than the agronomical genealogy.
+Neither of these features was the case in the plant breeding genealogy.
+Additionally, the academic genealogy is much larger than the plant breeding genealogy.
 Some of these differences may affect how one would approach \pkg{ggenealogy} plotting tools.
@@
 
-The ability to plot ancestors and descendants by generation was demonstrated using the agronomical genealogy in Figure \ref{fig:Lee}.
+The ability to plot ancestors and descendants by generation was demonstrated using the plant breeding genealogy in Figure \ref{fig:Lee}.
 As we believe this is the most novel plotting tool in the \pkg{ggenealogy} package, we will test it again here using the academic genealogy.
@@
 We need to choose a central individual of interest in order to create this plot.
-Perhaps we can use the academic statistician in the dataset that has the largest number of descendants.
+Perhaps we can use the academic statistician in the dataset that has the largest number of ``descendants".
 To determine the name of this individual, below we use the \pkg{ggenealogy} function \code{getNode()} to create a vector \code{indVec} that contains the names of all individuals in the dataset.
 We then use the \pkg{dplyr} package to apply the \pkg{ggenealogy} function \code{getDescendants()} on each individual in the \code{indVec} vector (\citealt{dplyr}).
-We set the parameter \code{gen} to a conservatively large value of 100 as this dataset is unlikely to have any individuals with more than 100 generations of descendants.
+We set the parameter \code{gen} to a conservatively large value of 100 as this dataset is unlikely to have any individuals with more than 100 generations of ``descendants".
 
-After that, we can generate a table to examine all values of descendant counts in the dataset, along with the number of individuals who have each of those values of descendant counts.
-Of the 8165 individuals in this dataset, 6252 of them have zero descendants, 322 of them have one descendant, and 145 of them have two descendants.
-There are only 17 individuals who have more than 30 descendants, and there is one individual who has the largest value of 159 descendants.
+After that, we can generate a table to examine all values of ``descendant" counts in the dataset, along with the number of individuals who have each of those values of ``descendant" counts.
+Of the 8165 individuals in this dataset, 6252 of them have zero ``descendants", 322 of them have one ``descendant", and 145 of them have two ``descendants".
+There are only 17 individuals who have more than 30 ``descendants", and there is one individual who has the largest value of 159 ``descendants".
 We determine that this individual is the prominent British statistician Sir David Cox, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.
@@
 
-We can now visualize how these 159 descendants are related to Sir David Cox by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}.
+We can now visualize how these 159 ``descendants" are related to Sir David Cox by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}.
 As such, we create Figure \ref{fig:dCox} using the code below.
@@
     \includegraphics[width=\textwidth]{dCox.png}
-    \caption{The 159 academic statistician descendants of Sir David Cox.}
+    \caption{The 159 academic statistician ``descendants" of Sir David Cox.}
     \label{fig:dCox}
@@
 
-We see from Figure \ref{fig:dCox} that Sir David Cox had 42 children, many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
-Of his children, the one who produced the most children of their own was Peter Bloomfield, who has 26 children and 49 descendants.
+We see from Figure \ref{fig:dCox} that Sir David Cox had 42 ``children", many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
+Of his ``children", the one who produced the most ``children" of their own was Peter Bloomfield, who has 26 ``children" and 49 ``descendants".
 In total, Sir David Cox had five generations of academic statistics mentees in this dataset.
@@
 
-At this point, it would be insightful to examine a more detailed view of one of the longest strings of parent-child relationships between Sir David Cox and one of the two individuals who are his fifth generation descendants.
-We do so with the code below, choosing his fifth generation descendant to be Petra Buzkova.
+At this point, it would be insightful to examine a more detailed view of one of the longest strings of ``parent-child" relationships between Sir David Cox and one of the two individuals who are his fifth generation ``descendants".
+We do so with the code below, choosing his fifth generation ``descendant" to be Petra Buzkova.
 
@@
     \includegraphics[width=\textwidth]{pathCB}
-    \caption{The shortest path between Sir David Cox and one of his fifth generation descendants, Petra Buzkova.}
+    \caption{The shortest path between Sir David Cox and one of his fifth generation ``descendants", Petra Buzkova.}
     \label{fig:pathCB}
@@
 This code results in Figure \ref{fig:pathCB}.
-We see that the shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional parent-child relationships that span about 55 years.
-We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than about 30 years in degree acquisition, whereas the three statisticians who earned their degrees later in this path only span less than about ten years in degree acquisition.
+We see that the shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional ``parent-child" relationships that span about 55 years.
+We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than 30 years in degree acquisition, whereas the three statisticians who earned their degrees later in this path only span less than ten years in degree acquisition.
 
@@
 We see from the resulting Figure \ref{fig:plotCBText} that almost all text labels for individuals who received their graduate-level statistics degrees between 1950 and 2015 are undecipherable.
-We also see that the year Sir David Cox acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for statistics degree acquisitions in this dataset occur around 1860.
+We also see that the year Sir David Cox acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for acquisition of statistics degrees in this dataset occur around 1860.
 However, the number of individuals who are documented as receiving their statistics degrees between 1860 and 1950 are few enough so that their text labels are somewhat readable.
@@
 We can also more clearly see that, even though only ten years span between the youngest individual in the genealogy and Petra Buzkova, there are many individuals in that last decade.
-Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of statistics degree acquisitions.
+Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of acquisition of statistics degrees.
 This figure is interactive with the use of Plotly (\citealt{plotly}).
@@
 
-The authors are grateful for the financial support from the United Soybean Board, North Central Soybean Research Program, the Iowa Soybean Association, the NSF Plant Genome Research Program (award number 0820642), and the USDA-Agricultural Research Service.
+The authors thank Drs.
+James E.
+Spect and Randy C.
+Shoemaker for helpful discussions of soybean genealogy.
+In addition, the authors are grateful for the financial support from the United Soybean Board (Project 1204), The North Central Soybean Research Program, the NSF Plant Genome Research Program (award number 0820642), and the USDA-ARS CRIS Project 3625-21220-005-00D.
+The USDA is an equal opportunity provider and employer.
+Mention of trade names or commercial products in this article is solely for the purpose of providing specific information and does not imply recommendation or endorsement by the U.S.
+Department of Agriculture.
 

### 2016-04-20 JSS/jss-style/article.tex
shiny app

@@
 
-It may also be of interest to generate matrices where the colors indicates a variable (such as the degree of the shortest path) between all pairwise combinations of inputted varieties.
+It may also be of interest to generate matrices where the colors indicates a variable between all pairwise combinations of inputted varieties.
 The package \pkg{ggenealogy} also provides a function \code{plotDegMatrix()} for that purpose.
+We can demonstrate this function with the variable being the shortest path degree between a given pair of varieties.
+The shortest path degree is calculated as the smallest number of parent-child edges needed to traverse between two varieties of interest.
+For instance, in Figure \ref{fig:pathTNZB}, the shortest path degree between \code{Tokyo} and \code{Narow} is four and the shortest path degree between \code{Bedford} and \code{Zane} is ten.
 
@@
 We see that the degree of the shortest path between varieties \code{Bedford} and \code{Zane} is 10, which is consistent with what we saw earlier in Figure \ref{fig:pathTNZB}.
-However, we now also see that a degree of 10 may be considered a large degree for this dataset.
+However, we now also see that a shortest path degree of 10 may be considered relative to the rest of this dataset.
 
@@
 Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of acquisition of statistics degrees.
-This figure is interactive with the use of Plotly (\citealt{plotly}).
 
@@
 
-Incorporation of the Shiny application would allow users to examine \pkg{ggenealogy} tools in a more interactive way (\citealt{shiny}).
-The reactive programming would save them the time of using command-line for each change of input as well as the inefficiency of rerunning code.
+Incorporation of the Shiny application allows users to examine \pkg{ggenealogy} tools in a more interactive way (\citealt{shiny}).
+The reactive programming saves them the time of using command-line for each change of input as well as the inefficiency of rerunning code.
+A Shiny application that uses certain \pkg{ggenealogy} functionality is available for users who wish to explore the soybean genealogy; the data can be viewed at \url{http://shiny.soybase.org/CNV/}.
+
 We also aim to incorporate plotting tools that can examine not only quantitative variables (such as our example variable of ``year"), but also categorical variables associated with individuals in datasets.
@@
 For example, after comparing several bin sizes on the soybean genealogy, we determined that the bin size of 6 produced the minimal textual overlap, as is seen in Figure \ref{fig:plotTNBin6}.
-If we could subsequently fine-tune the vertical positions of the small fraction of text labels that remained overlapped after the automated \pkg{gggenealogy} function, then we could potentially remove all overlaps, and the plot could be used in presentations and publications.
+If we could subsequently fine-tune the vertical positions of the small fraction of text labels that remained overlapped after the automated \pkg{ggenealogy} function, then we could potentially remove all overlaps, and the plot could be used in presentations and publications.
 Of course, it is impossible to eliminate textual overlap in larger datasets (see Figure \ref{fig:plotCBText}).

### 2016-04-22 JSS/jss-style/article.tex
statement about release year

@@
         Dianne Cook\\Monash University\And
-        Michelle Graham\\Iowa State University}
+        Michelle A. Graham\\Iowa State University}
 \title{\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data}
@@
 Each row contains information about a particular child soybean variety, including the name of the child, its yield, the year it was released, whether or not its release year was imputed, and the name of its parent.
+It should be noted that it typically requires many crosses over the span of one to two decades to develop a new variety that has introduced a desired trait and/or removed an undesired trait.
+Hence, the release year variable in this dataset represents the year in which the variety was released to the public after its development period.
 While the name of the child is required, the other four columns can have missing values (which are represented in \proglang{R} with the symbol NA for ``not available").

### 2016-04-28 JSS/jss-style/article.tex
\key in .bib file
pathTNZB figure fontFace

@@
 \begin{Code}
-R> plotPath(pathTN)
+R> plotPath(pathTN, fontFace = 2)
 \end{Code}
@@
 We can then create a plot showing the shortest path between these two varieties of interest.
+As this is a longer path, we may also consider setting the \code{fontFace} variable of the \code{plotPath()} to a value of 2, indicating we wish to boldface the two varieties of interest.
 
@@
 R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
-R> plotPath(pathBZ)
+R> plotPath(pathBZ, fontFace = 2)
 \end{Code}
@@
 James E.
-Spect and Randy C.
+Specht and Randy C.
 Shoemaker for helpful discussions of soybean genealogy.
@@
 Department of Agriculture.
+Author Cook was a faculty member at Iowa State University at the time that most of this work was conducted.
 

### 2016-04-29 JSS/jss-style/article.tex
ggplot2 add on consistency

@@
 \begin{Code}
-R> plotPath(pathTN, fontFace = 2)
+R> plotPath(pathTN)
 \end{Code}
@@
 We do so with the code below, choosing his fifth generation ``descendant" to be Petra Buzkova.
+We set the \code{fontFace} variable of the \code{plotPath()} to a value of 4, indicating we wish to boldface and italicize the two varieties of interest.
 
@@
   isDirected = FALSE)
-R> plotPath(pathCB) + ggplot2::theme(axis.text = ggplot2::element_text(size
-  = 12), axis.title = ggplot2::element_text(size = 12)) +
-  ggplot2::scale_x_continuous(breaks = seq(1940, 2010, by = 10))
+R> plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text =
+   ggplot2::element_text(size = 10), axis.title =
+   ggplot2::element_text(size = 10))
 \end{CodeInput}
@@
     \centering
-    \includegraphics[width=\textwidth]{pathCB}
+    \includegraphics[width=0.7\textwidth]{pathCB}
     \caption{The shortest path between Sir David Cox and one of his fifth generation ``descendants", Petra Buzkova.}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::
-  scale_x_continuous(limits = c(1850, 2030)) + ggplot2::theme(axis.text =
-  ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size
-  = 12))
+R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) +
+   ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+   ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+   (expand = c(.1, .2))
 \end{CodeInput}
@@
 The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{binVector} will not solve the text overlapping problem.
-Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying a value of \code{TRUE} for the input parameter \code{nodeLabel}.
-As is further detailed in the help manual of the function, doing so will generate a plot where only text labels for individuals on the path of interest are labeled; all other individual labels are represented as dots.
-We provide the code for this below, which results in Figure \ref{fig:plotCBNoText}.
+Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes thare not on the path of interest.
+Moreover, we can make the edges that are not on the path of interest to be represented in a less noticeable color (light gray) than the edges that are on the path of interest (default of black).
+The variable names and options for these aesthetics is further detailed in the help manual of the function.
+We provide one example code that alters the defaults of the text color and sizes of nodes and edges below, which results in Figure \ref{fig:plotCBNoText}.
 
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = 4,
-  pathNodeSize = 4, nodeLabel=FALSE) + ggplot2::theme(axis.text = ggplot2::
-  element_text(size = 12), axis.title = ggplot2::element_text(size = 12)) + 
-  ggplot2::xlab(1850, 2025)
+R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
+   pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
+   ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+   ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+   (expand = c(.1, .2))
 \end{CodeInput}

### 2016-04-29 JSS/jss-style/article.tex
switch dots to labels; add animate variable

@@
 Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes thare not on the path of interest.
-Moreover, we can make the edges that are not on the path of interest to be represented in a less noticeable color (light gray) than the edges that are on the path of interest (default of black).
+Moreover, we can make the edges that are not on the path of interest to be represented in a less noticeable color (light gray) than the edges that are on the path of interest (default of dark green).
 The variable names and options for these aesthetics is further detailed in the help manual of the function.
@@
     \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.
-Individuals on the shortest path have text labels; all other individuals have dot labels.}
+Individuals on the shortest path are labeled in large and black text and connected by dark green edges; all other individuals are labeled in small and gray text and connected by light gray edges.}
     \label{fig:plotCBNoText}
@@
 Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of acquisition of statistics degrees.
+
+We could still improve upon Figure \ref{fig:plotCBNoText}.
+Even though we may be primarily interested in understanding how the path of interest is overlaid across the entire genealogical structure, we could, upon viewing the entire structure, also develop an interest in nodes that are not on the path of interest but are revealed to stand out among the rest of the genealogical structure.
+For instance, in Figure \ref{fig:plotCBNoText}, it may be of interest for us to determine the names of the few individuals who obtained their statistics degrees before 1900.
+Fortunately, within the \code{plotPathOnAll()} function, there is a variable \code{animate} that we can set to a value of TRUE to create an interactive version of the figure that allows us to hover over individual illegible labels and immediately receive their labels in a readable format.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
+   pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate =
+   TRUE)
+\end{CodeInput}
+\end{CodeChunk}
 

### 2016-05-01 JSS/jss-style/article.tex
non-repeating nodes plot updated

@@
 
-This generates the plot in the left side of Figure \ref{fig:Lee}.
+This generates the top plot of Figure \ref{fig:Lee}.
 We notice that \code{Lee} has 3 generations of ancestors and 5 generations of descendants.
@@
 
-The right side of Figure \ref{fig:Lee} is not an output plot of \pkg{ggenealogy}.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{LeeAD3}
+    \caption{Top: All ancestors and descendants of the variety Lee are shown in this ggenealogy plot.
+Bottom: We now attempt to mimic the blue paths in the ggenealogy plot on the top, only now nodes cannot be repeated.
+The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee.
+The presence of red parentheses indicate that the plot on the bottom ambiguously display the example soybean genealogy in the way that the ggenealogy plot on the top can accomplish.}
+    \label{fig:Lee}
+\end{figure}
+
+The bottom plot of Figure \ref{fig:Lee} is not an output plot of \pkg{ggenealogy}.
 Instead, it was simply created for didactic purposes.
-Here, the paths that were manually highlighted in blue in the left plot produced by \pkg{ggenealogy} are shown again, only now nodes cannot be repeated.
+Here, the paths that were manually highlighted in blue in the top plot produced by \pkg{ggenealogy} are shown again, only now nodes cannot be repeated.
 The parenthetical number above each node represents the set of generation counts distancing that node from the center node \code{Lee}; green parentheses indicate that the node could be successfully placed in one horizontal position, but red parentheses indicate that the node could not be successfully placed in one horizontal position.
 We see that node \code{TN89-39} cannot simultaneously be represented as both a third and fourth descendent of node \code{Lee}, and node \code{5601T} cannot simultaneously be represented as a third, fourth, and fifth descendent of node \code{Lee}.
-Hence, without allowing nodes to repeat, this dataset cannot be presented in the graph on the right as it can be in the \pkg{ggenealogy} graph on the left.
+Hence, without allowing nodes to repeat, this dataset cannot be presented in the graph on the bottom as it can be in the \pkg{ggenealogy} graph on the top.
 This is a current limitation in other genealogy and graphical software that \pkg{ggenealogy} can now provide.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{Lee}
-    \caption{Left: All ancestors and descendants of the variety Lee are shown in this ggenealogy plot.
-Right: We now attempt to mimic the blue paths in the ggenealogy plot on the left, only now nodes cannot be repeated.
-The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee.
-The presence of red parentheses indicate that the plot on the right ambiguously display the example soybean genealogy in the way that the ggenealogy plot on the left can accomplish.}
-    \label{fig:Lee}
-\end{figure}
 

### 2016-05-01 JSS/jss-style/article.tex
Adobe Acrobat Reader DC in caption

@@
 \usepackage{float}
+\usepackage[]{media9}
 
@@
 Fortunately, within the \code{plotPathOnAll()} function, there is a variable \code{animate} that we can set to a value of TRUE to create an interactive version of the figure that allows us to hover over individual illegible labels and immediately receive their labels in a readable format.
+A short video demonstration of these interactive features can be viewed upon clicking on Figure \ref{fig:plotAnimate}.
 
@@
 \end{CodeChunk}
+
+\begin{figure}[H]
+    \centering
+    \includemedia[width=1.0\textwidth, addresource=animateGen.mov, deactivate=onclick, flashvars={source=animateGen.mov}]{\includegraphics{./plotCBNoText.png}}{VPlayer.swf}
+    \caption{Upon clicking on this figure twice, a short video demonstrating the animation features for this function can be viewed.
+Please note that to properly view this video, the PDF version of this paper must be opened in Adobe Acrobat Reader DC, which can be downloaded free of charge.}
+    \label{fig:plotAnimate}
+\end{figure}
 

### 2016-05-09 JSS/jss-style/article.tex
grammar

@@
   nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
-  element_text(size = 12), axis.title =ggplot2::element_text(size = 12))
+  element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 \end{Code}
@@
 The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{binVector} will not solve the text overlapping problem.
-Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes thare not on the path of interest.
+Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes that are not on the path of interest.
 Moreover, we can make the edges that are not on the path of interest to be represented in a less noticeable color (light gray) than the edges that are on the path of interest (default of dark green).

### 2016-05-17 JSS/jss-style/article.tex
play/pause toggle for \mediabutton

@@
 \usepackage[]{media9}
+
+\usepackage{tcolorbox}% http://ctan.org/pkg/tcolorbox
+\definecolor{frameCol}{rgb}{0.122, 0.435, 0.698}% Rule colour
+\definecolor{bgndCol}{rgb}{.66,.66,.66}% Rule colour
+\makeatletter
+\newcommand{\mybox}[1]{%
+  \setbox0=\hbox{#1}%
+  \setlength{\@tempdima}{\dimexpr\wd0+13pt}%
+  \begin{tcolorbox}[colframe=frameCol,colback=bgndCol,boxrule=0.5pt,arc=4pt,
+      left=6pt,right=6pt,top=6pt,bottom=6pt,boxsep=0pt,width=\@tempdima]
+    #1
+  \end{tcolorbox}
+}
+\makeatother
 
@@
     \centering
-    \includemedia[width=1.0\textwidth, addresource=animateGen.mov, deactivate=onclick, flashvars={source=animateGen.mov}]{\includegraphics{./plotCBNoText.png}}{VPlayer.swf}
+    \includemedia[label = v1, playbutton=none, width=1.0\textwidth, addresource=animateGen.mov, activate=pageopen, flashvars={source=animateGen.mov}]{\includegraphics{./plotCBNoText.png}}{VPlayer.swf}
+    \mediabutton[
+mediacommand=v1:playPause,
+]{\mybox{Play/Pause}}
     \caption{Upon clicking on this figure twice, a short video demonstrating the animation features for this function can be viewed.

### 2016-05-17 JSS/jss-style/article.tex
video caveats in caption

@@
     \centering
-    \includegraphics[width=0.65\textwidth]{kinshipFig}
+    \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
     \caption{Example pedigree chart from the \proglang{kinship2} package, where the vertical axis denotes generation count.
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{Graph}
+    \fbox{\includegraphics[width=0.6\textwidth]{Graph}}
     \caption{Example genealogical display using popular graph software like \proglang{GraphViz} and \proglang{Cytoscape}, with generation count denoted by the vertical axis.
@@
     \centering
-    \includegraphics[width=\textwidth]{pathTNZB}
+    \fbox{\includegraphics[width=\textwidth]{pathTNZB}}
     \caption{Left: The shortest path between varieties Tokyo and Narow is strictly composed of a unidirectional sequence of parent-child relationships.
@@
     \centering
-    \includegraphics[width=\textwidth]{plotTNBin3}
+    \fbox{\includegraphics[width=\textwidth]{plotTNBin3}}
     \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
@@
     \centering
-    \includegraphics[width=\textwidth]{plotTNBin6}
+    \fbox{\includegraphics[width=\textwidth]{plotTNBin6}}
     \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
@@
     \centering
-    \includegraphics[width=\textwidth]{LeeAD3}
+    \fbox{\includegraphics[width=\textwidth]{LeeAD3}}
     \caption{Top: All ancestors and descendants of the variety Lee are shown in this ggenealogy plot.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{degMatrix}
+    \fbox{\includegraphics[width=0.5\textwidth]{degMatrix}}
     \caption{The shortest path degree matrix between ten varieties of interest.}
@@
     \centering
-    \includegraphics[width=\textwidth]{dCox.png}
+    \fbox{\includegraphics[width=\textwidth]{dCox.png}}
     \caption{The 159 academic statistician ``descendants" of Sir David Cox.}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{pathCB}
+    \fbox{\includegraphics[width=0.7\textwidth]{pathCB}}
     \caption{The shortest path between Sir David Cox and one of his fifth generation ``descendants", Petra Buzkova.}
@@
     \centering
-    \includegraphics[width=\textwidth]{plotCBText}
+    \fbox{\includegraphics[width=\textwidth]{plotCBText}}
     \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.}
@@
     \centering
-    \includegraphics[width=\textwidth]{plotCBNoText}
+    \fbox{\includegraphics[width=\textwidth]{plotCBNoText}}
     \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.
@@
 Fortunately, within the \code{plotPathOnAll()} function, there is a variable \code{animate} that we can set to a value of TRUE to create an interactive version of the figure that allows us to hover over individual illegible labels and immediately receive their labels in a readable format.
-A short video demonstration of these interactive features can be viewed upon clicking on Figure \ref{fig:plotAnimate}.
+A short video demonstration of these interactive features can be viewed below.
 
@@
     \centering
-    \includemedia[label = v1, playbutton=none, width=1.0\textwidth, addresource=animateGen.mov, activate=pageopen, flashvars={source=animateGen.mov}]{\includegraphics{./plotCBNoText.png}}{VPlayer.swf}
-    \mediabutton[
-mediacommand=v1:playPause,
-]{\mybox{Play/Pause}}
-    \caption{Upon clicking on this figure twice, a short video demonstrating the animation features for this function can be viewed.
-Please note that to properly view this video, the PDF version of this paper must be opened in Adobe Acrobat Reader DC, which can be downloaded free of charge.}
+    \fbox{\includemedia[label = v1, playbutton=none, width=1.0\textwidth, addresource=animateGen.mov, activate=pageopen, flashvars={source=animateGen.mov}]{\includegraphics{./plotCBNoText.png}}{VPlayer.swf}}
+    \mediabutton[mediacommand=v1:playPause,]{\mybox{Play/Pause}}
+    \caption{A short video demonstrating the animation features for this function.
+Please note that to properly view this video, the PDF version of this document must be opened in Adobe Acrobat Reader DC (version >=9), which can be downloaded free of charge.
+This video can only be viewed on Windows or Mac OS X platforms; it cannot be viewed on mobile devices.}
     \label{fig:plotAnimate}

### 2016-05-20 JSS/jss-style/article.tex
replication script

@@
 
-\section{Available Software}
+\section{Available software}
 
@@
 
-The \proglang{R} package \proglang{pedigree} is named after the standardized chart used to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}).
+The \proglang{R} package \pkg{pedigree} is named after the standardized chart used to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}).
 This package does provide tools that perform methods on parent-child datasets, such as rapidly determining the generation count for each member in the pedigree.
@@
 
-Another \proglang{R} package called \proglang{kinship2} does produce basic pedigree charts (\citealt{kin}).
-In Figure \ref{fig:kinshipFig}, we provide an example pedigree chart from the \proglang{kinship2} package vignette.
+Another \proglang{R} package called \pkg{kinship2} does produce basic pedigree charts (\citealt{kin}).
+In Figure \ref{fig:kinshipFig}, we provide an example pedigree chart from the \pkg{kinship2} package vignette.
 This pedigree chart adheres to the standard set of symbols used for visualizing genealogical structures: Males are represented with squares and females with circles.
@@
 
-We demonstrate how the standardized pedigree charts in the \proglang{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational breeding case in Figure \ref{fig:kinshipFig}.
+We demonstrate how the standardized pedigree charts in the \pkg{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational breeding case in Figure \ref{fig:kinshipFig}.
 In that figure, each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202.
@@
     \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
-    \caption{Example pedigree chart from the \proglang{kinship2} package, where the vertical axis denotes generation count.
+    \caption{Example pedigree chart from the kinship2 package, where the vertical axis denotes generation count.
 We superimposed green-highlighted individual 215 for explanatory purposes.
@@
 
-In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}).
+In addition, popular graph drawing software such as \pkg{GraphViz} and \pkg{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}).
 Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated.
@@
 
-We again illustrate this problem in Figure \ref{fig:Graph} with an example genealogy using popular graph drawing software like \proglang{GraphViz} and \proglang{Cytoscape}.
+We again illustrate this problem in Figure \ref{fig:Graph} with an example genealogy using popular graph drawing software like \pkg{GraphViz} and \pkg{Cytoscape}.
 Here, generation count is denoted by the vertical axis.
@@
     \fbox{\includegraphics[width=0.6\textwidth]{Graph}}
-    \caption{Example genealogical display using popular graph software like \proglang{GraphViz} and \proglang{Cytoscape}, with generation count denoted by the vertical axis.
+    \caption{Example genealogical display using popular graph software like \pkg{GraphViz} and \pkg{Cytoscape}, with generation count denoted by the vertical axis.
 As was shown in Figure \ref{fig:kinshipFig}, the green node has parents from two different generations, and hence must be ambiguously positioned as one of two generation counts.}
@@
 
-\section{Example Datasets}
+\section{Example datasets}
 \label{exData}
@@
 
-\subsection{Soybean Genealogy}
+\subsection{Soybean genealogy}
 
@@
 R> library("ggenealogy")
-R> data(sbGeneal)
+R> data("sbGeneal")
 R> str(sbGeneal)
@@
 
-\subsection{Academic Genealogy of Statisticians}
+\subsection{Academic genealogy of statisticians}
 
@@
 \begin{CodeInput}
-R> data(statGeneal)
+R> data("statGeneal")
 R> dim(statGeneal)
@@
 
-\section{Genealogical Input Format}
+\section{Genealogical input format}
 
@@
 
-\section{Generating a Graphical Object}
+\section{Generating a graphical object}
 
@@
 
-\section{Plotting a Shortest Path}
+\section{Plotting a shortest path}
 
@@
 
-\section{Superimposing Shortest Path on Tree}
+\section{Superimposing shortest path on tree}
 
@@
 
-\section{Plotting Ancestors and Descendants by Generation}
+\section{Plotting ancestors and descendants by generation}
 \label{remedy}
@@
 
-\section{Plotting Distance Matrix}
+\section{Plotting distance matrix}
 
@@
 
-\section{Academic Genealogy of Statisticians}
+\section{Academic genealogy of statisticians}
 
@@
 \begin{CodeInput}
->R library(dplyr)
+>R library("dplyr")
 >R indVec <- getNodes(statGeneal)
@@
 Fortunately, within the \code{plotPathOnAll()} function, there is a variable \code{animate} that we can set to a value of TRUE to create an interactive version of the figure that allows us to hover over individual illegible labels and immediately receive their labels in a readable format.
+This interactive functionality comes from methods in the \pkg{plotly} package (\citealt{plotly}).
 A short video demonstration of these interactive features can be viewed below.
@@
 
-\section{Future Avenues}
+\section{Future avenues}
 

### 2016-05-20 JSS/jss-style/replicationScript.Rnw
replication script

@@
 \parindent 0pt
-\title{Replication Script to JSS Submission:\\
-\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
-\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
+\title{Replication Script to JSS Submission}
+\begin{document}
+
+\author{Lindsay Rutter, Susan VanderPlas, Dianne Cook, Michelle A. Graham}
 \date{\today}
-\begin{document}
 \maketitle
@@
 \tableofcontents
-<<>>=
-install.packages("ggenealogy")
-library("ggenealogy")
-@
+
 %%%%%%%%%%%%%%%
-%% Section 1 %%
+%% Section 3 %%
 %%%%%%%%%%%%%%%
-\section{Generating a Graphical Object}
+\section*{3. Example datasets}
 
-Examine structure of sbGeneal example dataset:
+First, we can load and examine the structure of the soybean genealogy example dataset (called sbGeneal):
 
 <<>>=
-data(sbGeneal)
+rm(list=)
+library("ggenealogy")
+data("sbGeneal")
 str(sbGeneal)
+@
+
+After that, we can load and examine the structure of the academic statistician genealogy example dataset (called statGeneal):
+
+<<>>=
+data("statGeneal")
+dim(statGeneal)
+colnames(statGeneal)
 @
@@
 
-<<>>=
-library("mstate")
-tra2 <- !is.na(transMat(list(c(2), c(1))))
-tra2
-@
-Use \texttt{mplskeleton} to construct a skeleton of the model parameter list \texttt{mpl}:
-<<>>=
-mpl <- mplskeleton(tmat = tra2)
-str(mpl)
-@
-Parametrize the model and complete \texttt{mpl}:
-<<>>=
-mpl[[1]]$bhr[[2]] <- mpl[[2]]$bhr[[1]] <- function(t){return(0.5)}
-## time-varying x2 and time-varying f(x2):
-mpl[[1]]$eta[[2]] <- function(x.i, t){
-  ifelse(t < 5, 
-         return(1.0 * x.i[1] + 0.5 * x.i[2]), 
-         return(1.0 * x.i[1] + 1.0 * x.i[3]))}
-## time-varying x2 and time-varying f(x1):
-mpl[[2]]$eta[[1]] <- function(x.i, t){
-  ifelse(t < 5, 
-         return(-0.5 * x.i[1] + 0.5 * x.i[2]), 
-         return( 1.0 * x.i[1] + 0.5 * x.i[3]))}
-@
-Simulate event histories:
-<<>>=
-set.seed(123)
-N <- 500
-X <- matrix(nrow = N, ncol = 2, rnorm(2 * N))
-X <- cbind(X, X[, 2] + runif(N)/10)
-colnames(X) <- c("x1", "x2.t1", "x2.t2")
-head(X)
-Xstruc <- matrix(ncol = 2, nrow = 2, data = 0)
-rownames(Xstruc) <- c("t1", "t2")
-colnames(Xstruc) <- c("x1", "x2")
-Xstruc[, 1] <- 1
-Xstruc[, 2] <- c(2, 3)
-Xstruc
-d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
-                       change.times = c(5), Xstruc = Xstruc)
-head(d$msm.basics)
-head(d$ttsce[, 1:7])
-head(d$tt.indicators)
-@
-Nelson-Aalen estimator of the cumulative hazard rate functions:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
-plotnae(d = d$msm.basics, mpl = mpl, xlim = c(0, 15))
-abline(a = 0, b = 0.5, col = "darkgrey")
-@
-
-Stratified partial likelihood modelling:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
-m <- coxph(Surv(entry, exit, delta) ~ x1.12.t1 + x2.12.t1 + x1.12.t2 + 
-             x2.12.t2 + x1.21.t1 + x2.21.t1 + x1.21.t2 + x2.21.t2 + 
-             strata(trans), data = data.frame(d$msm.basics, d$ttsce))
-round(summary(m)$coef[, c(1, 5)], 5)
-plotcph(m)
-points(c(c(1, 0.5, 1 - 1, 1 - 0.5, -0.5, 0.5, 1 - (-0.5), 0.5 - 0.5)), 
-       pch = 16, col = "darkgrey")
-@
-
-Breslow estimate for cumulative hazard:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=be>>=
-plotbe(m, mpl, xlim = c(0, 15), ylim = c(0, 10))
-abline(a = 0, b = 0.5, col = "darkgrey")
-@
-%%%%%%%%%%%%%%%
-%% Section 2 %%
-%%%%%%%%%%%%%%%
-\section{Illness-death Model with Recovery, a Non-linear Baseline Hazarad Rate Function and Non-linear Covariate Effects}
-Prepare the simulation of event histories:
-<<>>=
-library("simMSM")
-traIDM <- matrix(nrow = 3, ncol = 3, FALSE)
-traIDM[1, 2] <- traIDM[1, 3] <- traIDM[2, 1] <- traIDM[2, 3] <- TRUE
-## Or use mstate:
-library("mstate")
-traIDM <- !is.na(transMat(list(c(2, 3), c(1, 3), c())))
-traIDM
-## Build up mpl:
-mpl <- mplskeleton(tmat = traIDM)
-mpl[[1]]$bhr[[2]] <- function(t){0.5 + 0.4 * sin(t)}
-mpl[[1]]$bhr[[3]] <- mpl[[2]]$bhr[[1]] <- 
-  mpl[[2]]$bhr[[3]] <- function(t){0.5}
-mpl[[1]]$eta[[2]] <- function(x.i, t){ 3 * (x.i^2)}
-mpl[[1]]$eta[[3]] <- function(x.i, t){-3 * (x.i^2)}
-mpl[[2]]$eta[[1]] <- mpl[[2]]$eta[[3]] <- 
-  function(x.i, t){0.5 * x.i}
-## simulate design matrix and event history data:
-set.seed(123)
-N <- 500
-X <- matrix(ncol = 1, nrow = N, data = runif(N) - 0.5)
-colnames(X) <- c("x")
-Xstruc <- matrix(ncol = 1, nrow = 1, data = 1)
-rownames(Xstruc) <- c("t1")
-colnames(Xstruc) <- c("x")
-Xstruc
-d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
-                       change.times = NULL, Xstruc = Xstruc)
-head(d$msm.basics)
-@
-
-Nelson-Aalen estimators:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
-plotnae(d = d$msm.basics, mpl)
-@
-
-Convert to long data format and estimate a stratified Cox proportional hazards (partial likelihood) model:
-<<>>=
-ld <- tolongformat(d, mpl = mpl)
-head(ld$msm.basics)
-head(ld$tt.indicators)
-m <- coxph(Surv(entry, exit, delta) ~ I(x.12^2) + I(x.13^2) + 
-             x.21 + x.23 + strata(trans), 
-           data = data.frame(ld$msm.basics, ld$ttsce))
-round(summary(m)$coef[, c(1, 5)], 5)
-@
-
-Cox model coefficient illustration:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
-plotcph(m)
-points(c(3, -3, 0.5, 0.5), col = "darkgrey", pch = 16)
-@
-
-Breslow estimator:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
-plotbe(m, mpl)
-@
-
-BayesX:
-<<>>=
-d <- data.frame(d$msm.basics, d$tt.indicators)
-install.packages("BayesXsrc")
-library("BayesXsrc")
-p <- "delimiter = ;
-dataset dat;
-dat.infile using d.raw;
-remlreg simmsm;
-simmsm.mregress 
-trans12=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
-trans13=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
-trans21=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
-trans23=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2),
-family=multistate lefttrunc=entry state=from maxit=1000 using dat;
-delimiter = return;"
-write.table(d, row.names = FALSE, quote = FALSE, file = "d.raw")
-writeLines(p, con = "prg")
-run.bayesx(prg = "prg", verbose = FALSE)
-@
-
-Prepare plotting the results:
-<<>>=
-setwd("./output/")
-log.bhr <- beta.fixed <- beta.x <- vector("list", 4)
-log.bhr[[1]] <- read.table("simmsm_f_exit_logbaseline.res", header=T)
-log.bhr[[2]] <- read.table("simmsm_f_2_exit_logbaseline.res", header=T)
-log.bhr[[3]] <- read.table("simmsm_f_3_exit_logbaseline.res", header=T)
-log.bhr[[4]] <- read.table("simmsm_f_4_exit_logbaseline.res", header=T)
-beta.fixed[[1]] <- read.table("simmsm_FixedEffects.res", header=T)
-beta.fixed[[2]] <- read.table("simmsm_FixedEffects_2.res", header=T)
-beta.fixed[[3]] <- read.table("simmsm_FixedEffects_3.res", header=T)
-beta.fixed[[4]] <- read.table("simmsm_FixedEffects_4.res", header=T)
-beta.x[[1]] <- read.table("simmsm_f_x_pspline.res", header=T)
-beta.x[[2]] <- read.table("simmsm_f_2_x_pspline.res", header=T)
-beta.x[[3]] <- read.table("simmsm_f_3_x_pspline.res", header=T)
-beta.x[[4]] <- read.table("simmsm_f_4_x_pspline.res", header=T)
-setwd("../")
-plot.bhr <- function(log.bhr, beta.fixed, t.max){
-  index.for.plot <- c("12,0", "13,0", "21,0", "23,0")
-  par(mfrow = c(2, 2), mar = c(3.1, 3.1, 0.1, 0.1), tcl = -0.3)
-  for(q in 1:4){
-    plot(1, 1, ylim=c(0, 3), xlim=c(0, t.max), 
-         type = "n", xaxt = "n", yaxt = "n", 
-         xlab = "", ylab = "", bty = "n")
-    axis(1, at = c(0, 4, 8, 12), label = FALSE)
-    mtext(side = 1, text = c(0, 4, 8, 12), at = c(0, 4, 8, 12), 
-          line = 0.5, cex = 0.5)
-    axis(2, label = FALSE, at = 0:2)
-    mtext(side = 2, text = 0:2, at = 0:2, line = 0.5, cex = 0.5)
-    mtext("t", side = 1, line = 1.4, cex = 0.7)
-    mtext(bquote(paste(lambda[.(index.for.plot[q])],(t))), side = 2, 
-          line = 1.4, cex = 0.7)
-    hilf <- beta.fixed[[q]]
-    if(q == 1){
-      ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
-      intercept <- hilf$pmode[1] - 0.255
-    }
-    if(q == 2){
-      ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
-      intercept <- hilf$pmode[1] + 0.255
-    }
-    if(q %in% c(3, 4)){
-      intercept <- hilf$pmode[1]
-    }
-    hilf <- log.bhr[[q]]
-    polygon(c(hilf$exit, rev(hilf$exit)), 
-            exp(intercept + c(hilf$ci95lower, rev(hilf$ci95upper))), 
-            border = "NA", col = "lightgrey")
-    lines(hilf$exit, exp(hilf$pmode + intercept), col = 1, lwd = 2)
-    time <- seq(0, t.max, length = 100)
-    if(q == 1){lines(time, (0.5 + 0.4 * sin(time)), lwd = 2, lty = 2)}
-    if(q == 2){lines(time, (0.5 + 0*time), lwd = 2, lty = 2)}
-    if(q == 3){lines(time, 0.5 + 0*time, lwd = 2, lty = 2)}
-    if(q == 4){lines(time, 0.5 + 0*time, lwd = 2, lty = 2)}
-  }
-}
-plot.spline <- function(beta.x){
-  index.for.plot <- c("12,x", "13,x", "21,x", "23,x")
-  par(mfrow = c(2, 2), mar = c(3.1, 3.1, 0.1, 0.1), tcl = -0.3)
-  for(q in 1:4){
-    plot(1, 1, ylim = c(-1.2, 1.2), xlim=c(-0.5, 0.5), type = "n", 
-         xaxt = "n", yaxt = "n", xlab = "", ylab = "", bty = "n")
-    axis(1, at = c(-0.5, 0, 0.5), label = FALSE)
-    mtext(side = 1, text = c(-0.5, 0, 0.5), at = c(-0.5, 0, 0.5), 
-          line = 0.5, cex = 0.5)
-    axis(2, label = FALSE, at = -1:1)
-    mtext(side = 2, text = -1:1, at = -1:1, line = 0.5, cex = 0.5)
-    mtext("x", side = 1, line = 1.4, cex = 0.7)
-    mtext(bquote(f[.(index.for.plot[q])](x)), side = 2, 
-          line = 1.4, cex = 0.7)
-    if(q == 1){
-       ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
-      intercept <-  0.255
-    }
-    if(q == 2){
-      ## -0.255 = mean(-3 * (seq(-0.5, 0.5, length = 100)^2))
-      intercept <- -0.255
-    }
-    if(q %in% c(3, 4)){
-      intercept <- 0
-    }
-    hilf <- beta.x[[q]]
-    polygon(c(hilf$x, rev(hilf$x)), 
-            intercept + c(hilf$ci95lower, rev(hilf$ci95upper)), 
-            border = "NA", col = "lightgrey")
-    lines(hilf$x, hilf$pmode + intercept, col = 1, lwd = 2)
-    x <- seq(-0.5, 0.5, length = 100)
-    if(q == 1){lines(x, 3*(x^2), lwd = 2, lty = 2)}
-    if(q == 2){lines(x,-3*(x^2), lwd = 2, lty = 2)}
-    if(q == 3){lines(x, 0.5 * x, lwd = 2, lty = 2)}
-    if(q == 4){lines(x, 0.5 * x, lwd = 2, lty = 2)}
-  }
-}
-@
-
-BayesX baseline hazard rate function estimates:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=bayesX_bhr>>=
-plot.bhr(log.bhr, beta.fixed, t.max = print(max(d$exit)))
-@
-
-BayesX covariate effect function estimates:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=bayesX_x>>=
-plot.spline(beta.x)
-@
-%%%%%%%%%%%%%%%
-%% Section 3 %%
-%%%%%%%%%%%%%%%
-\section{Illness-death Model with Partial Markov Influences}
-Model build up in a general way:
-<<>>=
-traIDM <- matrix(nrow = 3, ncol = 3, FALSE)
-traIDM[1, 2] <- traIDM[1, 3] <- traIDM[2, 1] <- traIDM[2, 3] <- TRUE
-mpl <- mplskeleton(tmat = traIDM)
-mpl[[1]]$bhr[[2]] <- mpl[[1]]$bhr[[3]] <- mpl[[2]]$bhr[[1]] <- 
-                     mpl[[2]]$bhr[[3]] <- function(t){0.25}
-mpl[[1]]$eta[[2]] <- mpl[[1]]$eta[[3]] <- mpl[[2]]$eta[[1]] <- 
-  mpl[[2]]$eta[[3]] <- function(x.i, t){
-    ifelse(t < 5, 
-      return(0.5 * x.i[1]), 
-      return(0.5 * x.i[2]))}
-set.seed(123)
-N <- 500
-X <- matrix(nrow = N, ncol = 1, rnorm(N))
-X <- cbind(X, X[, 1] + rnorm(N)/10)
-colnames(X) <- c("x1.t1", "x1.t2")
-Xstruc <- matrix(ncol = 1, nrow = 2, data = 0)
-rownames(Xstruc) <- c("t1", "t2")
-colnames(Xstruc) <- c("x1")
-Xstruc[, 1] <- c(1, 2)
-Xstruc
-@
-Now set-up the partial Markov influences. 
-Function \texttt{partial.markov.x} has to take 5 input arguments representig vectors of past history information. 
-They have to take names \texttt{entry}, \texttt{exit}, \texttt{from}, \texttt{to}, and \texttt{delta}:
-<<>>=
-partial.markov.x <- function(entry, exit, from, to, delta){
-  count.12 <- sum(as.numeric((from == 1) & (to == 2) & (delta == 1)))
-  count.21 <- sum(as.numeric((from == 2) & (to == 1) & (delta == 1)))
-  return(c(count.12, count.21))}
-@
-The object \texttt{partial.markov.eta} is a list of lists in analogy to \texttt{mpl}:
-<<>>=
-partial.markov.eta <- pmeskeleton(traIDM)
-str(partial.markov.eta)
-partial.markov.eta[[1]][[2]] <- function(x){return( 0.25 * x[1])}
-partial.markov.eta[[1]][[3]] <- function(x){return( 0.5 * x[1])}
-partial.markov.eta[[2]][[1]] <- function(x){return(-0.5 * x[1] + 0.25 * x[2])}
-partial.markov.eta[[2]][[3]] <- function(x){return(0)}
-@
-Event history simulation:
-<<>>=
-d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
-                       change.times = c(5), Xstruc = Xstruc, 
-                       partial.markov.x = partial.markov.x, 
-                       partial.markov.eta = partial.markov.eta)
-@
-Nelson-Aalen estimator of the cumulative hazard rate functions:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
-plotnae(d$msm.basics, mpl = mpl)
-abline(a = 0, b = 0.25, col = "darkgrey")
-@
-
-Stratified partial likelihood modelling:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
-ld <- tolongformat(d = d, mpl = mpl)
-head(ld$msm.basics, n = 10)
-head(ld$ttsce[, 1:6], n = 10)
-m <- coxph(Surv(entry, exit, delta) ~ x1 + 
-             pm.x1.12 + pm.x1.13 + pm.x1.21 + pm.x2.21 + 
-             strata(trans), 
-           data = data.frame(ld$msm.basics, ld$ttsce))
-plotcph(m)
-points(c(0.5, 0.25, 0.5, -0.5, 0.25), pch = 16, col = "darkgrey")
-@
-
-Breslow estimate for cumulative hazard:
-<<fig=TRUE, pdf=TRUE, height=5, width=7, label=be>>=
-plotbe(m, mpl = mpl)
-abline(a = 0, b = 0.25, col = "darkgrey")
-@
 \end{document}

### 2016-05-20 JSS/jss-style/replicationScript.tex
replication script

--- 2016-03-09 JSS/jss-style/replicationScript.Rnw
@@
 \parindent 0pt
-\title{Replication Script to JSS Submission:\\
-\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
-\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
+\title{Replication Script to JSS Submission}
+\usepackage{Sweave}
+\begin{document}
+
+\author{Lindsay Rutter, Susan VanderPlas, Dianne Cook, Michelle A. Graham}
 \date{\today}
-\begin{document}
 \maketitle
-\SweaveOpts{concordance=TRUE}
+\input{replicationScript-concordance}
 \tableofcontents
-<<>>=
-install.packages("ggenealogy")
-library("ggenealogy")
-@
+
 %%%%%%%%%%%%%%%
-%% Section 1 %%
+%% Section 3 %%
 %%%%%%%%%%%%%%%
-\section{Generating a Graphical Object}
+\section*{3. Example datasets}
 
-Examine structure of sbGeneal example dataset:
+First, we can load and examine the structure of the soybean genealogy example dataset (called sbGeneal):
 
-<<>>=
-data(sbGeneal)
-str(sbGeneal)
-@
+\begin{Schunk}
+\begin{Sinput}
+> rm(list=)
+> library("ggenealogy")
+> data("sbGeneal")
+> str(sbGeneal)
+\end{Sinput}
+\begin{Soutput}
+'data.frame':	390 obs. of  5 variables:
+ $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
+ $ year        : num  1981 1948 1910 1912 1968 ...
+ $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
+ $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
+ $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
+\end{Soutput}
+\end{Schunk}
+
+After that, we can load and examine the structure of the academic statistician genealogy example dataset (called statGeneal):
+
+\begin{Schunk}
+\begin{Sinput}
+> data("statGeneal")
+> dim(statGeneal)
+\end{Sinput}
+\begin{Soutput}
+[1] 8165    6
+\end{Soutput}
+\begin{Sinput}
+> colnames(statGeneal)
+\end{Sinput}
+\begin{Soutput}
+[1] "child"   "parent"  "year"    "country" "school"  "thesis" 
+\end{Soutput}
+\end{Schunk}
 

### 2016-05-20 JSS/jss-style/ggenealogy/R/allFunctions.R

@@ -1,0 +1,1215 @@
+#' Returns the coordinate positions of all ancestors and descendants of a variety.
+#' 
+#' Calculates coordinates to plot each ancestors and descendant of a variety in a lineage. The x and y values describe the coordinates of the label, while the xstart, ystart, xend, and yend values describe the edges of the label.
+#' 
+#' @param df the data frame of the ancestors and descendants of a variety (from function buildAncDesTotalDF)
+#' @seealso \code{\link{buildAncList}} for information on determining ancestors
+#' @seealso \code{\link{buildDesList}} for information on determining descendants
+buildAncDesCoordDF = function(df){
+  id.offset <- NULL
+  root.gen <- gen <- par.id <- label <- NULL 
+  # This gets rid of redundancy and creates a "center"
+  if(nrow(subset(df, root.gen==0 & gen==0))>1){
+    temp = subset(df, root.gen==0 & gen==0)
+    temp$type = "center"
+    old.ids = temp$id[-1]
+    df$par.id[which(df$par.id%in%old.ids)] = temp$id[1]
+    temp$id[-1] = temp$id[1]
+    temp = unique(temp)
+    df = plyr::rbind.fill(subset(df, !(root.gen==0 & gen==0)), temp)
+  }
+  # This adds a leaf boolean column
+  df$leaf = !df$id%in%df$par.id
+  # Initialize x, y coords, and genside
+  df$genside = df$gen*(df$type=="descendant")-df$gen*(df$type=="ancestor")
+  df$x = 0
+  df$y = 0
+  # Ensure y coords are spread among all branches
+  df$y[df$leaf & df$genside<0] = seq(1, sum(df$leaf), length.out=sum(df$leaf & df$genside<0))
+  # but later, 2 ancestors are taken off b/c too previous
+  df$y[df$leaf & df$genside>0] = seq(1, sum(df$leaf), length.out=sum(df$leaf & df$genside>0))
+  # Ensure a single branch on one side is centrally placed (what if only 2 or 3?)
+  if(sum(df$leaf & df$genside<0)==1) df$y[df$leaf & df$genside<0] = sum(df$leaf)/2
+  if(sum(df$leaf & df$genside>0)==1) df$y[df$leaf & df$genside>0] = sum(df$leaf)/2
+  
+  df$xstart = 0
+  df$ystart = 0
+  df$xend = 0
+  df$yend = 0
+  df$branchx = 0
+  df$branchy = 0
[…]

### 2016-05-21 JSS/jss-style/article.tex
first draft rep script

@@
 
+\clearpage
+
 \begin{figure}[H]
@@
    ggplot2::element_text(size = 10), axis.title =
-   ggplot2::element_text(size = 10))
+   ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
+   = c(.1, .2))
 \end{CodeInput}
@@
 
+\clearpage
+
 \bibliography{article}

### 2016-05-21 JSS/jss-style/replicationScript.Rnw
first draft rep script

@@
 %%%%%%%%%%%%%%%
+%% Section 2 %%
+%%%%%%%%%%%%%%%
+\section*{2. Available software}
+
+The two figures in this section of the paper (Figure 1 and 2) were not produced using source code from our ggenealogy package.
+Instead, they were produced from external software for demonstration purposes.
+
+%%%%%%%%%%%%%%%
 %% Section 3 %%
@@
 
-First, we can load and examine the structure of the soybean genealogy example dataset (called sbGeneal):
+First, we can load and examine the structure of the example soybean genealogy dataset (called sbGeneal):
 
@@
 
-After that, we can load and examine the structure of the academic statistician genealogy example dataset (called statGeneal):
+After that, we can load and examine the structure of the example academic statistician genealogy dataset (called statGeneal):
 
@@
 
-Convert sbGeneal into igraph object:
+%%%%%%%%%%%%%%%
+%% Section 5 %%
+%%%%%%%%%%%%%%%
+\section*{5. Generating a graphical object}
+
+We can convert our example soybean genealogy dataset (called sbGeneal) into an igraph object:
+
+<<>>=
+sbIG <- dfToIG(sbGeneal)
+sbIG
+@
+
+Now that we have an igraph object, we can obtain basic summary statistics about it:
+
+<<>>=
+getBasicStatistics(sbIG)
+@
+
+%%%%%%%%%%%%%%%
+%% Section 6 %%
+%%%%%%%%%%%%%%%
+\section*{6. Plotting a shortest path}
+
+At this point, we can determine the shortest path of parent-child relationships between two labels of interest ("Tokyo" and "Narow").
+
+<<>>=
+pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
+pathTN
+@
+
+The returned path can then be plotted.
+
+<<>>=
+plotPath(pathTN)
+@
+
+We can do the same for a different pair of two labels of interest ("Bedford" and "Zane"). First, we can determine the years these two labels were identified.
+
+<<>>=
+getYear("Bedford", sbGeneal)
+getYear("Zane", sbGeneal)
+@
+
+Next, we can determine the shortest path of parent-child relationships between these two labels of interest and plot it.
+
+<<>>=
+pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
+plotPath(pathBZ, fontFace = 2)
+@
+
+%%%%%%%%%%%%%%%
+%% Section 7 %%
+%%%%%%%%%%%%%%%
+\section*{7. Superimposing shortest path on tree}
+
+In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN.
+Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
+
+<<>>=
+plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+@
+
+We can repeat this process, only now instead of setting the binVector variable to 1:3 (as we did earlier), we can set it to 1:6.
+
+<<>>=
+plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
+@
+
+%%%%%%%%%%%%%%%
+%% Section 8 %%
+%%%%%%%%%%%%%%%
+\section*{8. Plotting ancestors and descendants by generation}
+
+As is explained in the article, only the top part of Figure 6 (the figure from this section) is produced by ggenealogy code.
+In contrast, the bottom part of Figure 6 was produced by tools outside of ggenealogy for didactic purposes.
+Below, we recreate the top part of Figure 6, which was to generate a plot of the ancestors and descendants of the label Lee.
+
+<<>>=
+plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
+@
+
+%%%%%%%%%%%%%%%
+%% Section 9 %%
+%%%%%%%%%%%%%%%
+\section*{9. Plotting distance matrix}
+
+We can plot the distance matrix for a set of 10 varieties.
+
+<<>>=
+varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow", "Pella", "Tokyo", "Young", "Zane")
+plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") + ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") + ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 15))
+@
+
+%%%%%%%%%%%%%%%%
+%% Section 10 %%
+%%%%%%%%%%%%%%%%
+\section*{10. Academic genealogy of statisticians}
+
+We can now explore some of the plotting functions in ggenealogy, only now with the academic statistican genealogy dataset.
+This second example dataset is much larger than the first example dataset of soybean genealogy.
+For example purposes, we would like to view the ancestor and descendant plot for the individual who has the largest number of descendants.
+To identify the name of this individual, we run the following code:
+
+<<>>=
+library("dplyr")
+indVec <- getNodes(statGeneal)
+indVec <- indVec[which(indVec != "", )]
+dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
+numDesc <- sapply(indVec, dFunc)
+table(numDesc)
+which(numDesc == 159)
+@
+
+We see the individual with the largest number of descendants is Sir David Cox, who has 159 descendants. Now, we can plot the "ancestors" and "descendants" of Sir David Cox.
+
+<<>>=
+plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
+@
+
+It seems that of the 42 "children" of Sir David Cox, the one who went on to have the largest number of "children" of his own was Peter Bloomfield.
+We can verify below that Peter Bloomfield had 26 "children" and 49 "descendants".
+
+<<>>=
+length(getChild("Peter Bloomfield", statGeneal))
+nrow(getDescendants("Peter Bloomfield", statGeneal, gen = 100))
+@
+
+It would be of interest now to examine the shortest path between Sir David Cox and one of his newest "descendants" Petra Buzkova.
+To do so, we first need to obtain the corresponding igraph object of the example academic statistican genealogy dataset.
+
+<<>>=
+statIG <- dfToIG(statGeneal)
+@
+
+After doing so, we can now determine the shortest path between Sir David Cox and Petra Buzkova, and plot it.
+
+<<>>=
+pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected = FALSE)
+plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
+@
+
+We can now superimpose this shortest path between Sir David Cox and Petra Buzkova across the entire genealogical structure.
+
+<<>>=
+plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
+@
+
+We notice, however, that we cannot read the text of the nodes on the path of interest.
+To solve this problem, we can create the same plot, only now specifying that any nodes that are not on our path of interest are deemphasized with smaller text.
+
+<<>>=
+plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
+@
+
+Even though we can now read the text labels on the path of interest, we lost the ability to read labels that are not on our path of interest.
+At this point, we can keep the best of both worlds: We can create a plot that demphasizes the nodes that are not on the path of interest by assigning them small text font size, but incoroporate an interactive function so that we can hover over these non-path nodes if we wish to obtain their label information.
+This is the plot we used and interacted with to create the video embedded in Figure 12.
+
+<<>>=
+plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate = TRUE)
+@
 

### 2016-05-21 JSS/jss-style/ggenealogy/R/allFunctions.R
remove excess movies/figures

@@ -1,1215 +1,0 @@
-#' Returns the coordinate positions of all ancestors and descendants of a variety.
-#' 
-#' Calculates coordinates to plot each ancestors and descendant of a variety in a lineage. The x and y values describe the coordinates of the label, while the xstart, ystart, xend, and yend values describe the edges of the label.
-#' 
-#' @param df the data frame of the ancestors and descendants of a variety (from function buildAncDesTotalDF)
-#' @seealso \code{\link{buildAncList}} for information on determining ancestors
-#' @seealso \code{\link{buildDesList}} for information on determining descendants
-buildAncDesCoordDF = function(df){
-  id.offset <- NULL
-  root.gen <- gen <- par.id <- label <- NULL 
-  # This gets rid of redundancy and creates a "center"
-  if(nrow(subset(df, root.gen==0 & gen==0))>1){
-    temp = subset(df, root.gen==0 & gen==0)
-    temp$type = "center"
-    old.ids = temp$id[-1]
-    df$par.id[which(df$par.id%in%old.ids)] = temp$id[1]
-    temp$id[-1] = temp$id[1]
-    temp = unique(temp)
-    df = plyr::rbind.fill(subset(df, !(root.gen==0 & gen==0)), temp)
-  }
-  # This adds a leaf boolean column
-  df$leaf = !df$id%in%df$par.id
-  # Initialize x, y coords, and genside
-  df$genside = df$gen*(df$type=="descendant")-df$gen*(df$type=="ancestor")
-  df$x = 0
-  df$y = 0
-  # Ensure y coords are spread among all branches
-  df$y[df$leaf & df$genside<0] = seq(1, sum(df$leaf), length.out=sum(df$leaf & df$genside<0))
-  # but later, 2 ancestors are taken off b/c too previous
-  df$y[df$leaf & df$genside>0] = seq(1, sum(df$leaf), length.out=sum(df$leaf & df$genside>0))
-  # Ensure a single branch on one side is centrally placed (what if only 2 or 3?)
-  if(sum(df$leaf & df$genside<0)==1) df$y[df$leaf & df$genside<0] = sum(df$leaf)/2
-  if(sum(df$leaf & df$genside>0)==1) df$y[df$leaf & df$genside>0] = sum(df$leaf)/2
-  
-  df$xstart = 0
-  df$ystart = 0
-  df$xend = 0
-  df$yend = 0
-  df$branchx = 0
-  df$branchy = 0
[…]

### 2016-05-21 JSS/jss-style/ggenealogy/R/data-sbGeneal.R

@@ -1,25 +1,0 @@
-#' Soybean genealogical data
-#' 
-#' This data contains information on copy number variants, single nucleotide polymorphisms, protein content, and yield, of soybeans. The available data consists of a data frame structure that contains 412 direct child-parent relationships between pairs of soybean varieties. These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-
-#' @name sbGeneal
-#' @title Soybean data
-#' @description This data set contains soy bean genealogical information maintained by the United States Department of Agriculture to be used by plant breeders, geneticists, bioinformaticians, pathologists,and many other research workers.
-#' @docType data
-#' @format a \code{RData} instance, 1 row per each child-parent relationship between soybean varieties
-#' @details \itemize{
-  #' \item child name of child soybean variety
-  #' \item year child variety was introduced
-  #' \item yield protein yield
-  #' \item year.imputed whether or not the introduced year of the child variety was imputed
-  #' \item parent name of parent soybean variety
-  #' }
-  #'
-#' @docType data
-#' @keywords datasets
-#' @name sbGeneal
-#' @usage data(sbGeneal)
-#' @format A data frame with 412 rows and 7 variables
-#' @references
-#' "Pedigrees of soybean cultivars released in the United States and Canada." Theodore Hyivitz, C.A. Newell, S.G. Carmer. College of Agriculture, University of Illinois at Urbana-Champaign (1977).
-NULL

### 2016-05-21 JSS/jss-style/ggenealogy/R/data-statGeneal.R

@@ -1,26 +1,0 @@
-#' Academic statistics genealogical data
-#' 
-#' This data contains information on copy number variants, single nucleotide polymorphisms, protein content, and yield, of soybeans. The available data consists of a data frame structure that contains 412 direct child-parent relationships between pairs of soybean varieties. These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-
-#' @name statGeneal
-#' @title Academic statistics data
-#' @description This data set contains academic genealogical information from the Mathematics Genealogy Project, a web-based database serviced by the North Dakota State University Department of Mathematics and the American Mathematical Society. Specifically, this data set represents a subset of the Mathematical Genealogy Project that contains all the parent-child relationships where both parent and child received an advanced degree of statistics as of June 6, 2015.
-#' @docType data
-#' @format a \code{RData} instance, 1 row per each individual in the Mathematics Genealogy Project with an advanced degree in statistics. If the child had a parent who was also listed as having received an advanced degree in statistics in the Mathematics Genealogy project, then the row also contains the parental information.
-#' @details \itemize{
-  #' \item child name of the individual who received an advanced degree in statistics
-  #' \item parent name of the individual who mentored the child and also received an advanced degree in statistics. If the child has no such parent, then this field is an empty string
-  #' \item year year the child received their advanced degree in statistics
-  #' \item country country from which the child received their advanced degree in statistics
-  #' \item school school from which the child received their advanced degree in statistics
-  #' \item thesis title of the thesis the child submitted to receive their advanced degree in statistics. If this information is not available, then this field is an empty string
-#' }
-#'
-#' @docType data
-#' @keywords datasets
-#' @name statGeneal
-#' @usage data(statGeneal)
-#' @format A data frame with 8165 rows and 6 variables
-#' @references
-#' North Dakota State University and American Mathematical Society (2010). The Mathematics Genealogy Project. Archived Web Site. Retrieved from the Library of Congress, Accessed on March 6, 2015, URL http://www.genealogy.math.ndsu.nodak.edu.
-NULL

### 2016-05-21 JSS/submission/replicationScript.Rnw
updated statGeneal.rda

--- 2016-05-21 JSS/jss-style/replicationScript.Rnw
@@
 \documentclass{article}
-\parindent 0pt
+\usepackage{float, hyperref}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{caption}
+\usepackage{pgffor}
+
 \title{Replication Script to JSS Submission}
 \begin{document}
+
+<<options, echo=FALSE>>=
+library(knitr)
+opts_chunk$set(concordance=TRUE)
+@
 
@@
 \SweaveOpts{concordance=TRUE}
-\tableofcontents
 
@@
 
-<<>>=
+<<eval=TRUE>>=
 rm(list=)
+library("devtools")
+load_all("ggenealogy")
 library("ggenealogy")
@@
 
-<<>>=
+<<eval=TRUE>>=
 data("statGeneal")
@@
 
-<<>>=
+<<eval=TRUE>>=
 sbIG <- dfToIG(sbGeneal)
@@
 
-<<>>=
+<<eval=TRUE>>=
 getBasicStatistics(sbIG)
@@
 
-<<>>=
+<<eval=TRUE>>=
 pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotPath(pathTN)
@@
 
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
+plotPath(pathTN)
+@
+\end{figure}
+
 We can do the same for a different pair of two labels of interest ("Bedford" and "Zane"). First, we can determine the years these two labels were identified.
 
-<<>>=
+<<eval=TRUE>>=
 getYear("Bedford", sbGeneal)
@@
 
-<<>>=
+<<eval=TRUE>>=
 pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
@@
 
-<<>>=
+<<eval=TRUE>>=
 varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow", "Pella", "Tokyo", "Young", "Zane")
@@
 
-<<>>=
+<<eval=TRUE>>=
 library("dplyr")
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
@@
 
-<<>>=
+<<eval=TRUE>>=
 length(getChild("Peter Bloomfield", statGeneal))
@@
 
-<<>>=
+<<eval=TRUE>>=
 statIG <- dfToIG(statGeneal)
@@
 
-<<>>=
+<<eval=TRUE>>=
 pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected = FALSE)
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
@@
 
-<<>>=
+<<eval=TRUE>>=
 plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate = TRUE)

### 2016-05-21 JSS/submission/replicationScript.tex
updated statGeneal.rda

--- 2016-05-21 JSS/jss-style/replicationScript.Rnw
@@
 \documentclass{article}
-\parindent 0pt
+\usepackage{float, hyperref}
+\usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{caption}
+\usepackage{pgffor}
+
 \title{Replication Script to JSS Submission}
+\usepackage{Sweave}
 \begin{document}
@@
 \maketitle
-\SweaveOpts{concordance=TRUE}
-\tableofcontents
+\input{replicationScript-concordance}
 
@@
 
-<<>>=
-rm(list=)
-library("ggenealogy")
-data("sbGeneal")
-str(sbGeneal)
-@
+\begin{Schunk}
+\begin{Sinput}
+> rm(list=)
+> library("devtools")
+> load_all("ggenealogy")
+> library("ggenealogy")
+> data("sbGeneal")
+> str(sbGeneal)
+\end{Sinput}
+\begin{Soutput}
+'data.frame':	390 obs. of  5 variables:
+ $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
+ $ year        : num  1981 1948 1910 1912 1968 ...
+ $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
+ $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
+ $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-data("statGeneal")
-dim(statGeneal)
-colnames(statGeneal)
-@
+\begin{Schunk}
+\begin{Sinput}
+> data("statGeneal")
+> dim(statGeneal)
+\end{Sinput}
+\begin{Soutput}
+[1] 8165    6
+\end{Soutput}
+\begin{Sinput}
+> colnames(statGeneal)
+\end{Sinput}
+\begin{Soutput}
+[1] "child"   "parent"  "year"    "country" "school"  "thesis" 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-sbIG <- dfToIG(sbGeneal)
-sbIG
-@
+\begin{Schunk}
+\begin{Sinput}
+> sbIG <- dfToIG(sbGeneal)
+> sbIG
+\end{Sinput}
+\begin{Soutput}
+IGRAPH UNW- 230 340 -- 
++ attr: name (v/c), weight (e/n)
++ edges (vertex names):
+ [1] 5601T    --Hutcheson        Adams    --Dunfield        
+ [3] A.K.     --A.K. (Harrow)    Altona   --Flambeau        
+ [5] Amcor    --Amsoy 71         Adams    --Amsoy           
+ [7] Amsoy 71 --C1253            Anderson --Lincoln         
+ [9] Bay      --York             Bedford  --Forrest         
+[11] Beeson   --Kent             Blackhawk--Richland        
+[13] Bonus    --C1266R           Bradley  --J74-39          
+[15] Bragg    --Jackson          Bragg    --Bragg x D60-7965
++ ... omitted several edges
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-getBasicStatistics(sbIG)
-@
+\begin{Schunk}
+\begin{Sinput}
+> getBasicStatistics(sbIG)
+\end{Sinput}
+\begin{Soutput}
+$isConnected
+[1] FALSE
+
+$numComponents
+[1] 11
+
+$avePathLength
+[1] 5.333746
+
+$graphDiameter
+[1] 13
+
+$numNodes
+[1] 230
+
+$numEdges
+[1] 340
+
+$logN
+[1] 5.438079
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
-pathTN
-@
+\begin{Schunk}
+\begin{Sinput}
+> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
+> pathTN
+\end{Sinput}
+\begin{Soutput}
+$pathVertices
+[1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"
+
+$yearVertices
+[1] "1907"   "1942"   "1954.5" "1971.5" "1985"  
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotPath(pathTN)
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotPath(pathTN)
+\end{Sinput}
+\end{Schunk}
+
+\begin{figure}[H]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+> plotPath(pathTN)
+\end{Sinput}
+\end{Schunk}
+\includegraphics{replicationScript-008}
+\end{figure}
 
@@
 
-<<>>=
-getYear("Bedford", sbGeneal)
-getYear("Zane", sbGeneal)
-@
+\begin{Schunk}
+\begin{Sinput}
+> getYear("Bedford", sbGeneal)
+\end{Sinput}
+\begin{Soutput}
+[1] 1978
+\end{Soutput}
+\begin{Sinput}
+> getYear("Zane", sbGeneal)
+\end{Sinput}
+\begin{Soutput}
+[1] 1985
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
-plotPath(pathBZ, fontFace = 2)
-@
+\begin{Schunk}
+\begin{Sinput}
+> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
+> plotPath(pathBZ, fontFace = 2)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow", "Pella", "Tokyo", "Young", "Zane")
-plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") + ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") + ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 15))
-@
+\begin{Schunk}
+\begin{Sinput}
+> varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow", "Pella", "Tokyo", "Young", "Zane")
+> plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") + ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") + ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 15))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-library("dplyr")
-indVec <- getNodes(statGeneal)
-indVec <- indVec[which(indVec != "", )]
-dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
-numDesc <- sapply(indVec, dFunc)
-table(numDesc)
-which(numDesc == 159)
-@
+\begin{Schunk}
+\begin{Sinput}
+> library("dplyr")
+> indVec <- getNodes(statGeneal)
+> indVec <- indVec[which(indVec != "", )]
+> dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
+> numDesc <- sapply(indVec, dFunc)
+> table(numDesc)
+\end{Sinput}
+\begin{Soutput}
+numDesc
+   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14   15 
+6251  322  145   88   58   36   31   22   23   14   17   13   14   10    9    6 
+  16   17   18   19   20   21   22   23   24   25   26   27   29   30   34   37 
+   4    3    2    5    7    4    3    3    2    2    6    1    1    3    2    1 
+  38   40   41   44   45   48   49   60   61   75   77   84  159 
+   1    1    1    1    1    1    2    1    1    1    1    1    1 
+\end{Soutput}
+\begin{Sinput}
+> which(numDesc == 159)
+\end{Sinput}
+\begin{Soutput}
+David Cox 
+     1980 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-length(getChild("Peter Bloomfield", statGeneal))
-nrow(getDescendants("Peter Bloomfield", statGeneal, gen = 100))
-@
+\begin{Schunk}
+\begin{Sinput}
+> length(getChild("Peter Bloomfield", statGeneal))
+\end{Sinput}
+\begin{Soutput}
+[1] 26
+\end{Soutput}
+\begin{Sinput}
+> nrow(getDescendants("Peter Bloomfield", statGeneal, gen = 100))
+\end{Sinput}
+\begin{Soutput}
+[1] 49
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-statIG <- dfToIG(statGeneal)
-@
+\begin{Schunk}
+\begin{Sinput}
+> statIG <- dfToIG(statGeneal)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected = FALSE)
-plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
-@
+\begin{Schunk}
+\begin{Sinput}
+> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected = FALSE)
+> plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate = TRUE)
-@
+\begin{Schunk}
+\begin{Sinput}
+> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate = TRUE)
+\end{Sinput}
+\end{Schunk}
 

### 2016-05-22 JSS-ggenealogy/ReplicationScript.Rnw
Retain necessary files

--- 2016-05-21 JSS/submission/replicationScript.Rnw
@@
 library("devtools")
-load_all("ggenealogy")
+load_all("SourceCode")
 library("ggenealogy")
@@
 
-<<eval=TRUE>>=
-plotPath(pathTN)
-@
-
 \begin{figure}[H]
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
@@
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow", "Pella", "Tokyo", "Young", "Zane")
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotAncDes("David Cox", statGeneal, mAnc = 6, mDes = 6, vCol = "blue")
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected = FALSE)
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title = ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
 @
+\end{figure}
 
@@
 
-<<eval=TRUE>>=
+\begin{figure}[H]
+\centering
+<<fig=TRUE, echo=TRUE>>=
 plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate = TRUE)
 @
+\end{figure}
 

### 2016-07-13 article.tex
editorial changes

--- 2016-05-21 JSS/jss-style/article.tex
@@
 \end{CodeChunk}
-
 \subsection{Academic genealogy of statisticians}
-
 The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns.
@@
 With the final data frame of 8165 entries, we only maintained six of the original 26 variables.
-
 Each row of the final data frame contains information about a particular child who received a graduate-level academic degree in statistics, including the name of the child, the year the child obtained the degree, the country and school from which the child obtained the degree, the thesis title of the degree awarded to the child, and the name of its parent.
@@
 There are 3291 parent-child relationships in the dataset.
-
 We can load the example dataset of academic genealogy of statisticians (\code{statGeneal}) and examine its structure. 
-
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
 \section{Genealogical input format}
-
 As is the case with both example data files introduced above, \pkg{ggenealogy} requires that the genealogy input file is a data frame structure with at least two columns.
@@
 At this point, any \pkg{ggenealogy} plot that only requires information about parent-child relationships can be used.
-
 However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy.
@@
 At this point, any \pkg{ggenealogy} plot can be used.
-
 \section{Generating a graphical object}
-
 Most functions in the \pkg{ggenealogy} software package require an input parameter of a graph structure.
@@
 Below, we read in the \proglang{R} data file \code{sbGeneal} that is included in the package as a sample data set of soybean genealogy.
-
 We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function \code{dfToIG()}.
-
 \begin{CodeChunk}
@@
 \end{CodeChunk}
-
 There are many statistics about the \code{sbGeneal} dataset that we may wish to know that cannot easily be obtained through images and tables.
@@
 In addition to these statistics, the \code{getBasicStatistics()} function will also return the number of nodes, the number of edges, the average path length, the graph diameter, and other graph theoretical information.
-
 \begin{CodeChunk}
@@
 [1] 11
-
 $avePathLength
@@
 [1] 13
-
 $numNodes
@@
 [1] 340
-
 $logN
@@
 [1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"   
-
 $yearVertices
@@
     \centering
-    \fbox{\includegraphics[width=\textwidth]{LeeAD3}}
+    \fbox{\includegraphics[width=\textwidth]{LeeAD}}
     \caption{Top: All ancestors and descendants of the variety Lee are shown in this ggenealogy plot.
@@
    0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
-6252  322  145   88   58   36   31   22   23   14   17   13   14    9    9 
+6251  322  145   88   58   36   31   22   23   14   17   13   14   10    9 
   15   16   17   18   19   20   21   22   23   24   25   26   27   29   30 
-   6    4    3    2    5    7    5    3    3    2    2    6    1    1    3 
-  34   37   38   40   41   44   45   48   49   61   62   75   77   84  159 
+   6    4    3    2    5    7    4    3    3    2    2    6    1    1    3 
+  34   37   38   40   41   44   45   48   49   60   61   75   77   84  159 
    2    1    1    1    1    1    1    1    2    1    1    1    1    1    1 

### 2016-07-14 article.tex
abstract

@@
 The tools are being tested on a dataset with milestone cultivars of soybean varieties (\citealt{soybean}) as well as on a web-based database of the academic genealogy of mathematicians (\citealt{mgp}).
-The software package is currently available on the Comprehensive \proglang{R} Archive Network.}
+The latest stable package version is available in source and binary form on the Comprehensive \proglang{R} Archive Network.}
 
@@
 With the final data frame of 8165 entries, we only maintained six of the original 26 variables.
+
 Each row of the final data frame contains information about a particular child who received a graduate-level academic degree in statistics, including the name of the child, the year the child obtained the degree, the country and school from which the child obtained the degree, the thesis title of the degree awarded to the child, and the name of its parent.
@@
 There are 3291 parent-child relationships in the dataset.
+
 We can load the example dataset of academic genealogy of statisticians (\code{statGeneal}) and examine its structure.
+
 \begin{CodeChunk}
@@
 \end{CodeChunk}
+
 \section{Genealogical input format}
+
 As is the case with both example data files introduced above, \pkg{ggenealogy} requires that the genealogy input file is a data frame structure with at least two columns.
@@
 At this point, any \pkg{ggenealogy} plot that only requires information about parent-child relationships can be used.
+
 However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy.
@@
 At this point, any \pkg{ggenealogy} plot can be used.
+
 \section{Generating a graphical object}
+
 Most functions in the \pkg{ggenealogy} software package require an input parameter of a graph structure.
@@
 Below, we read in the \proglang{R} data file \code{sbGeneal} that is included in the package as a sample data set of soybean genealogy.
-We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function \code{dfToIG()}.
+
+We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function 
+\code{dfToIG()}.
 \begin{CodeChunk}
@@
 \end{CodeChunk}
+
 There are many statistics about the \code{sbGeneal} dataset that we may wish to know that cannot easily be obtained through images and tables.
@@
 In addition to these statistics, the \code{getBasicStatistics()} function will also return the number of nodes, the number of edges, the average path length, the graph diameter, and other graph theoretical information.
+
 \begin{CodeChunk}

### 2016-07-15 article.tex
editorial recommended change

@@
 
+\section{Package overview}
+
+We will now provide a brief overview of the functionality of the \pkg{ggenealogy} package before going into more detail with examples later in this paper.
+With the \pkg{ggenealogy} package, users can convert genealogical data into graph structures.
+This allows for users to efficiently traverse, analyze, and elicit graph theoretical measurements on genealogical lineages (see Section \ref{ggo}).
+These capabilities were developed by building upon the \pkg{igraph} software (\citealt{igraph}).
+
+Additionally, the \pkg{ggenealogy} package allows for users to plot genealogical data in multiple ways.
+Users can obtain and plot the shortest path between two nodes of interest constrained on a variable of interest (see Section \ref{psp}), superimpose a shortest path between two nodes of interest with the entire genealogical structure constrained on a variable of interest (see Section \ref{sspt}), plot the ancestors and descendants of a node of interest constrained on generation count (see Section \ref{remedy}), and plot distance matrices based on a variable of interest (see Section \ref{pdm}).
+Most of these plotting tools were developed by building upon the \pkg{ggplot2} package (\citealt{ggplot2}).
+As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in example code throughout this paper.
+Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Subsection \ref{interaction}).
+
 \section{Example datasets}
@@
 \section{Generating a graphical object}
+\label{ggo}
 
@@
 
-We now convert it into an \pkg{igraph} object (\citealt{igraph}) \code{sbIG} using the function 
+We now convert it into an \pkg{igraph} object \code{sbIG} using the function 
 \code{dfToIG()}.
@@
 \section{Plotting a shortest path}
+\label{psp}
 
@@
 \section{Superimposing shortest path on tree}
+\label{sspt}
 
@@
 
-In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions (\citealt{ggplot2}).
+In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions.
 While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation.
@@
 \section{Plotting distance matrix}
+\label{pdm}
 
@@
 
+\subsection{Plotting animation}
+\label{interaction}
+
 We could still improve upon Figure \ref{fig:plotCBNoText}.

### 2016-07-15 article.tex
reference fix

@@
 Users can obtain and plot the shortest path between two nodes of interest constrained on a variable of interest (see Section \ref{psp}), superimpose a shortest path between two nodes of interest with the entire genealogical structure constrained on a variable of interest (see Section \ref{sspt}), plot the ancestors and descendants of a node of interest constrained on generation count (see Section \ref{remedy}), and plot distance matrices based on a variable of interest (see Section \ref{pdm}).
+
 Most of these plotting tools were developed by building upon the \pkg{ggplot2} package (\citealt{ggplot2}).
-As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in example code throughout this paper.
+As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in demonstrative code throughout this paper.
 Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Subsection \ref{interaction}).

### 2016-07-15 article.tex
code input indents

@@
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red",
-  nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
-  element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
++    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
++    element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 \end{Code}
@@
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol =
-  "seagreen2", nodeSize = 1, pathNodeSize = 3)
++    "seagreen2", nodeSize = 1, pathNodeSize = 3)
 \end{Code}
@@
 >R varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
-  "Pella", "Tokyo", "Young", "Zane")
++    "Pella", "Tokyo", "Young", "Zane")
 >R plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") +
-  ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
-  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text
-  = ggplot2::element_text(size = 15))
++    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
++    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
++    legend.text = ggplot2::element_text(size = 15))
 \end{Code}
@@
 R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
-  isDirected = FALSE)
++    isDirected = FALSE)
 R> plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text =
-   ggplot2::element_text(size = 10), axis.title =
-   ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
-   = c(.1, .2))
++    ggplot2::element_text(size = 10), axis.title =
++    ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
++    = c(.1, .2))
 \end{CodeInput}
@@
 R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) +
-   ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
-   ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-   (expand = c(.1, .2))
++    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
++    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
++    (expand = c(.1, .2))
 \end{CodeInput}
@@
 R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
-   pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
-   ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
-   ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-   (expand = c(.1, .2))
++    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
++    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
++    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
++    (expand = c(.1, .2))
 \end{CodeInput}
@@
 R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
-   pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray", animate =
-   TRUE)
++    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
++    animate = TRUE)
 \end{CodeInput}

### 2016-07-16 article.tex
\pkg formatting

@@
     \fbox{\includegraphics[width=\textwidth]{LeeAD}}
-    \caption{Top: All ancestors and descendants of the variety Lee are shown in this ggenealogy plot.
-Bottom: We now attempt to mimic the blue paths in the ggenealogy plot on the top, only now nodes cannot be repeated.
+    \caption{Top: All ancestors and descendants of the variety Lee are shown in this \pkg{ggenealogy} plot.
+Bottom: We now attempt to mimic the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated.
 The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee.
-The presence of red parentheses indicate that the plot on the bottom ambiguously display the example soybean genealogy in the way that the ggenealogy plot on the top can accomplish.}
+The presence of red parentheses indicate that the plot on the bottom ambiguously display the example soybean genealogy in the way that the \pkg{ggenealogy} plot on the top can accomplish.}
     \label{fig:Lee}

### 2016-07-18 article.tex
kinship2 format consistency
title case and caps after colon

@@
     \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
-    \caption{Example pedigree chart from the kinship2 package, where the vertical axis denotes generation count.
+    \caption{Example pedigree chart from the \pkg{kinship2} package, where the vertical axis denotes generation count.
 We superimposed green-highlighted individual 215 for explanatory purposes.
@@
     \fbox{\includegraphics[width=\textwidth]{pathTNZB}}
-    \caption{Left: The shortest path between varieties Tokyo and Narow is strictly composed of a unidirectional sequence of parent-child relationships.
-Right: The shortest path between varieties Zane and Bedford is not strictly composed of unidirectional parent-child relationships; they instead have a cousin-like relationship.}
+    \caption{Left: The shortest path between varieties \code{Tokyo} and \code{Narow} is strictly composed of a unidirectional sequence of parent-child relationships.
+Right: The shortest path between varieties \code{Zane} and \code{Bedford} is not strictly composed of unidirectional parent-child relationships; they instead have a cousin-like relationship.}
     \label{fig:pathTNZB}
@@
     \fbox{\includegraphics[width=\textwidth]{plotTNBin3}}
-    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 3.}
+    \caption{The shortest path between \code{Tokyo} and \code{Narow}, superimposed over the data structure, using a bin size of 3.}
     \label{fig:plotTNBin3}
@@
     \fbox{\includegraphics[width=\textwidth]{plotTNBin6}}
-    \caption{The shortest path between Tokyo and Narow, superimposed over the data structure, using a bin size of 6.}
+    \caption{The shortest path between \code{Tokyo} and \code{Narow}, superimposed over the data structure, using a bin size of 6.}
     \label{fig:plotTNBin6}
@@
     \fbox{\includegraphics[width=\textwidth]{LeeAD}}
-    \caption{Top: All ancestors and descendants of the variety Lee are shown in this \pkg{ggenealogy} plot.
+    \caption{Top: All ancestors and descendants of the variety \code{Lee} are shown in this \pkg{ggenealogy} plot.
 Bottom: We now attempt to mimic the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated.
-The parenthetical numbers above each node represents the set of generation counts that node is away from the center node Lee.
+The parenthetical numbers above each node represents the set of generation counts that node is away from the center node \code{Lee}.
 The presence of red parentheses indicate that the plot on the bottom ambiguously display the example soybean genealogy in the way that the \pkg{ggenealogy} plot on the top can accomplish.}
@@
 There are only 17 individuals who have more than 30 ``descendants", and there is one individual who has the largest value of 159 ``descendants".
-We determine that this individual is the prominent British statistician Sir David Cox, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.
+We determine that this individual is the prominent British statistician Sir \code{David Cox}, who is known for the Box-Cox transformation and Cox processes, as well as for mentoring many younger researchers who later became notable statisticians themselves.
 
@@
 
-We can now visualize how these 159 ``descendants" are related to Sir David Cox by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}.
+We can now visualize how these 159 ``descendants" are related to Sir \code{David Cox} by calling the \code{plotAncDes()} function of \pkg{ggenealogy}, similar to what we did to generate Figure \ref{fig:Lee}.
 As such, we create Figure \ref{fig:dCox} using the code below.
@@
     \fbox{\includegraphics[width=\textwidth]{dCox.png}}
-    \caption{The 159 academic statistician ``descendants" of Sir David Cox.}
+    \caption{The 159 academic statistician ``descendants" of Sir \code{David Cox}.}
     \label{fig:dCox}
@@
 
-We see from Figure \ref{fig:dCox} that Sir David Cox had 42 ``children", many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
+We see from Figure \ref{fig:dCox} that Sir \code{David Cox} had 42 ``children", many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
 Of his ``children", the one who produced the most ``children" of their own was Peter Bloomfield, who has 26 ``children" and 49 ``descendants".
-In total, Sir David Cox had five generations of academic statistics mentees in this dataset.
+In total, Sir \code{David Cox} had five generations of academic statistics mentees in this dataset.
 
@@
 
-At this point, it would be insightful to examine a more detailed view of one of the longest strings of ``parent-child" relationships between Sir David Cox and one of the two individuals who are his fifth generation ``descendants".
-We do so with the code below, choosing his fifth generation ``descendant" to be Petra Buzkova.
+At this point, it would be insightful to examine a more detailed view of one of the longest strings of ``parent-child" relationships between Sir \code{David Cox} and one of the two individuals who are his fifth generation ``descendants".
+We do so with the code below, choosing his fifth generation ``descendant" to be \code{Petra Buzkova}.
 We set the \code{fontFace} variable of the \code{plotPath()} to a value of 4, indicating we wish to boldface and italicize the two varieties of interest.
@@
     \fbox{\includegraphics[width=0.7\textwidth]{pathCB}}
-    \caption{The shortest path between Sir David Cox and one of his fifth generation ``descendants", Petra Buzkova.}
+    \caption{The shortest path between Sir \code{David Cox} and one of his fifth generation ``descendants", \code{Petra Buzkova}.}
     \label{fig:pathCB}
@@
 This code results in Figure \ref{fig:pathCB}.
-We see that the shortest path between Sir David Cox and Petra Buzkova is strictly composed of five unidirectional ``parent-child" relationships that span about 55 years.
+We see that the shortest path between Sir \code{David Cox} and \code{Petra Buzkova} is strictly composed of five unidirectional ``parent-child" relationships that span about 55 years.
 We see that the time difference between when an advisor and student earned their degrees is not consistent across this path: The three statisticians who earned their degrees earliest in this path span more than 30 years in degree acquisition, whereas the three statisticians who earned their degrees later in this path only span less than ten years in degree acquisition.
 
-We also notice in Figure \ref{fig:pathCB} that Sir David Cox received his statistics degree in about 1950, and Petra Buzkova received her statistics degree in about 2005.
+We also notice in Figure \ref{fig:pathCB} that Sir \code{David Cox} received his statistics degree in about 1950, and \code{Petra Buzkova} received her statistics degree in about 2005.
 This genealogy only contains historical information about obtained degrees, and does not project into the future.
-Hence, we can be assured that Petra Buzkova is one of the younger individuals in the dataset, at least in the sense that the youngest individual could only have received his or her degree ten years after Petra Buzkova.
-However, we cannot be assured that Sir David Cox is one of the oldest individuals in the dataset.
+Hence, we can be assured that \code{Petra Buzkova} is one of the younger individuals in the dataset, at least in the sense that the youngest individual could only have received his or her degree ten years after \code{Petra Buzkova}.
+However, we cannot be assured that Sir \code{David Cox} is one of the oldest individuals in the dataset.
 As such, it would be informative to superimpose this path of interest onto the entire dataset, using the \code{plotPathOnAll()} function of the \pkg{ggenealogy} package, as we did for the soybean genealogy in Figures \ref{fig:plotTNBin3} and \ref{fig:plotTNBin6}.
@@
     \fbox{\includegraphics[width=\textwidth]{plotCBText}}
-    \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.}
+    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.}
     \label{fig:plotCBText}
@@
 We see from the resulting Figure \ref{fig:plotCBText} that almost all text labels for individuals who received their graduate-level statistics degrees between 1950 and 2015 are undecipherable.
-We also see that the year Sir David Cox acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for acquisition of statistics degrees in this dataset occur around 1860.
+We also see that the year Sir \code{David Cox} acquired his statistics degree is somewhere in the later half of the variable year for this dataset, as the oldest dates for acquisition of statistics degrees in this dataset occur around 1860.
 However, the number of individuals who are documented as receiving their statistics degrees between 1860 and 1950 are few enough so that their text labels are somewhat readable.
@@
     \fbox{\includegraphics[width=\textwidth]{plotCBNoText}}
-    \caption{The shortest path between Sir David Cox and Petra Buzkova, superimposed over the data structure, using a bin size of 200.
+    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.
 Individuals on the shortest path are labeled in large and black text and connected by dark green edges; all other individuals are labeled in small and gray text and connected by light gray edges.}
@@
 In Figure \ref{fig:plotCBNoText}, we can now see each individual on the path of interest, and how their values for the variable year are overlaid on the entire genealogy structure.
-We can also more clearly see that, even though only ten years span between the youngest individual in the genealogy and Petra Buzkova, there are many individuals in that last decade.
+We can also more clearly see that, even though only ten years span between the youngest individual in the genealogy and \code{Petra Buzkova}, there are many individuals in that last decade.
 Indeed, the decade from 2005 to 2015 appears to be the densest in this dataset in terms of acquisition of statistics degrees.

### 2016-07-21 JSS-ggenealogy/ReplicationScript.R
replication script .R

@@ -1,0 +1,78 @@
+# Load and examine the structure of the example soybean genealogy dataset (called sbGeneal)
+rm(list=)
+library("devtools")
+load_all("SourceCode")
+library("ggenealogy")
+data("sbGeneal")
+str(sbGeneal)
+
+# Load and examine the structure of the example academic statistician genealogy dataset (called statGeneal)
+data("statGeneal")
+dim(statGeneal)
+colnames(statGeneal)
+
+# Convert our example soybean genealogy dataset (called sbGeneal) into an igraph object:
+sbIG <- dfToIG(sbGeneal)
+sbIG
+
+# Obtain basic summary statistics about the igraph object
+getBasicStatistics(sbIG)
+
+# Determine shortest path of parent-child relationships between two labels of interest ("Tokyo" and "Narow")
+pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
+pathTN
+
+# Plot returned path
+plotPath(pathTN)
+
+# Do the same for a different pair of two labels of interest ("Bedford" and "Zane"). First, we can determine the years these two labels were identified
+getYear("Bedford", sbGeneal)
+getYear("Zane", sbGeneal)
+
+# Next, we can determine the shortest path of parent-child relationships between these two labels of interest and plot it.
+pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
+plotPath(pathBZ, fontFace = 2)
+
+# In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
+plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+
+# We can repeat this process, only now instead of setting the binVector variable to 1:3 (as we did earlier), we can set it to 1:6.
+plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
+
[…]

### 2016-07-21 article.tex
explain not from ggenealogy plot

@@
     \caption{Top: All ancestors and descendants of the variety \code{Lee} are shown in this \pkg{ggenealogy} plot.
-Bottom: We now attempt to mimic the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated.
+Bottom: This plot was constructed manually outside of the \pkg{ggenealogy} plot for didactic purposes.
+It mimics the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated.
 The parenthetical numbers above each node represents the set of generation counts that node is away from the center node \code{Lee}.

### 2016-07-21 JSS-ggenealogy/ReplicationScript.R

@@ -1,5 +1,5 @@
 # Load and examine the structure of the example soybean genealogy dataset (called sbGeneal)
-rm(list=)
+rm(list=ls())
 library("devtools")
 load_all("SourceCode")
 library("ggenealogy")

### 2016-07-22 article.tex
\clearpage for layout

@@
 Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Subsection \ref{interaction}).
+
+\clearpage
 
@@
 
-\clearpage
-
 \begin{figure}[H]
@@
 We can immediately see that Figure \ref{fig:plotTNBin6} more successfully mitigates text overlap compared to Figure \ref{fig:plotTNBin3}.
-We can also confirm what we saw in the previous plot that indeed most varieties were released between 1950 and 1970, and any textual overlap is confined to this range of years.
+We also confirm what we saw in the previous plot that most varieties were released between 1950 and 1970, and any textual overlap is confined to this range of years.
 
@@
 The further from the center that a variety is located, the more generations that variety is distanced from the centered variety of interest.
-
-This particular \pkg{ggenealogy} tool is unique because most available genealogy and graph visualization software do not allow for repeated labels.
-It is a useful tool because, as was demonstrated in Figures \ref{fig:kinshipFig} and \ref{fig:Graph}, some genealogical datasets require repeated node labels if they are to be visualized by generation counts.
-Indeed, our example soybean genealogy is one such dataset.
+This particular \pkg{ggenealogy} tool is uniquely beneficial because most genealogy and graph visualization software do not allow for repeated node labels even though some genealogical datasets require repeated node labels in order to be visualized by generation counts (as was shown in Figures \ref{fig:kinshipFig} and \ref{fig:Graph}).
 
@@
 We notice that \code{Lee} has 3 generations of ancestors and 5 generations of descendants.
-We also notice that some varieties are repeated in the plot, which is a unique feature provided by \pkg{ggenealogy}.
+We also notice that some varieties are repeated, which is a unique feature provided by \pkg{ggenealogy}.
 For example, the variety \code{5601T} is represented four times - once as a third generation descendant of \code{Lee}, once as a fourth generation descendant of \code{Lee}, and twice as a fifth generation descendant of \code{Lee}.
@@
     \caption{Top: All ancestors and descendants of the variety \code{Lee} are shown in this \pkg{ggenealogy} plot.
-Bottom: This plot was constructed manually outside of the \pkg{ggenealogy} plot for didactic purposes.
+Bottom: For didactic purposes, this plot was constructed manually outside of the \pkg{ggenealogy} package.
 It mimics the blue paths in the \pkg{ggenealogy} plot on the top, only now nodes cannot be repeated.
@@
 
-\subsection{Plotting animation}
+\subsection{Interactive plotting functions}
 \label{interaction}

### 2016-07-22 JSS-ggenealogy/ReplicationScript.R

@@ -64,6 +64,7 @@ nrow(getDescendants("Peter Bloomfield", statGeneal, gen = 100))
 
 # It would be of interest now to examine the shortest path between Sir David Cox and one of his newest "descendants" Petra Buzkova. To do so, we first need to obtain the corresponding igraph object of the example academic statistican genealogy dataset.
 statIG <- dfToIG(statGeneal)
+pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected = FALSE)
 
 # After doing so, we can now determine the shortest path between Sir David Cox and Petra Buzkova, and plot it.
 plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))

### 2016-11-06 article.tex
Reviewer wording suggestions

@@
 However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy.
-For these plots, the input data frame should also contain a third column.
-In both example data files, this column is labeled ``year," and each case in that column can either be of type numeric, type NA, or type ``".
-At this point, any \pkg{ggenealogy} plot can be used.
+For these plots, the input data frame should also contain a third quantitative variable column.
+Each case in this quantitative variable column can be of type numeric, type NA, or type ``".
+In both example data files, this column is labeled ``year,".
+However, we would like to emphasize that the name of quantitative variable column to be plotted is user-specified.
+For instance, should the data of interest contain time-related values on a shorter scale than years (such as days or months), then the column could be named ``day" or ``month".
+This flexibilitiy allows for more diverse datasets to easily be examined with the package.
 
@@
 
-The \pkg{ggenealogy} package comes with two example datasets, and while we have introduced the plant breeding genealogy, we have yet to introduce the academic genealogy.
-As was demonstrated in Section \ref{exData}, every parent in the academic genealogy is also a child, and some children in the academic genealogy have more than two parents.
+An academic genealogy is the second dataset provided in the package, where every parent is also a child and some children have more than two parents, as was briefly described in Section \ref{exData}.
 Neither of these features was the case in the plant breeding genealogy.
@@
 We do so with the code below, choosing his fifth generation ``descendant" to be \code{Petra Buzkova}.
-We set the \code{fontFace} variable of the \code{plotPath()} to a value of 4, indicating we wish to boldface and italicize the two varieties of interest.
+We set the \code{fontFace} variable of the \code{plotPath()} to a value of 4, indicating we wish to boldface and italicize the two statisticians of interest.
 

### 2016-11-06 article.tex
binVector to bin

@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red",
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:3, pathEdgeCol = "red",
 +    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
@@
 In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions.
-While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{binVector}) requires some explanation.
+While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{bin}) requires some explanation.
 The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, year of release).
@@
 Hence, in an attempt to mitigate text overlapping, the \code{plotPathOnAll()} function does not randomly assign the y coordinate.
-Instead, it allows users to partially control the y coordinates with a user-determined number of bins (\code{binVector}).
+Instead, it allows users to partially control the y coordinates with a user-determined number of bins (\code{bin}).
 
@@
 
-In the code above, \code{binVector} was assigned a value of 3, and \code{pathEdgeCol} was assigned a value of ``red".
+In the code above, \code{bin} was assigned a value of 3, and \code{pathEdgeCol} was assigned a value of ``red".
 Additionally, we specified a size of 2.5 for the non-path node test using the \code{nodeSize} parameter, and a size of 4 for the path node text using the \code{pathNodeSize} parameter.
@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol =
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:6, pathEdgeCol =
 +    "seagreen2", nodeSize = 1, pathNodeSize = 3)
@@
 We can achieve this using the below code.
-After trial and error, we use a \code{binVector} of size 200, and append \pkg{ggplot2} syntax to define suitable x-axis limits.
+After trial and error, we use a \code{bin} of size 200, and append \pkg{ggplot2} syntax to define suitable x-axis limits.
 The output of this process is illustrated in Figure \ref{fig:plotCBText}.
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200) +
+R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200) +
 +    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
@@
 
-The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{binVector} will not solve the text overlapping problem.
+The text labels are so numerous in Figure \ref{fig:plotCBText} that simply trying different values for the input parameter \code{bin} will not solve the text overlapping problem.
 Instead, one approach we can try is to reconstruct the plot using the same \pkg{ggenealogy} function \code{plotPathOnAll()}, only now specifying variables to render the size (2.5) and color (default of black) of the text for nodes that are on the path of interest to be more noticeable than the size (0.5) and color (dark gray) of the text for nodes that are not on the path of interest.
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
+R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
 +    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, binVector = 1:200, nodeSize = .5,
+R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
 +    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",

### 2016-11-06 JSS-ggenealogy/ReplicationScript.R

@@ -34,10 +34,10 @@ pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
 plotPath(pathBZ, fontFace = 2)
 
 # In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
-plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 
-# We can repeat this process, only now instead of setting the binVector variable to 1:3 (as we did earlier), we can set it to 1:6.
-plotPathOnAll(pathTN, sbGeneal, sbIG, binVector = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
+# We can repeat this process, only now instead of setting the bin variable to 1:3 (as we did earlier), we can set it to 1:6.
+plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
 
 # As is explained in the article, only the top part of Figure 6 (the figure from this section) is produced by ggenealogy code. In contrast, the bottom part of Figure 6 was produced by tools outside of ggenealogy for didactic purposes. Below, we recreate the top part of Figure 6, which was to generate a plot of the ancestors and descendants of the label Lee.
 plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
@@ -70,10 +70,10 @@ pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, isDirected =
 plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
 
[…]

### 2016-11-06 article.tex
replication script update

@@
  $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
- $ year        : num  1981 1948 1910 1912 1968 ...
+ $ date        : num  1981 1948 1910 1912 1968 ...
  $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
- $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
+ $ date.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
  $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
@@
 \begin{CodeOutput}
-[1] "child"   "parent"  "year"    "country" "school"  "thesis"
+[1] "child"   "parent"  "date"    "country" "school"  "thesis" 
 \end{CodeOutput}
@@
 Each case in this quantitative variable column can be of type numeric, type NA, or type ``".
-In both example data files, this column is labeled ``year,".
-However, we would like to emphasize that the name of quantitative variable column to be plotted is user-specified.
-For instance, should the data of interest contain time-related values on a shorter scale than years (such as days or months), then the column could be named ``day" or ``month".
-This flexibilitiy allows for more diverse datasets to easily be examined with the package.
+In both example data files, this column is labeled ``date," and each case in that column can be of type numeric, type NA, or type ``".
+At this point, any \pkg{ggenealogy} plot can be used.
 
@@
 [1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"   
-$yearVertices
+$dateVertices
 [1] "1907"   "1942"   "1954.5" "1971.5" "1985"  
@@
 Next, we can run the same set of functions on a different pair of varieties.
-First, a call to the \pkg{ggenealogy} function \code{getYear()} indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.
+First, a call to the \pkg{ggenealogy} function \code{getDate()} indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.
 
@@
 \begin{CodeInput}
-R> getYear("Bedford", sbGeneal)
+R> getDate("Bedford", sbGeneal)
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> getYear("Zane", sbGeneal)
+R> getDate("Zane", sbGeneal)
 \end{CodeInput}
@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:3, pathEdgeCol = "red",
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 3, pathEdgeCol = "red",
 +    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:6, pathEdgeCol =
-+    "seagreen2", nodeSize = 1, pathNodeSize = 3)
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 6, pathEdgeCol =
++    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Year")
 \end{Code}
@@
 We also confirm what we saw in the previous plot that most varieties were released between 1950 and 1970, and any textual overlap is confined to this range of years.
+
+Notice from Figure~\ref{fig:plotTNBin3} that the default horizontal axis label for the \texttt{plotPath()} method has a value of ``Date".
+Given that the ``date" variable in this example dataset is on the timescale of years, we wanted to change the default value of the horizontal axis label to ``Year".
+We did this in the code above for Figure~\ref{fig:plotTNBin6} by appending the \texttt{ggplot2::xlab()} function.
 
@@
 +    isDirected = FALSE)
-R> plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text =
-+    ggplot2::element_text(size = 10), axis.title =
+R> plotPath(pathCB, fontFace = 4) + ggplot2::xlab("Year") +
++    ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title =
 +    ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200) +
+R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200)
 +    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
 +    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-+    (expand = c(.1, .2))
++    (expand = c(.1, .2)) + ggplot2::xlab("Year")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
+R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
 +    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
@@
 +    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-+    (expand = c(.1, .2))
++    (expand = c(.1, .2)) + ggplot2::xlab("Year")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
+R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
 +    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",

### 2016-11-06 JSS-ggenealogy/ReplicationScript.R

@@ -26,18 +26,18 @@ pathTN
 plotPath(pathTN)
 
 # Do the same for a different pair of two labels of interest ("Bedford" and "Zane"). First, we can determine the years these two labels were identified
-getYear("Bedford", sbGeneal)
-getYear("Zane", sbGeneal)
+getDate("Bedford", sbGeneal)
+getDate("Zane", sbGeneal)
 
 # Next, we can determine the shortest path of parent-child relationships between these two labels of interest and plot it.
 pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
 plotPath(pathBZ, fontFace = 2)
 
 # In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
-plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 
 # We can repeat this process, only now instead of setting the bin variable to 1:3 (as we did earlier), we can set it to 1:6.
-plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3)
+plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Year")
 
 # As is explained in the article, only the top part of Figure 6 (the figure from this section) is produced by ggenealogy code. In contrast, the bottom part of Figure 6 was produced by tools outside of ggenealogy for didactic purposes. Below, we recreate the top part of Figure 6, which was to generate a plot of the ancestors and descendants of the label Lee.
 plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
@@ -67,13 +67,14 @@ statIG <- dfToIG(statGeneal)
[…]

### 2016-11-06 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -153,7 +153,7 @@ buildAncDesTotalDF = function(v1, geneal, mAnc=3, mDes=3){
     })
     
     # This creates a unique color for the variety label of interest
-    cols = hcl(h=seq(0, 300, by=50), c=80, l=55, fixup=TRUE)
+    cols = grDevices::hcl(h=seq(0, 300, by=50), c=80, l=55, fixup=TRUE)
     temp2$color = "#000000"
     
     # The number 7 was selected as it is the average working memory maximum
@@ -254,20 +254,20 @@ buildDesList = function(v1, geneal, gen=0){
 #' 
 #' @param geneal the full genealogy  (in data frame format)
 #' @param ig the graph representation of the data genealogy (in igraph format)
-#' @param binVector the number of bins between 1 and length(binVector) (default is 12). For more information on choosing binVector size, please visit the ggenealogy vignette.
+#' @param bin the number of bins to determine the vertical positions of nodes (default is 12). For more information on choosing bin size, please visit the ggenealogy vignette.
 #' @seealso \code{\link{dfToIG}} for information on producing ig from the genealogy
 #' @seealso \url{http://www.r-project.org} for iGraph information
-buildEdgeTotalDF = function(geneal, ig, binVector = 1:12){
+buildEdgeTotalDF = function(geneal, ig, bin = 12){
   
   if(class(ig)!="igraph"){
     stop("ig must be an igraph object")
   }
   
-  if(sum(1:length(binVector)%in%binVector)!=length(binVector)){
-    stop("binVector must contain all numbers 1:length(binVector)")
+  if(class(bin) != "numeric"){
+    stop("bin must contain a number")
   }
   
-  tG <- buildSpreadTotalDF(geneal, ig, binVector)
+  tG <- buildSpreadTotalDF(geneal, ig, bin)
   eG <- igraph::get.data.frame(ig, "edges")
   
   # edgeTotalDF used in function plotPathOnAll()
@@ -280,8 +280,8 @@ buildEdgeTotalDF = function(geneal, ig, binVector = 1:12){
   for (i in 1:numEdges){
     xname = as.character(eG[i,]$from)
     xendname = as.character(eG[i,]$to)
-    x_i = getYear(xname, tG)
-    xend_i = getYear(xendname, tG)
+    x_i = getDate(xname, tG)
+    xend_i = getDate(xendname, tG)
     if(!xname%in%tG$name) {
[…]

### 2016-11-10 article.tex
kinship2 reviewer remarks

@@
 Siblings are connected by horizontal sibship lines.
-Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational breeding occurs, as is often the case in agronomic genealogical lineages.
+Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in many situations where inter-generational breeding occurs, as is often the case in agronomic genealogical lineages.
 
@@
 However, standard pedigree tools only allow for an individual to be displayed once.
-As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset.
+As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, most standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset.
+
+In the \pkg{kinship2} package, if an individual cannot be represented with only one instance, then it will be completely copied and connected with dotted lines to relevant individuals.
+However, the package requires that each child has exactly zero or two parents; if a child has two parents, then one must be female and one must be male.
+These requirements preclude certain genealogical datasets from being plotted by generation count, especially when their complexity increases with inter-generational breeding.
 
@@
 Hence, it should be displayed twice on the vertical axis, once for each of its generation counts.
-However, standard pedigree tools only allow for an individual to be displayed once, leading to ambiguous portrayal of the genealogical dataset.}
+However, most standard pedigree tools only allow for an individual to be displayed once.
+In the \pkg{kinship2} package, individuals can be displayed more than once.
+However, each child must have zero or two parents (one male and one female).
+These restrictions make it impossible to plot genealogical data by generation count in cases where there are many inter-generational breedings.
     \label{fig:kinshipFig}

### 2016-11-10 article.tex
kinship2 second response

@@
 Hence, individual 215 should be displayed in both second and third generational positions on the vertical axis.
-However, standard pedigree tools only allow for an individual to be displayed once.
-As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, most standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset.
+However, most standard pedigree tools only allow for an individual to be displayed once.
+As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, most standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset by generation count.
 
-In the \pkg{kinship2} package, if an individual cannot be represented with only one instance, then it will be completely copied and connected with dotted lines to relevant individuals.
+In the \pkg{kinship2} package, if an individual cannot be represented with only one instance, then it will be completely copied and connected with dotted lines to the relevant individuals.
 However, the package requires that each child has exactly zero or two parents; if a child has two parents, then one must be female and one must be male.
@@
 However, most standard pedigree tools only allow for an individual to be displayed once.
-In the \pkg{kinship2} package, individuals can be displayed more than once.
+In the \pkg{kinship2} package, individuals can indeed be displayed more than once.
 However, each child must have zero or two parents (one male and one female).

### 2016-11-12 JSS-ggenealogy/SourceCode/R/allFunctions.R
Changed QV column to be variable

@@ -255,9 +255,10 @@ buildDesList = function(v1, geneal, gen=0){
 #' @param geneal the full genealogy  (in data frame format)
 #' @param ig the graph representation of the data genealogy (in igraph format)
 #' @param bin the number of bins to determine the vertical positions of nodes (default is 12). For more information on choosing bin size, please visit the ggenealogy vignette.
+#' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
 #' @seealso \code{\link{dfToIG}} for information on producing ig from the genealogy
 #' @seealso \url{http://www.r-project.org} for iGraph information
-buildEdgeTotalDF = function(geneal, ig, bin = 12){
+buildEdgeTotalDF = function(geneal, ig, bin = 12, colName){
   
   if(class(ig)!="igraph"){
     stop("ig must be an igraph object")
@@ -267,7 +268,7 @@ buildEdgeTotalDF = function(geneal, ig, bin = 12){
     stop("bin must contain a number")
   }
   
-  tG <- buildSpreadTotalDF(geneal, ig, bin)
+  tG <- buildSpreadTotalDF(geneal, ig, colName, bin)
   eG <- igraph::get.data.frame(ig, "edges")
   
   # edgeTotalDF used in function plotPathOnAll()
@@ -280,8 +281,8 @@ buildEdgeTotalDF = function(geneal, ig, bin = 12){
   for (i in 1:numEdges){
     xname = as.character(eG[i,]$from)
     xendname = as.character(eG[i,]$to)
-    x_i = getDate(xname, tG)
-    xend_i = getDate(xendname, tG)
+    x_i = getVariable(xname, tG, colName)
+    xend_i = getVariable(xendname, tG, colName)
     if(!xname%in%tG$name) {
       stop(paste(xname, "cannot be found in ig vertices"))
     }
@@ -310,10 +311,11 @@ buildEdgeTotalDF = function(geneal, ig, bin = 12){
 #' @param path path as returned from getPath() or a vector of two variety names which exist in the ig object
 #' @param geneal the full genealogy  (in data frame format)
 #' @param ig the graph representation of the data genealogy (in igraph format)
+#' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
[…]

### 2016-11-12 JSS-ggenealogy/ReplicationScript.R
plotPath(), plotPathOnAll() input changed

@@ -19,25 +19,25 @@ sbIG
 getBasicStatistics(sbIG)
 
 # Determine shortest path of parent-child relationships between two labels of interest ("Tokyo" and "Narow")
-pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
+pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal, "devYear")
 pathTN
 
 # Plot returned path
-plotPath(pathTN)
+plotPath(pathTN, "devYear")
 
 # Do the same for a different pair of two labels of interest ("Bedford" and "Zane"). First, we can determine the years these two labels were identified
-getDate("Bedford", sbGeneal)
-getDate("Zane", sbGeneal)
+getVariable("Bedford", sbGeneal, "devYear")
+getVariable("Zane", sbGeneal, "devYear")
 
 # Next, we can determine the shortest path of parent-child relationships between these two labels of interest and plot it.
-pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
-plotPath(pathBZ, fontFace = 2)
+pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
+plotPath(pathBZ, "devYear", fontFace = 2)
 
 # In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
-plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 
 # We can repeat this process, only now instead of setting the bin variable to 1:3 (as we did earlier), we can set it to 1:6.
-plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Year")
[…]

### 2016-11-12 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -258,7 +258,7 @@ buildDesList = function(v1, geneal, gen=0){
 #' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
 #' @seealso \code{\link{dfToIG}} for information on producing ig from the genealogy
 #' @seealso \url{http://www.r-project.org} for iGraph information
-buildEdgeTotalDF = function(geneal, ig, bin = 12, colName){
+buildEdgeTotalDF = function(geneal, ig, colName, bin = 12){
   
   if(class(ig)!="igraph"){
     stop("ig must be an igraph object")
@@ -331,7 +331,7 @@ buildMinusPathDF = function(path, geneal, ig, colName, bin = 12){
     }
     varieties <- path
     path <- getPath(varieties[1], varieties[2], ig, colName)
-  } else if(sum(names(path)%in%c("pathVertices", "dateVertices"))!=2){
+  } else if(sum(names(path)%in%c("pathVertices", "variableVertices"))!=2){
     stop("path does not appear to be a result of the getPath() function")
   } 
   
@@ -339,7 +339,7 @@ buildMinusPathDF = function(path, geneal, ig, colName, bin = 12){
   eG <- igraph::get.data.frame(ig, "edges")
   
   label=tG$name
-  x=tG$date
+  x=tG[[colName]]
   y=tG$y
   # If the label is part of the path, then we change the its value to NA
   for (i in 1:length(label)){
@@ -371,7 +371,7 @@ buildPathDF = function(path){
     # The labels of the nodes are the names of the varieties in the path
     label=path$pathVertices
     # The x-axis position of the node labels are the dates of the varieties in the path
-    x=as.numeric(path$dateVertices)
+    x=as.numeric(path$variableVertices)
     # The y-axis position of the node labels are incremented by unity for each new connected
     # node in the path
     y=seq(1, length(label), 1)
@@ -433,7 +433,7 @@ buildPlotTotalDF = function(path, geneal, ig, colName, bin = 12){
     }
     varieties <- path
     path <- getPath(varieties[1], varieties[2], ig, colName)
-  } else if(sum(names(path)%in%c("pathVertices", "dateVertices"))!=2){
+  } else if(sum(names(path)%in%c("pathVertices", "variableVertices"))!=2){
[…]

### 2016-11-12 article.tex
plotPath(), plotPathOnAll() input changed
all but animate in replication script

@@
 However, each child must have zero or two parents (one male and one female).
-These restrictions make it impossible to plot genealogical data by generation count in cases where there are many inter-generational breedings.
+These restrictions make it impossible to plot genealogical data by generation count in cases where there are many inter-generational breedings.}
     \label{fig:kinshipFig}
@@
  $ child      : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
- $ date        : num  1981 1948 1910 1912 1968 ...
+ $ devYear    : num  1981 1948 1910 1912 1968 ...
  $ yield      : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
- $ date.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
+ $ yearImputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
  $ parent     : chr  "Hutcheson" "Dunfield" NA "A.K." ...
@@
 \end{CodeChunk}
+
 \subsection{Academic genealogy of statisticians}
@@
 \begin{CodeOutput}
-[1] "child"   "parent"  "date"    "country" "school"  "thesis" 
+[1] "child"    "parent"   "gradYear" "country"  "school"   "thesis" 
 \end{CodeOutput}
@@
 Each case in this quantitative variable column can be of type numeric, type NA, or type ``".
-In both example data files, this column is labeled ``date," and each case in that column can be of type numeric, type NA, or type ``".
+In the first example dataset, columns that could be used for this purpose include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear".
 At this point, any \pkg{ggenealogy} plot can be used.
@@
 The \pkg{ggenealogy} package allows users to select two varieties of interest, and determine the shortest pathway of parent-child relationships between them, using the \code{getPath()} function.
-This will return a list that contains the variety names and their years in the path.
+This will return a list that contains the path, along with the variety name and quantitative variable value of interest for each variety in the path.
+For this example, we will use the development year (from the column ``devYear") as our quantitative variable.
 
@@
 \begin{CodeInput}
-R> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
+R> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal, "devYear")
 R> pathTN
@@
 [1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"   
-$dateVertices
+$variableVertices
 [1] "1907"   "1942"   "1954.5" "1971.5" "1985"   
@@
 \begin{Code}
-R> plotPath(pathTN)
+plotPath(pathTN, "devYear")
 \end{Code}
@@
 Next, we can run the same set of functions on a different pair of varieties.
-First, a call to the \pkg{ggenealogy} function \code{getDate()} indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.
+First, we can call the \pkg{ggenealogy} function \code{getVariable()} using the input quantitative variable of development year.
+This indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.
 
@@
 \begin{CodeInput}
-R> getDate("Bedford", sbGeneal)
+R> getVariable("Bedford", sbGeneal, "devYear")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> getDate("Zane", sbGeneal)
+R> getVariable("Zane", sbGeneal, "devYear")
 \end{CodeInput}
@@
 As this is a longer path, we may also consider setting the \code{fontFace} variable of the \code{plotPath()} to a value of 2, indicating we wish to boldface the two varieties of interest.
+In addition, as is the case with plotting tools in \pkg{ggenealogy}, we can append \pkg{ggplot2} syntax.
+In this case, we may wish to hard code the x-axis label from its default value of ``devYear" (the inputted quantitative variable column name) to the more readable ``Development Year".
 
 \begin{Code}
-R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal)
-R> plotPath(pathBZ, fontFace = 2)
+R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
+R> plotPath(pathBZ, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
 \end{Code}
@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 3, pathEdgeCol = "red",
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red",
 +    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 6, pathEdgeCol =
-+    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Year")
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol =
++    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development Year")
 \end{Code}
@@
 
-Notice from Figure~\ref{fig:plotTNBin3} that the default horizontal axis label for the \texttt{plotPath()} method has a value of ``Date".
-Given that the ``date" variable in this example dataset is on the timescale of years, we wanted to change the default value of the horizontal axis label to ``Year".
+Notice again from Figure~\ref{fig:plotTNBin3} that the default horizontal axis label for the \texttt{plotPath()} method has a value of ``devYear".
+We wanted to change the default value of the horizontal axis label to ``Development Year".
 We did this in the code above for Figure~\ref{fig:plotTNBin6} by appending the \texttt{ggplot2::xlab()} function.
@@
 +    "Pella", "Tokyo", "Young", "Zane")
->R plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") +
+>R plotDegMatrix(varieties, sbIG, sbGeneal, "devYear") +
 +    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
 +    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
-+    legend.text = ggplot2::element_text(size = 15))
++    legend.text = ggplot2::element_text(size = 15)) +
++    ggplot2::labs(x = "Variety", y = "Variety")
 \end{Code}
@@
 R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
-+    isDirected = FALSE)
-R> plotPath(pathCB, fontFace = 4) + ggplot2::xlab("Year") +
++    "gradYear", isDirected = FALSE)
+R> plotPath(pathCB, "gradYear", fontFace = 4) + ggplot2::xlab("Graduation Year") +
 +    ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title =
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200)
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200)
 +    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
 +    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-+    (expand = c(.1, .2)) + ggplot2::xlab("Year")
++    (expand = c(.1, .2)) + ggplot2::xlab("Graduation Year")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize = .5,
 +    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
@@
 +    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-+    (expand = c(.1, .2)) + ggplot2::xlab("Year")
++    (expand = c(.1, .2)) + ggplot2::xlab("Graduation Year")
 \end{CodeInput}

### 2016-11-12 JSS-ggenealogy/ReplicationScript.R

@@ -31,20 +31,20 @@ getVariable("Zane", sbGeneal, "devYear")
 
 # Next, we can determine the shortest path of parent-child relationships between these two labels of interest and plot it.
 pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
-plotPath(pathBZ, "devYear", fontFace = 2)
+plotPath(pathBZ, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
 
 # In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
 plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
 
 # We can repeat this process, only now instead of setting the bin variable to 1:3 (as we did earlier), we can set it to 1:6.
-plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Year")
+plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development Year")
 
 # As is explained in the article, only the top part of Figure 6 (the figure from this section) is produced by ggenealogy code. In contrast, the bottom part of Figure 6 was produced by tools outside of ggenealogy for didactic purposes. Below, we recreate the top part of Figure 6, which was to generate a plot of the ancestors and descendants of the label Lee.
 plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
 
 # We can plot the distance matrix for a set of 10 varieties.
 varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow", "Pella", "Tokyo", "Young", "Zane")
-plotDegMatrix(varieties, sbIG, sbGeneal, "Variety", "Variety", "Degree") + ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") + ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 15))
[…]

### 2016-11-12 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -652,16 +652,16 @@ getChild = function(v1, geneal){
 #' @examples
 #' data(sbGeneal)
 #' ig <- dfToIG(sbGeneal)
-#' getDegree("Brim", "Bedford", ig, sbGeneal, "devYear")
+#' getDegree("Brim", "Bedford", ig, sbGeneal)
 #' @export
 getDegree = function(v1, v2, ig, geneal, colName){
   if(is.null(geneal)){
-    stop("Please input a genealogy data frame where the first two columns are nodes and at least one other column is your inputted quantitative variable colName")
+    stop("Please input a genealogy data frame")
   }
   if(is.null(ig)){
     stop("Please input an igraph object formatted by dfToIG()")
   }
-  path <- getPath(v1=v1, v2=v2, ig=ig, geneal = geneal, colName = colName, isDirected=F)
+  path <- getPath(v1=v1, v2=v2, ig=ig, geneal = geneal, isDirected=F, colName = colName)
   # The degree between two vertices is equal to one less than the number of nodes in the shortest path
   return(length(path$pathVertices)-1)
 }
@@ -964,20 +964,17 @@ plotAncDes = function(v1, geneal, mAnc=3, mDes=3, vColor="#D35C79"){
 #' @param ig the graph representation of the data genealogy (in igraph format)
 #' @param geneal the full genealogy  (in data frame format)
 #' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
-#' @param xLab string label on the x axis (default is "Variety")
-#' @param yLab string label on the y axis (default is "Variety")
-#' @param legendLab string label on the legend (default is "Degree")
 #' 
 #' @seealso \url{http://www.r-project.org} for iGraph information
 #' @examples
 #' data(sbGeneal)
 #' ig <- dfToIG(sbGeneal)
 #' varieties <- c("Bedford", "Calland", "Narow", "Pella", "Tokyo", "Young", "Zane")
-#' p <- plotDegMatrix(varieties, ig, sbGeneal, "devYear", "Soybean label", "Soybean label", "Degree")
+#' p <- plotDegMatrix(varieties, ig, sbGeneal, "devYear")
 #' p + ggplot2::scale_fill_continuous(low = "white", high = "darkgreen")
 #' 
 #' @export
[…]

### 2016-11-15 JSS-ggenealogy/ReplicationScript.R
Yield for plotPath()

@@ -39,6 +39,23 @@ plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red", n
 # We can repeat this process, only now instead of setting the bin variable to 1:3 (as we did earlier), we can set it to 1:6.
 plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development Year")
 
+##########################################################
+
+row.has.na <- apply(sbGeneal, 1, function(x){any(is.na(x))})
+sbFilt <- sbGeneal[!row.has.na,]
+dim(sbFilt)
+sbFiltIG <- dfToIG(sbFilt)
+sbFiltIG
+getBasicStatistics(sbFiltIG)
+
+pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
+pathCL
+plotPath(pathCL, "yield") + ggplot2::xlab("Yield")
+
+
+
+##########################################################
+
 # As is explained in the article, only the top part of Figure 6 (the figure from this section) is produced by ggenealogy code. In contrast, the bottom part of Figure 6 was produced by tools outside of ggenealogy for didactic purposes. Below, we recreate the top part of Figure 6, which was to generate a plot of the ancestors and descendants of the label Lee.
 plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
 

### 2016-11-15 JSS-ggenealogy/ReplicationScript.R
Fixed misaligned plotPath edges

@@ -52,7 +52,7 @@ pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
 pathCL
 plotPath(pathCL, "yield") + ggplot2::xlab("Yield")
 
-
+plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 4, pathEdgeCol = "purple")
 
 ##########################################################
 

### 2016-11-15 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -1107,6 +1107,16 @@ plotPathOnAll = function(path, geneal, ig, colName, bin = 12, edgeCol = "gray84"
   )
   
   textFrame <- stats::na.omit(textFrame) #remove any row that has at least one NA
+  eTDF <- eTDF[(eTDF$x %in% textFrame$x) & (eTDF$xend %in% textFrame$x),]
+  
+  rowETDF <- rownames(eTDF)
+  rowTextFrame <- rownames(textFrame)
+  rownames(eTDF) <- 1:nrow(eTDF)
+  rownames(textFrame) <- 1:nrow(textFrame)
+  eTDF$y <- textFrame[match(eTDF$x, textFrame$x),]$y
+  eTDF$yend <- textFrame[match(eTDF$xend, textFrame$x),]$y
+  rownames(eTDF) <- rowETDF
+  rownames(textFrame) <- rowTextFrame
   
   # The plotTotalImage object creates two line segments (geom_segment), one to create grey
   # edges for non-path connections between pairs of nodes, the other to create light-green
@@ -1114,7 +1124,8 @@ plotPathOnAll = function(path, geneal, ig, colName, bin = 12, edgeCol = "gray84"
   # create labels of size 2 for non-path connections between pairs of nodes, the other to
   # create labels of size 2.5 and boldfaced for path connections between pairs of nodes.
   plotTotalImage = ggplot2::ggplot(data = pMPDF, ggplot2::aes(x = x, y = y)) +
-    ggplot2::geom_segment(data = eTDF, ggplot2::aes(x=x, y=y-.1, xend=xend, yend=yend+.1), colour = edgeCol) +
+    #ggplot2::geom_segment(data = eTDF, ggplot2::aes(x=x, y=y-.1, xend=xend, yend=yend+.1), colour = edgeCol) +
+    ggplot2::geom_segment(data = eTDF, ggplot2::aes(x=x, y=y, xend=xend, yend=yend), colour = edgeCol) +
     ggplot2::geom_segment(data = pTDF, ggplot2::aes(x=xstart, y=ystart, xend=xend, yend=yend), colour = pathEdgeCol, size = 1) +
       ggplot2::geom_text(data = textFrame, ggplot2::aes(x = x, y = y, label = label), size = nodeSize, colour = nodeCol)
       ggplot2::geom_text(data = textFrame, ggplot2::aes(x = x, y = y, label = label), size = nodeSize, colour = nodeCol)

### 2016-11-15 JSS-ggenealogy/SourceCode/R/allFunctions.R
eTDF factor problem fixed

@@ -277,6 +277,8 @@ buildEdgeTotalDF = function(geneal, ig, colName, bin = 12){
   y=as.numeric(rep("",numEdges))
   xend=as.numeric(rep("",numEdges))
   yend=as.numeric(rep("",numEdges))
+  name=as.numeric(rep("",numEdges))
+  nameEnd=as.numeric(rep("",numEdges))
   # For each edge in the graph
   for (i in 1:numEdges){
     xname = as.character(eG[i,]$from)
@@ -295,9 +297,11 @@ buildEdgeTotalDF = function(geneal, ig, colName, bin = 12){
     xend[i] = xend_i
     y[i] = y_i
     yend[i] = yend_i
+    name[i] = xname
+    nameEnd[i] = xendname
   }
   # Create a dataframe containing the start and end positions of the x and y axes
-  edgeTotalDF = as.data.frame(cbind(x, y, xend, yend))
+  edgeTotalDF = as.data.frame(cbind(x, y, xend, yend, name, nameEnd))
   
   edgeTotalDF
 }
@@ -341,7 +345,7 @@ buildMinusPathDF = function(path, geneal, ig, colName, bin = 12){
   label=tG$name
   x=tG[[colName]]
   y=tG$y
-  # If the label is part of the path, then we change the its value to NA
+  # If the label is part of the path, then we change its value to NA
   for (i in 1:length(label)){
     if (label[i]%in%path$pathVertices){
       label[i]=NA
@@ -1113,10 +1117,13 @@ plotPathOnAll = function(path, geneal, ig, colName, bin = 12, edgeCol = "gray84"
   rowTextFrame <- rownames(textFrame)
   rownames(eTDF) <- 1:nrow(eTDF)
   rownames(textFrame) <- 1:nrow(textFrame)
-  eTDF$y <- textFrame[match(eTDF$x, textFrame$x),]$y
-  eTDF$yend <- textFrame[match(eTDF$xend, textFrame$x),]$y
+  eTDF$y <- textFrame[match(eTDF$name, textFrame$label),]$y
+  eTDF$yend <- textFrame[match(eTDF$nameEnd, textFrame$label),]$y
+  eTDF <- eTDF[,-c(5,6)]
   rownames(eTDF) <- rowETDF
   rownames(textFrame) <- rowTextFrame
+  eTDF$x <- as.integer(as.character(eTDF$x))
+  eTDF$xend <- as.integer(as.character(eTDF$xend)) 
   
[…]

### 2016-11-15 JSS-ggenealogy/ReplicationScript.R
Example of new qv in plotPathOnAll() working

@@ -41,18 +41,25 @@ plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol = "seagree
 
 ##########################################################
 
-row.has.na <- apply(sbGeneal, 1, function(x){any(is.na(x))})
-sbFilt <- sbGeneal[!row.has.na,]
+#works
+# row.has.na <- apply(sbGeneal, 1, function(x){any(is.na(x))})
+# sbFilt <- sbGeneal[!row.has.na,]
+# dim(sbFilt)
+# sbFiltIG <- dfToIG(sbFilt)
+# pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
+# pathCL
+# plotPath(pathCL, "yield") + ggplot2::xlab("Yield")
+# plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 4, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
+
+sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
 dim(sbFilt)
 sbFiltIG <- dfToIG(sbFilt)
-sbFiltIG
-getBasicStatistics(sbFiltIG)
-
 pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
 pathCL
 plotPath(pathCL, "yield") + ggplot2::xlab("Yield")
+plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
 
-plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 4, pathEdgeCol = "purple")
+
 
 ##########################################################
 

### 2016-11-15 JSS-ggenealogy/SourceCode/R/allFunctions.R
Two variables for plotPathOnAll()

@@ -114,7 +114,7 @@ buildAncDesCoordDF = function(df){
 #' 
 #' Returns the data frame that includes labels and plot coordinates of all ancestors and descendants of a variety. Users can specify the maximum number of ancestors and descendants to display.
 #' 
-#' @param v1 the label of the variety/vertex of interest (in character string format)
+#' @param v1 the label of the vertex of interest (in character string format)
 #' @param mAnc the maximum number of generations of ancestors of v1 to be displayed (in numeric format)
 #' @param mDes the maximum number of generations of descendants of v1 to be displayed (in numeric format)
 #' @param geneal the full genealogy (in data frame format)
@@ -186,7 +186,7 @@ buildAncDesTotalDF = function(v1, geneal, mAnc=3, mDes=3){
 #' 
 #' This function returns a nested list of the ancestors of the inputted variety.
 #' 
-#' @param v1 the label of the variety/vertex of interest (in character string format)
+#' @param v1 the label of the vertex of interest (in character string format)
 #' @param geneal the full genealogy  (in data frame format)
 #' @param gen the generation (note: This should be left as default, as any other input will not affect results anyway)
 #' @seealso \code{\link{getParent}} for information on determining parents
@@ -219,7 +219,7 @@ buildAncList = function(v1, geneal, gen = 0){
 #' 
 #' This function returns a nested list of the descendants of the inputted variety.
 #' 
-#' @param v1 the label of the variety/vertex of interest (in character string format)
+#' @param v1 the label of the vertex of interest (in character string format)
 #' @param geneal the full genealogy  (in data frame format)
 #' @param gen the generation (note: This should be left as default, as any other input will not affect results)
 #' @seealso \code{\link{getChild}} for information on determining children
@@ -316,10 +316,11 @@ buildEdgeTotalDF = function(geneal, ig, colName, bin = 12){
 #' @param geneal the full genealogy  (in data frame format)
 #' @param ig the graph representation of the data genealogy (in igraph format)
 #' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
+#' @param colNameY the name of the second optional column of the data frame that contains the second optional quantitative variable of interest (in character string format). This optional quantitative variable will be plotted on the vertical axis.
[…]

### 2016-11-16 JSS-ggenealogy/ReplicationScript.R
plotPath() 2 Qvs

@@ -23,7 +23,7 @@ pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal, "devYear")
 pathTN
 
 # Plot returned path
-plotPath(pathTN, "devYear")
+plotPath(pathTN, sbGeneal, "devYear")
 
 # Do the same for a different pair of two labels of interest ("Bedford" and "Zane"). First, we can determine the years these two labels were identified
 getVariable("Bedford", sbGeneal, "devYear")
@@ -31,7 +31,7 @@ getVariable("Zane", sbGeneal, "devYear")
 
 # Next, we can determine the shortest path of parent-child relationships between these two labels of interest and plot it.
 pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
-plotPath(pathBZ, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
+plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
 
 # In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
 plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
@@ -41,26 +41,15 @@ plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol = "seagree
 
 ##########################################################
 
-#works
-# row.has.na <- apply(sbGeneal, 1, function(x){any(is.na(x))})
-# sbFilt <- sbGeneal[!row.has.na,]
-# dim(sbFilt)
-# sbFiltIG <- dfToIG(sbFilt)
-# pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
-# pathCL
-# plotPath(pathCL, "yield") + ggplot2::xlab("Yield")
-# plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 4, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
-
 sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
-dim(sbFilt)
[…]

### 2016-11-16 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -381,7 +381,10 @@ buildMinusPathDF = function(path, geneal, ig, colName, colNameY, bin = 12){
 #' to connect to the node at the next largest y-value), "yend" (the y-axis position
 #' of the outgoing edge (connected to the node at the next largest y-value))).
 #' @param path path object representing the path between two vertices
-buildPathDF = function(path){
+#' @param geneal the full genealogy  (in data frame format)
+#' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
+#' @param colNameY the name of the second optional column of the data frame that contains the second optional quantitative variable of interest (in character string format). This optional quantitative variable will be plotted on the vertical axis.
+buildPathDF = function(path, geneal, colName, colNameY=""){
   if(length(path) > 0){
     # The labels of the nodes are the names of the varieties in the path
     label=path$pathVertices
@@ -389,14 +392,18 @@ buildPathDF = function(path){
     x=as.numeric(path$variableVertices)
     # The y-axis position of the node labels are incremented by unity for each new connected
     # node in the path
-    y=seq(1, length(label), 1)
+    if (colNameY==""){
+      y=seq(1, length(label), 1)
+    }
+    else{
+      y =geneal[match(label, geneal$child),][[colNameY]]
+    }
     # The starting x-axis position of the edge between two nodes will be the x-axis position
     # of the label of the first node
     xstart=x
     xend=rep(0,length(label))
     yend=rep(0,length(label))
     for (i in 1:length(label)){
-      y[i] = i
       xend[i] = xstart[i]
       if (i < length(label)){
         # The ending x-axis position of the edge between two nodes will be the x-axis position
@@ -1027,7 +1034,9 @@ plotDegMatrix = function(varieties,ig,geneal,colName){
 #' parent-child relationships between those nodes. For visual appeal, there is a grey
 #' box that outlines the node label, as well as an underline and overline for each label.
 #' @param path object created from function getPath
+#' @param geneal the full genealogy  (in data frame format)
[…]

### 2016-11-17 article.tex
multivariate animation in plotPathOnAll()

@@
 \begin{Code}
-plotPath(pathTN, "devYear")
+plotPath(pathTN, sbGeneal, "devYear")
 \end{Code}
@@
 R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
-R> plotPath(pathBZ, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
+R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
 \end{Code}

### 2016-11-17 JSS-ggenealogy/ReplicationScript.R

@@ -45,11 +45,11 @@ sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
 sbFiltIG <- dfToIG(sbFilt)
 pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
 plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
+pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
+plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
 plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
 plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
 
-pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
-plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
 ##########################################################
 
 # As is explained in the article, only the top part of Figure 6 (the figure from this section) is produced by ggenealogy code. In contrast, the bottom part of Figure 6 was produced by tools outside of ggenealogy for didactic purposes. Below, we recreate the top part of Figure 6, which was to generate a plot of the ancestors and descendants of the label Lee.
@@ -80,7 +80,7 @@ statIG <- dfToIG(statGeneal)
 pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal, "gradYear", isDirected = FALSE)
 
 # After doing so, we can now determine the shortest path between Sir David Cox and Petra Buzkova, and plot it.
-plotPath(pathCB, "gradYear", fontFace = 4) + ggplot2::xlab("Graduation Year") + ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, 1.5))
[…]

### 2016-11-17 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -1194,12 +1194,22 @@ plotPathOnAll = function(path, geneal, ig, colName, colNameY = "", bin = 12, edg
   }
   # Return the animatePlotTotalImage, if animate is TRUE
   else{
-    animatePlotTotalImage <- plotly::plotly_build(plotly::ggplotly(plotTotalImage, tooltip = c("x", "label")))
-    animatePlotTotalImage$data[[1]]$hoverinfo <- "none"
-    animatePlotTotalImage$data[[2]]$hoverinfo <- "none"
-    animatePlotTotalImage$data[[3]]$hoverinfo <- c("x+text")
-    animatePlotTotalImage$data[[4]]$hoverinfo <- c("x+text")
-    animatePlotTotalImage
+    if (colNameY == ""){
+      animatePlotTotalImage <- plotly::plotly_build(plotly::ggplotly(plotTotalImage, tooltip = c("x", "label")))
+      animatePlotTotalImage$data[[1]]$hoverinfo <- "none"
+      animatePlotTotalImage$data[[2]]$hoverinfo <- "none"
+      animatePlotTotalImage$data[[3]]$hoverinfo <- c("x+text")
+      animatePlotTotalImage$data[[4]]$hoverinfo <- c("x+text")
+      animatePlotTotalImage 
+    }
+    else{
+      animatePlotTotalImage <- plotly::plotly_build(plotly::ggplotly(plotTotalImage, tooltip = c("x", "label")))
+      animatePlotTotalImage$data[[1]]$hoverinfo <- "none"
+      animatePlotTotalImage$data[[2]]$hoverinfo <- "none"
+      animatePlotTotalImage$data[[3]]$hoverinfo <- c("x+y+text")
+      animatePlotTotalImage$data[[4]]$hoverinfo <- c("x+y+text")
+      animatePlotTotalImage       
+    }
   }
 }
 

### 2016-11-17 article.tex
New plots in script

@@
 For these plots, the input data frame should also contain a third quantitative variable column.
-Each case in this quantitative variable column can be of type numeric, type NA, or type ``".
-In the first example dataset, columns that could be used for this purpose include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear".
+Each case in this quantitative variable column should be of type numeric.
+In the first example dataset, columns that could be used for this purpose (once subsetted to filter out values of ``'' and NA) include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear".
 At this point, any \pkg{ggenealogy} plot can be used.
@@
 
+We would like to note that the \code{plotPath()} method can be created with one or two quantitative variables that the user hard-codes.
+We illustrate this with the \code{sbGeneal} data frame after we subset it to remove observations that do not have values for the quantitative variable ``yield".
+Then, we can examine the path between the varieties \code{Clark} and \code{Lawrence} and plot how the quantitative varible ``yield" changes along the parent-child relationships of that path.
+Likewise, we plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path.
+The output of these two calls to the \code{plotPath()} function can be viewed in Figure \ref{fig:plotPathQV2}.
+
+\begin{Code}
+R> sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
+R> sbFiltIG <- dfToIG(sbFilt)
+R> pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
+R> plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
+R> pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
+R> plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
+\end{Code}
+
+\begin{figure}[h]
+    \centering
+    \fbox{\includegraphics[width=\textwidth]{plotPathQV2}}
+    \caption{Left: Plotting a path of interest using a new quantitative variable of interest, ``Yield".
+Right: Plotting a path of interest using two quantitative variables.
+We see that the varieties \code{Clark} and \code{Lawrence} have lower yields than the varieties in the middle of the path.}
+    \label{fig:plotPathQV2}
+\end{figure}
+
 \section{Superimposing shortest path on tree}
@@
 In the code above, syntax from the \pkg{ggplot2} package was appended to the \code{plotPathOnAll()} function; this can be done for most \pkg{ggenealogy} functions.
-While the first three explicit parameters have been introduced earlier in this paper, the fourth parameter (\code{bin}) requires some explanation.
-The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, year of release).
-As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels.
+While the first four explicit parameters have been introduced earlier in this paper, the fifth parameter (\code{bin}) requires some explanation.
+The motivation of the \code{plotPathOnAll()} function is to write node labels on a plot, with the center of each node label constricted on the horizontal axis to its quantitative variable of interest (in this case, development year).
+As is the case for the plots before, the vertical axis has no meaning other than providing a plotting area in which to draw the node labels (unless a user specifies a second quantitative variable of interest, as we will demonstrate later).
 Unfortunately, for large datasets, this motivation can be a difficult task because the text labels of the varieties can overlap if they are assigned a similar y coordinate, have a similar year (x coordinate), and have long text labels (width of x coordinate).
@@
 We did this in the code above for Figure~\ref{fig:plotTNBin6} by appending the \texttt{ggplot2::xlab()} function.
+
+We would like to emphasize that the \code{plotPathOnAll()} method can be created with one or two quantitative variables that the user hard-codes.
+We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created by subsetting the soy bean genealogy data to only include observations that contain a value for the quantitative variable ``yield".
+Then, we can examine the path across the entire genealogical structure and how the quantitative varible ``yield" changes along the parent-child relationships of that path.
+Likewise, we can plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path.
+The output of these two calls to the \code{plotPathOnAll()} function can be viewed in Figure \ref{fig:plotPathOnAllQV2}.
+
+\begin{Code}
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
+\end{Code}
+
+\begin{figure}[h]
+    \centering
+    \fbox{\includegraphics[width=\textwidth]{plotPathOnAllQV2}}
+    \caption{Left: Plotting a path of interest across the genealogical structure using a new quantitative variable of interest, ``Yield".
+Right: Plotting a path of interest across the genealogical structure using two quantitative variables.
+We see that an increase in development year might correlate to an increase in yield for the varieties in this dataset, and that our path of interest is composed of generally younger varieties compared to the dataset as a whole.}
+    \label{fig:plotPathOnAllQV2}
+\end{figure}
 
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize = .5,
 +    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
@@
 \end{figure}
+
+We would like to state that users can still hard code one or two variables in the \code{plotPathOnAll()} function, even with the \code{animate} option set to a value of ``TRUE".
+The first call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis.
+When hovering over labels, the label name and yield value is revealed.
+The second call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis and development year on the vertical axis.
+When hoering over labels, the label name, yield value, and development year is revealed.
+
+\begin{Code}
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange", animate = TRUE)
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange", animate = TRUE)
+\end{Code}
 

### 2016-11-18 JSS-ggenealogy/SourceCode/R/allFunctions.R
Fix case of null getAncestors() data.frame

@@ -556,6 +556,9 @@ buildSpreadTotalDF = function(geneal, ig, colName, bin = 12){
 #' getAncestors("Essex", sbGeneal, 5)
 getAncestors = function(v1, geneal, gen = 3){
   id.offset <- NULL
+  if (is.null(buildAncList(v1, geneal))){
+    return(data.frame())
+  }
   aDF = buildAncDesCoordDF(nodeToDF(buildAncList(v1, geneal)))
   subDF = aDF[aDF$gen <= gen & aDF$gen != 0,]
   keep = c("label","gen")

### 2016-11-18 JSS-ggenealogy/testBranches.R
getBranchStatistics - no NA, not including child value

@@ -1,0 +1,55 @@
+rm(list=ls())
+library("devtools")
+load_all("SourceCode")
+library("ggenealogy")
+data("sbGeneal")
+
+getBranches = function(v1, geneal, ancGen = 3, desGen = 3){
+  id.offset <- NULL
+  if (!is.null(buildAncList(v1, geneal))){
+    aDF = buildAncDesCoordDF(nodeToDF(buildAncList(v1, geneal)))
+    subDF = aDF[aDF$gen <= ancGen & aDF$gen != 0,]
+    keep = c("label","gen")
+    subDF = subDF[keep]
+    row.names(subDF) = NULL
+    aDF <- subDF[order(subDF$gen,subDF$label),]
+  }
+  id.offset <- NULL
+  if (!is.null(buildDesList(v1, geneal))){
+    dDF = buildAncDesCoordDF(nodeToDF(buildDesList(v1, geneal)))
+    subDF = dDF[dDF$gen <= desGen & dDF$gen != 0,]
+    keep = c("label","gen")
+    subDF = subDF[keep]
+    row.names(subDF) = NULL
+    dDF <- subDF[order(subDF$gen,subDF$label),]
+  }
+  allDF <- rbind(aDF,dDF)
+  if (is.null(allDF)){
+    return(data.frame())
+  }
+  return(allDF)
+}
+
+
+
+sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
+
+getBranchStatistics = function(v1, geneal, colName, gen=3){
+  id.offset <- NULL
+  if (is.null(getChild(v1, geneal))){
+    return(data.frame())
+  }
+  childList <- getChild(v1,geneal)
+  datRet <- data.frame()
+  for (i in 1:length(childList)){
+    dat <- getDescendants(childList[i], geneal, gen)
+    dat[[colName]] <- geneal[match(dat$label, geneal$child),][[colName]]
+    Labels <- dat$label
+    Count <- length(Labels)
+    Mean <- mean(dat[[colName]])
+    SD <- sd(dat[[colName]])
+    datRet <- rbind(datRet, data.frame(Name=childList[i],Count=Count, Mean=Mean, SD=SD, Labels=paste(Labels, collapse = ',')))
+  }
+  return(datRet) 
+}
+

### 2016-11-19 JSS-ggenealogy/testBranches.R
GetBranchStatQuant()

@@ -3,38 +3,13 @@ library("devtools")
 load_all("SourceCode")
 library("ggenealogy")
 data("sbGeneal")
+data("statGeneal")
+library(stringr) # for str_count
 
-getBranches = function(v1, geneal, ancGen = 3, desGen = 3){
-  id.offset <- NULL
-  if (!is.null(buildAncList(v1, geneal))){
-    aDF = buildAncDesCoordDF(nodeToDF(buildAncList(v1, geneal)))
-    subDF = aDF[aDF$gen <= ancGen & aDF$gen != 0,]
-    keep = c("label","gen")
-    subDF = subDF[keep]
-    row.names(subDF) = NULL
-    aDF <- subDF[order(subDF$gen,subDF$label),]
-  }
-  id.offset <- NULL
-  if (!is.null(buildDesList(v1, geneal))){
-    dDF = buildAncDesCoordDF(nodeToDF(buildDesList(v1, geneal)))
-    subDF = dDF[dDF$gen <= desGen & dDF$gen != 0,]
-    keep = c("label","gen")
-    subDF = subDF[keep]
-    row.names(subDF) = NULL
-    dDF <- subDF[order(subDF$gen,subDF$label),]
-  }
-  allDF <- rbind(aDF,dDF)
-  if (is.null(allDF)){
-    return(data.frame())
-  }
-  return(allDF)
-}
+v1="Tokyo";geneal=sbGeneal;colName="devYear";gen=15
+v1="David Cox";geneal=statGeneal;colName="gradYear";gen=15
 
-
-
-sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
-
-getBranchStatistics = function(v1, geneal, colName, gen=3){
+getBranchStatQuant = function(v1, geneal, colName, gen=3){
   id.offset <- NULL
   if (is.null(getChild(v1, geneal))){
     return(data.frame())
@@ -46,10 +21,69 @@ getBranchStatistics = function(v1, geneal, colName, gen=3){
     dat[[colName]] <- geneal[match(dat$label, geneal$child),][[colName]]
     Labels <- dat$label
     Count <- length(Labels)
-    Mean <- mean(dat[[colName]])
-    SD <- sd(dat[[colName]])
-    datRet <- rbind(datRet, data.frame(Name=childList[i],Count=Count, Mean=Mean, SD=SD, Labels=paste(Labels, collapse = ',')))
+    Mean <- mean(dat[[colName]], na.rm=TRUE)
[…]

### 2016-11-19 JSS-ggenealogy/testBranches.R
Branch examples

@@ -38,12 +38,7 @@ CoxBranches <- getBranchStatQuant("David Cox", statGeneal, "gradYear", 15)
 ##############################################################################
 ##############################################################################
 
-
-v1="David Cox";geneal=statGeneal;colName="thesis";gen=15
-
-rExpr = "grepl('Stochastic', geneal$colName)"
-
-getBranchStatQual = function(v1, rExpr, gen=3){
+getBranchStatQual = function(v1, geneal, colName, rExpr, gen=3){
   id.offset <- NULL
     if (is.null(getChild(v1, geneal))){
     return(data.frame())
@@ -55,35 +50,34 @@ getBranchStatQual = function(v1, rExpr, gen=3){
     dat[[colName]] <- geneal[match(dat$label, geneal$child),][[colName]]
     Labels <- dat$label
     Count <- length(Labels)
-    
     rExpr = gsub("geneal[$]colName", "dat[[colName]]", rExpr)
-    
-    eval(parse(text=rExpr))
-    
-    Mean <- mean(eval(parse(text=rExpr)),na.rm=TRUE)
-    SD <- sd(eval(parse(text=rExpr)),na.rm=TRUE)
+    CountTrue <- sum(eval(parse(text=rExpr)),na.rm=TRUE)
     NACount <- sum(is.na(eval(parse(text=rExpr))))
     PercentNotNA <- round(100*(1-(NACount/Count)),digits=2)
-    datRet <- rbind(datRet, data.frame(Name=childList[i], Mean=Mean, SD=SD, Count=Count, NACount=NACount, PercentNotNA=PercentNotNA, Labels=paste(Labels, collapse = ',')))
+    datRet <- rbind(datRet, data.frame(Name=childList[i], CountTrue = CountTrue, Count=Count, NACount=NACount, PercentNotNA=PercentNotNA, Labels=paste(Labels, collapse = ',')))
   }
-  datRet <- datRet[order(-datRet$Mean),]
+  datRet <- datRet[order(-datRet$CountTrue),]
   return(datRet)
 }
 
-grepl("Stochastic", statGeneal$thesis) #150 \\147
-rExpr = "grepl('Stochastic', statGeneal$thesis)" #works
+v1="David Cox";geneal=statGeneal;colName="thesis";gen=15
+rExpr = "grepl('Stochastic', geneal$colName)"
[…]

### 2016-11-20 article.tex
testBranches.R revisions

@@
 
-\subsection{Interactive plotting functions}
+\section{Interactive plotting functions}
 \label{interaction}
@@
 The second call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis and development year on the vertical axis.
-When hoering over labels, the label name, yield value, and development year is revealed.
+When hovering over labels, the label name, yield value, and development year is revealed.
 
@@
 \end{Code}
+
+\section{Branch calculations and parsings}
+
+It may be helpful for users to search through the descendant branches of an individual of interest to compare and contrast how a variable of interest changes within these branches.
+For instance, which descending branches of a particular soybean variety are producing the highest yields?
+Which are developing new varieties in recent years?
+Which descending branches of a particular academic statistician have large proportions of individuals graduating from a certain school or country?
+What propotion have thesis titles containing a word of interest?
+Which are graduating new students in recent years?
 

### 2016-11-20 JSS-ggenealogy/testBranches.R

@@ -52,7 +52,7 @@ getBranchStatQual = function(v1, geneal, colName, rExpr, gen=3){
     Count <- length(Labels)
     rExpr = gsub("geneal[$]colName", "dat[[colName]]", rExpr)
     CountTrue <- sum(eval(parse(text=rExpr)),na.rm=TRUE)
-    NACount <- sum(is.na(eval(parse(text=rExpr))))
+    NACount <- sum(dat[[colName]]=="")
     PercentNotNA <- round(100*(1-(NACount/Count)),digits=2)
     datRet <- rbind(datRet, data.frame(Name=childList[i], CountTrue = CountTrue, Count=Count, NACount=NACount, PercentNotNA=PercentNotNA, Labels=paste(Labels, collapse = ',')))
   }
@@ -62,17 +62,17 @@ getBranchStatQual = function(v1, geneal, colName, rExpr, gen=3){
 
 v1="David Cox";geneal=statGeneal;colName="thesis";gen=15
 rExpr = "grepl('Stochastic', geneal$colName)"
-DCBranches_Stochastic <- getBranchStatQual(v1, geneal, colName, rExpr, gen)
+DCBranches_Sto <- getBranchStatQual(v1, geneal, colName, rExpr, gen)
 
 rExpr = "grepl('(?i)Stochastic', geneal$colName)"
-DCBranches_StochasticCaps <- getBranchStatQual(v1, geneal, colName, rExpr, gen)
+DCBranches_StoCaps <- getBranchStatQual(v1, geneal, colName, rExpr, gen)
 
 colName = "country"
 rExpr = "geneal$colName=='UnitedKingdom'" #works
 DCBranches_UK <- getBranchStatQual(v1, geneal, colName, rExpr, gen)
 
 desDC <- getDescendants("David Cox", statGeneal,15)
-sort(table(statGeneal[match(desDC$label,statGeneal$child),]$school))
+tail(sort(table(statGeneal[match(desDC$label,statGeneal$child),]$school)),5)
 sum(statGeneal[match(getChild("David Cox",statGeneal),statGeneal$child),]$school=="University of London")
 
 colName = "school"

### 2016-11-21 article.tex
Remove percentNA

--- 2016-11-17 article.tex
@@
 \end{Code}
+
+\section{Branch parsing and calculations}
+
+It may be helpful for users to search through descendant branches of a certain individual to compare and contrast how a variable of interest changes along these branches.
+For instance, which descending branches of a particular soybean variety are producing the highest yields?
+Which branches are developing new varieties in recent years?
+Which descending branches of a particular academic statistician have large proportions of students graduating from certain schools or countries?
+Which branches are graduating new students in recent years?
+Which branches have the highest proportion of students with thesis titles containing a word of interest?
+
+Answering these questions in a straightforward manner requires more than basic data frame manipulation; it also requires methods that can easily traverse parent-child relationships.
+The \pkg{ggenealogy} package has two methods that can easily answer these questions using branch traversal.
+One method can be used to track a quantitative variable across branches (\code{getBranchQuant()}) and the other method can be used to track a qualitative variables across branches (\code{getBranchQual()}).
+
+We can demonstrate the (\code{getBranchQuant()}) method by examining the yield variable across the descendant branches of the soybean variety ``A.K.".
+To understand more about the output of this function, please consult the \pkg{ggenealogy} package documentation.
+In the code below, we remove the output column ``DesNames" because it verbosely lists all descendant names, which is not necessary for this demonstration.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> AKBranches <- getBranchQuant("A.K.", sbGeneal, "yield", 15)
+R> select(AKBranches,-DesNames)
+\end{CodeInput}
+\begin{CodeOutput}
+           Name     Mean       SD Count NACount PercentNotNA
+1 A.K. (Harrow) 2932.154 197.0092    54      41        24.07
+2        Illini 2856.667 210.7801   131     125         4.58
+\end{CodeOutput}
+\end{Code}
+
+We see from the output that ``A.K." has two children named ``A.K.
+(Harrow)" and "Illini".
+Descendants from the ``A.K.
+(Harrow)" branch have a higher mean yield than the ``Illini" branch (2932.154 versus 2856.667).
+However, the
 

### 2016-11-21 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -1296,3 +1296,77 @@ dfToIG = function(geneal, vertexinfo = NULL, edgeweights = 1, isDirected=FALSE){
   
   igraph::graph.data.frame(d=edges, directed=isDirected, vertices=nodes)
 }
+
+#' Descendant branch calculations for quantitative variable 
+#' 
+#' Returns a data frame containing the names of all children of an individual of interest ("Name"). The mean and standard deviation ("Mean" and "SD") of a quantitative variable across all descendents of each child is reported. In addition, for each child, the number of its descendants is reported ("Count"), the number of its descendants who do not have a value for the quantitative variable ("NACount") is reported, and the names of all of its descendants is reported ("DesNames").
+#' 
+#' @param v1 the label of the vertex of interest (in character string format)
+#' @param geneal the full genealogy  (in data frame format)
+#' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
+#' @param gen the number of generations back to include as ancestors
+#' @examples
+#' data(statGeneal)
+#' DC_Year <- getBranchQuant("David Cox", statGeneal, "gradYear", 15)
+#' 
+#' @export
+getBranchQuant = function(v1, geneal, colName, gen=3){
+  id.offset <- NULL
+  if (is.null(getChild(v1, geneal))){
+    return(data.frame())
+  }
+  childList <- getChild(v1,geneal)
+  datRet <- data.frame()
+  for (i in 1:length(childList)){
+    dat <- getDescendants(childList[i], geneal, gen)
+    dat[[colName]] <- geneal[match(dat$label, geneal$child),][[colName]]
+    DesNames <- dat$label
+    Count <- length(DesNames)
+    Mean <- mean(dat[[colName]], na.rm=TRUE)
+    SD <- sd(dat[[colName]], na.rm = TRUE)
+    NACount <- sum(is.na(dat[[colName]]))
+    datRet <- rbind(datRet, data.frame(Name=childList[i], Mean=Mean, SD=SD, Count=Count, NACount=NACount, DesNames=paste(DesNames, collapse = ',')))
+  }
+  datRet <- datRet[order(-datRet$Mean),]
+  return(datRet)
+}
+
+#' Descendant branch calculations for quantitative variable 
+#' 
[…]

### 2016-11-21 JSS-ggenealogy/SourceCode/R/allFunctions.R
Source folder passes check()

@@ -683,22 +683,22 @@ getChild = function(v1, geneal){
 #' @param v2 the label of the second vertex of interest (in character string format)
 #' @param ig the graph representation of the data genealogy (in igraph format)
 #' @param geneal the full genealogy  (in data frame format)
-#' @param colName the name of the column of the data frame that contains the quantitative variable of interest (in character string format)
+#'
 #' @examples
 #' data(sbGeneal)
 #' ig <- dfToIG(sbGeneal)
 #' getDegree("Brim", "Bedford", ig, sbGeneal)
 #' @export
-getDegree = function(v1, v2, ig, geneal, colName){
+getDegree = function(v1, v2, ig, geneal){
   if(is.null(geneal)){
     stop("Please input a genealogy data frame")
   }
   if(is.null(ig)){
     stop("Please input an igraph object formatted by dfToIG()")
   }
-  path <- getPath(v1=v1, v2=v2, ig=ig, geneal = geneal, isDirected=F, colName = colName)
+  path <- getPathOnly(v1=v1, v2=v2, ig=ig, geneal = geneal, isDirected=F)
   # The degree between two vertices is equal to one less than the number of nodes in the shortest path
-  return(length(path$pathVertices)-1)
+  return(length(path)-1)
 }
 
 #' Returns the nodes for a full genealogy
@@ -833,6 +833,87 @@ getPath = function(v1, v2, ig, geneal, colName, silent=FALSE, isDirected=FALSE){
         }
       }
       retPath = list(pathVertices = pathVertices, variableVertices = variableVertices)
+    }
+  }
+  if(length(retPath)==0 & !silent){
+    message("Warning: There is no path between those two vertices")
+  }
+  # Return the shortest path, if it exists
+  retPath
+}
+
+
+#' Determine the path between two varieties
+#' 
+#' Determines the shortest path between the two inputted vertices, and takes into
+#' account whether or not the graph is directed. If there is a path, the list of vertices of the
+#' path will be returned. If there is not a path, a list of character(0) will be returned. Note:
+#' For a directed graph, the direction matters. However, this function will check both directions
+#' and return the path if it exists.
[…]

### 2016-11-22 article.tex
Branch additions to manuscript

--- 2016-11-17 article.tex
@@
 \end{Code}
+
+\section{Branch parsing and calculations}
+
+It may be helpful for users to search through descendant branches of a certain individual to compare and contrast how a variable of interest changes along those branches.
+For instance, which descending branches of a particular soybean variety are producing the highest yields?
+Which branches are developing new varieties in recent years?
+Which descending branches of a particular academic statistician have large proportions of students graduating from certain universities or countries?
+Which branches are graduating new students in recent years?
+Which branches have the highest proportion of thesis titles containing a word of interest?
+
+Answering these questions in a straightforward manner requires more than basic data frame manipulation: It also requires methods that can easily traverse parent-child relationships.
+The \pkg{ggenealogy} package has two methods that can easily answer these questions using branch traversal.
+One method can be used to track a quantitative variable across branches (\code{getBranchQuant()}) and the other method can be used to track a qualitative variable across branches (\code{getBranchQual()}).
+
+\subsection{Quantitative variable parsing and calculations}
+
+We can demonstrate the \code{getBranchQuant()} function by examining the quantitative variable ``yield" across the descendant branches of the soybean variety ``A.K.".
+To understand more about the output of this function, please consult the \pkg{ggenealogy} package documentation.
+In the code below, we remove the output column ``DesNames" because it verbosely lists all descendant names, which is not necessary for this demonstration.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> AKBranchYield <- getBranchQuant("A.K.", sbGeneal, "yield", 15)
+R> select(AKBranchYield,-DesNames)
+\end{CodeInput}
+\begin{CodeOutput}
+           Name     Mean       SD Count NACount
+1 A.K. (Harrow) 2932.154 197.0092    54      41
+2        Illini 2856.667 210.7801   131     125
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see from the output that ``A.K." has two children named ``A.K.
+(Harrow)" and ``Illini".
+Descendants from the ``A.K.
+(Harrow)" branch have a higher mean yield than the ``Illini" branch (2932.154 versus 2856.667).
+However, we should recognize that even though the branches contain a large number of descendants (54 and 131), most of these descendants did not include a yield value in the dataset (41 and 125).
+As a result, the mean values are calculated from a small proportion of the descendants.
+
+As another example, we can examine the mean graduation year for the ``descendant" branches of academic statistician ``David Cox".
+We know from earlier that ``David Cox" had 42 ``children", so as expected the \code{CoxBranchYear} object below contains 42 rows.
+However, only 8 of these rows have any ``descendants" of their own.
+As a result, only the first 8 rows of the object contain branch information.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> CoxBranchYear <- getBranchQuant("David Cox", statGeneal, "gradYear", 15)
+R> head(select(CoxBranchYear,-DesNames),10)
+\end{CodeInput}
+\begin{CodeOutput}
+                         Name     Mean        SD Count NACount
+1                 Mark Berman 2007.200  6.340347     5       0
+2                  Henry Wynn 2005.333  7.637626     3       0
+3                Rodney Wolff 2003.500  2.121320     2       0
+4                 Jane Hutton 2003.000        NA     1       0
+5              Gauss Cordeiro 2002.643  7.722167    14       0
+6             Peter McCullagh 2001.231  8.778645    26       0
+7             Basilio Pereira 2000.647 10.074356    17       0
+8            Peter Bloomfield 1999.918 11.707969    49       0
+9            Adelchi Azzalini      NaN        NA     0       0
+10 Amy Berrington de Gonzales      NaN        NA     0       0
+\end{CodeOutput}
+\end{CodeChunk}
+
+In this case, we see that of the 8 ``children" of David Cox who had ``children" of their own, ``Mark Berman" has the ``descendants" (n=5) who on average have graduated the most recently (2007.200), whereas ``Peter Bloomfield" has the ``descendants" (n=49) who on average have graduated the least recently (1999.918).
+We see that, for all branches, there are no ``descendants" who contain a NA value for graduation year.
+
+\subsection{Qualitative variable parsing and calculations}
+
+The \code{getBranchQual()} function requires similar inputs as the \code{getBranchQuant()} function, except that it also requires an input parameter called \code{rExpr}.
+The user must initialize this input parameter to a regular expression that can be applied to the column containing the qualitative variable of interest.
+The regular expression syntax must work on a data frame column of type character.
+It must be saved as a double quotation string, and any quotation marks within the regular expression must be single quotations.
+The term ``geneal\$colName" must be used in the regular expression.
+
+We can demonstrate the \code{getBranchQual()} function by examining the qualitative variable ``thesis" across the ``descendant" branches of the academic statistican ``David Cox".
+Since one of the primary research areas for ``David Cox" was stochastic processes, we can determine if any descendant branches of his ``children" contained thesis titles that included the word ``stochastic".
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> v1 = "David Cox"; geneal = statGeneal; colName = "thesis"; gen = 15
+R> rExpr = "grepl('(?i)Stochastic', geneal$colName)"
+R> CoxBranchStochastic <- getBranchQual(v1, geneal, colName, rExpr, gen)
+R> head(select(CoxBranchStochastic, -DesNames))
+\end{CodeInput}
+\begin{CodeOutput}
+                        Name CountTrue Count NACount
+1           Peter Bloomfield         4    49       0
+2            Basilio Pereira         1    17       0
+3           Adelchi Azzalini         0     0       0
+4 Amy Berrington de Gonzales         0     0       0
+5              Andrew Roddam         0     0       0
+6            Angela Mariotto         0     0       0
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see that only two ``children" of ``David Cox" had any ``descendants" with thesis titles containing the word ``Stochastic" (4 out of 49 ``descendants" of ``Peter Bloomfield" and 1 out of 17 ``descendants" of ``Basilio Periera").
+We see again that none of the ``descendants" from either branches contained values that were NA for the variable ``thesis".
+
+In many string parsing applications, the choice of the regular expression can be tricky.
+This is probably true when the string variable we are parsing is thesis titles.
+For instance, notice that in our regular expression, we accounted for all instances of the substring ``Stochastic".
+Hence, words that contain "Stochastic" as a substring (such as ``Stochastics" and ``Stochastically") will also be returned.
+In addition, we defined our regular expression to return matches whether the first letter was upper or lower case.
+When initializing the \code{rExpr} parameter, users would need to consider what variants of their search criteria they would like to define as matches.
+
+We will demonstrate one more example of the \code{getBranchQual()} function by searching the qualitative variable ``school" across the ``descendant" branches of the academic statistican ``David Cox".
+The Mathematics Genealogy Project coding system for the ``school" variable was non-ambiguous, and so we do not have to worry about various ways the same school could be named in the dataset.
+As a result, we can now simply use a regular expression that equates specifically to one value.
+
+It may be interesting to examine the school that is represented the most among all descendants of ``David Cox".
+To accomplish this, we use the \code{getDescendants()} function to create a data frame called \code{desDC} that contains the names of all 159 ``descendants" of ``David Cox".
+Then, we use the basic R function \code{match()} to match the school names from the original dataset to each of the 159 ``descendants" in the \code{desDC} data frame.
+We then use the basic R functions \code{sort()} and \code{table()} to examine the five schools that were represented the most out of the 159 ``descendants".
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> desDC <- getDescendants("David Cox", statGeneal, 15)
+R> tail(sort(table(statGeneal[match(desDC$label,statGeneal$child),]$school)),5)
+\end{CodeInput}
+\begin{CodeOutput}
+          The Johns Hopkins University 
+                                     9 
+Universidade Federal do Rio de Janeiro 
+                                    17 
+       North Carolina State University 
+                                    18 
+             Universidade de São Paulo 
+                                    28 
+                  University of London 
+                                    35
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see from this table that the most common school of the 159 ``descendants" of ``David Cox" was the University of London with a count of 35.
+We can now determine which of the 42 branches from the ``children" of ``David Cox" have the largest proportion of ``descendants" graduating from the University of London.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> colName = "school"
+R> rExpr = "geneal$colName=='University of London'" #works
+R> DCBranchUL <- getBranchQual(v1, geneal, colName, rExpr, gen)
+R> head(select(DCBranchUL, -DesNames))
+\end{CodeInput}
+\begin{CodeOutput}
+                        Name CountTrue Count NACount
+1            Peter McCullagh         1    26       0
+2           Adelchi Azzalini         0     0       0
+3 Amy Berrington de Gonzales         0     0       0
+4              Andrew Roddam         0     0       0
+5            Angela Mariotto         0     0       0
+6             Basil Springer         0     0       0
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see that ``Peter McCullagh" is the only ``child" of ``David Cox" that has a ``descendant" branch with one student graduating from the University of London; the rest of the 41 children of ``David Cox" have ``descendant" branches with zero students graduating from the University of London.
+This must mean the other 34 ``descendants" of ``David Cox" that graduated from the University of London were direct ``children" of ``David Cox".
+We can verify this below:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> sum(statGeneal[match(getChild("David Cox",statGeneal),statGeneal$child),]$school=="University of London")
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 34
+\end{CodeOutput}
+\end{CodeChunk}
+
+The examples above demonstrate that users can quickly and flexibly parse descendant branches.
+The swiftness comes from \pkg{ggenealogy} functions that allow for fast parent-child traversals, such as \code{getChild()}, \code{getDescendants()}, \code{getBranchQuant()}, and \code{getBranchQual()}.
+The flexibilty comes from data frame manipulation functions in base R that can be used in conjunction with the parent-child traversals.
 

### 2016-11-22 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -1412,6 +1412,7 @@ getBranchQuant = function(v1, geneal, colName, gen=3){
     datRet <- rbind(datRet, data.frame(Name=childList[i], Mean=Mean, SD=SD, Count=Count, NACount=NACount, DesNames=paste(DesNames, collapse = ',')))
   }
   datRet <- datRet[order(-datRet$Mean),]
+  rownames(datRet) <- 1:nrow(datRet)
   return(datRet)
 }
 
@@ -1452,5 +1453,6 @@ getBranchQual = function(v1, geneal, colName, rExpr, gen=3){
     datRet <- rbind(datRet, data.frame(Name=childList[i], CountTrue = CountTrue, Count=Count, NACount=NACount, DesNames=paste(DesNames, collapse = ',')))
   }
   datRet <- datRet[order(-datRet$CountTrue),]
+  rownames(datRet) <- 1:nrow(datRet)
   return(datRet)
 }

### 2016-11-22 article.tex
Revisions and formatting complete

--- 2016-11-20 article.tex
@@
 Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Subsection \ref{interaction}).
-
-\clearpage
 
@@
 R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
-R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
+R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) +
++    ggplot2::xlab("Development Year")
 \end{Code}
@@
 
-Furthermore, as determined by the figure, for both \code{Zane} and \code{Bedford}, there are four varieties of unidirectional parent-child relationships between each of them and their common ancestor \code{D55-4090}.
-Hence, any feature of interest that differentiates \code{Zane} and \code{Bedford} (protein content, yield, disease resistance, etc.) can also be examined across these two separate lineage histories.
+Furthermore, as seen in the figure, for both \code{Zane} and \code{Bedford}, there are four varieties of unidirectional parent-child relationships between each of them and their common ancestor \code{D55-4090}.
+Hence, any feature that differentiates \code{Zane} and \code{Bedford} (protein content, yield, disease resistance, etc.) can also be examined across these two separate lineage histories.
 
@@
 We illustrate this with the \code{sbGeneal} data frame after we subset it to remove observations that do not have values for the quantitative variable ``yield".
-Then, we can examine the path between the varieties \code{Clark} and \code{Lawrence} and plot how the quantitative varible ``yield" changes along the parent-child relationships of that path.
+Then, we can examine the path between the varieties \code{Clark} and \code{Lawrence} and plot how the quantitative variable ``yield" changes along the parent-child relationships of that path.
 Likewise, we plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path.
@@
 R> pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
-R> plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
+R> plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development
++    Year") + ggplot2::ylab("Yield")
 \end{Code}
@@
 \begin{Code}
-R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red",
-+    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
-+    element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
+R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol =
++    "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text =
++    ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size
++    = 12))
 \end{Code}
@@
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol =
-+    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development Year")
++    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development
++    Year")
 \end{Code}
@@
 We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created by subsetting the soy bean genealogy data to only include observations that contain a value for the quantitative variable ``yield".
-Then, we can examine the path across the entire genealogical structure and how the quantitative varible ``yield" changes along the parent-child relationships of that path.
+Then, we can examine the path across the entire genealogical structure and how the quantitative variable ``yield" changes along the parent-child relationships of that path.
 Likewise, we can plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path.
@@
 \begin{Code}
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol = "purple") + ggplot2::xlab("Yield")
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol =
++    "purple") + ggplot2::xlab("Yield")
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol =
++    "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
 \end{Code}
@@
 For explanation purposes, all paths between \code{Lee} and \code{5601T} were manually highlighted in blue.
+
+\clearpage
 
@@
 \end{figure}
+
+\clearpage
 
@@
 +    "gradYear", isDirected = FALSE)
-R> plotPath(pathCB, "gradYear", fontFace = 4) + ggplot2::xlab("Graduation Year") +
-+    ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title =
-+    ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
-+    = c(.1, .2))
+R> plotPath(pathCB, "gradYear", fontFace = 4) + ggplot2::xlab("Graduation
++    Year") + ggplot2::theme(axis.text = ggplot2::element_text(size = 10),
++    axis.title = ggplot2::element_text(size = 10)) +
++    ggplot2::scale_x_continuous(expand = c(.1, .2))
 \end{CodeInput}
 \end{CodeChunk}
+
+\clearpage
 
@@
 The output of this process is illustrated in Figure \ref{fig:plotCBText}.
+
+\clearpage
 
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize = .5,
-+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize =
++    .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
 +    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize = .5,
-+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize =
++    .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
 +    animate = TRUE)
@@
 \begin{Code}
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange", animate = TRUE)
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange", animate = TRUE)
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange",
++    animate = TRUE)
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol =
++    "orange", animate = TRUE)
 \end{Code}
 
-\section{Branch calculations and parsings}
+\section{Branch parsing and calculations}
 
-It may be helpful for users to search through the descendant branches of an individual of interest to compare and contrast how a variable of interest changes within these branches.
+It may be helpful for users to search through descendant branches of a certain individual to compare and contrast how a variable of interest changes along those branches.
 For instance, which descending branches of a particular soybean variety are producing the highest yields?
-Which are developing new varieties in recent years?
-Which descending branches of a particular academic statistician have large proportions of individuals graduating from a certain school or country?
-What propotion have thesis titles containing a word of interest?
-Which are graduating new students in recent years?
+Which branches are developing new varieties in recent years?
+Which descending branches of a particular academic statistician have large proportions of students graduating from certain universities or countries?
+Which branches are graduating new students in recent years?
+Which branches have the highest proportion of thesis titles containing a word of interest?
+
+Answering these questions in a straightforward manner requires more than basic data frame manipulation: It also requires methods that can easily traverse parent-child relationships.
+The \pkg{ggenealogy} package has two methods that can answer these questions using branch traversal.
+The\code{getBranchQuant()} function can be used to track a quantitative variable across branches and the \code{getBranchQual()} method can be used to track a qualitative variable across branches.
+
+\subsection{Quantitative variable parsing and calculations}
+
+We can demonstrate the \code{getBranchQuant()} function by examining the quantitative variable ``yield" across the descendant branches of the soybean variety \code{A.K.}.
+To understand more about the output of this function, please consult the \pkg{ggenealogy} package documentation.
+In the code below, we remove the output column ``DesNames" because it verbosely lists all descendant names, which is not necessary for this demonstration.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> AKBranchYield <- getBranchQuant("A.K.", sbGeneal, "yield", 15)
+R> select(AKBranchYield,-DesNames)
+\end{CodeInput}
+\begin{CodeOutput}
+           Name     Mean       SD Count NACount
+1 A.K. (Harrow) 2932.154 197.0092    54      41
+2        Illini 2856.667 210.7801   131     125
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see from the output that \code{A.K.} has two children named \code{A.K.
+(Harrow)} and \code{Illini}.
+Descendants from the \code{A.K.
+(Harrow)} branch have a higher mean yield than the \code{Illini} branch (2932.154 versus 2856.667).
+However, we should recognize that even though the branches contain a large number of descendants (54 and 131), most of these descendants did not come with a yield value (41 and 125).
+As a result, the mean values are calculated from a small proportion of the descendants.
+
+As another example, we can examine the mean graduation year for the ``descendant" branches of the academic statistician \code{David Cox}.
+We know from earlier that \code{David Cox} had 42 ``children", so as expected, the \code{CoxBranchYear} object below contains 42 rows.
+However, only 8 of these rows have any ``descendants" of their own.
+As a result, only the first 8 rows of the \code{CoxBranchYear} object contain branch information.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> CoxBranchYear <- getBranchQuant("David Cox", statGeneal, "gradYear", 15)
+R> head(select(CoxBranchYear,-DesNames),10)
+\end{CodeInput}
+\begin{CodeOutput}
+                         Name     Mean        SD Count NACount
+1                 Mark Berman 2007.200  6.340347     5       0
+2                  Henry Wynn 2005.333  7.637626     3       0
+3                Rodney Wolff 2003.500  2.121320     2       0
+4                 Jane Hutton 2003.000        NA     1       0
+5              Gauss Cordeiro 2002.643  7.722167    14       0
+6             Peter McCullagh 2001.231  8.778645    26       0
+7             Basilio Pereira 2000.647 10.074356    17       0
+8            Peter Bloomfield 1999.918 11.707969    49       0
+9            Adelchi Azzalini      NaN        NA     0       0
+10 Amy Berrington de Gonzales      NaN        NA     0       0
+\end{CodeOutput}
+\end{CodeChunk}
+
+In this case, we see that of the 8 ``children" of \code{David Cox} who had ``children" of their own, \code{Mark Berman} had the ``descendants" (n=5) who have on average graduated the most recently (2007.200), whereas \code{Peter Bloomfield} has the ``descendants" (n=49) who on average have graduated the least recently (1999.918).
+We see that, for all branches, there are no ``descendants" who contain a NA value for graduation year.
+
+\subsection{Qualitative variable parsing and calculations}
+
+The \code{getBranchQual()} function requires similar inputs as the \code{getBranchQuant()} function, except that it also requires an input parameter called \code{rExpr}.
+The user must initialize this input parameter to a regular expression that can be applied to the column containing the qualitative variable of interest.
+The regular expression syntax must work on a data frame column of type character.
+It must be saved as a double quotation string, and any quotation marks within it must be single quotations.
+The term ``geneal\$colName" must be used in the regular expression.
+
+We can demonstrate the \code{getBranchQual()} function by examining the qualitative variable ``thesis" across the ``descendant" branches of the academic statistician \code{David Cox}.
+Since one of the primary research areas for \code{David Cox} was stochastic processes, we can determine if any descendant branches of his ``children" contained thesis titles that included the word ``stochastic".
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> v1 = "David Cox"; geneal = statGeneal; colName = "thesis"; gen = 15
+R> rExpr = "grepl('(?i)Stochastic', geneal$colName)"
+R> CoxBranchStochastic <- getBranchQual(v1, geneal, colName, rExpr, gen)
+R> head(select(CoxBranchStochastic, -DesNames))
+\end{CodeInput}
+\begin{CodeOutput}
+                        Name CountTrue Count NACount
+1           Peter Bloomfield         4    49       0
+2            Basilio Pereira         1    17       0
+3           Adelchi Azzalini         0     0       0
+4 Amy Berrington de Gonzales         0     0       0
+5              Andrew Roddam         0     0       0
+6            Angela Mariotto         0     0       0
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see that only two ``children" of \code{David Cox} had any ``descendants" with thesis titles containing the word ``Stochastic" (4 out of 49 ``descendants" of \code{Peter Bloomfield} and 1 out of 17 ``descendants" of \code{Basilio Periera}).
+We see again that none of the ``descendants" from either branches contained values that were NA for the variable ``thesis".
+
+In many string parsing applications, the choice of the regular expression can be tricky.
+This is true when the string variable we are parsing is thesis titles.
+For instance, notice that in our regular expression, we accounted for all instances of the substring ``Stochastic".
+Hence, words that contain "Stochastic" (such as ``Stochastics" and ``Stochastically") will also be returned.
+In addition, we defined our regular expression to return matches whether the first letter was upper or lower case.
+When initializing the \code{rExpr} parameter, users would need to consider what nuances of their search criteria they would like to define as matches.
+
+We will demonstrate one more example of the \code{getBranchQual()} function by searching the qualitative variable ``school" across the ``descendant" branches of the academic statistician \code{David Cox}.
+The Mathematics Genealogy Project coding system for the ``school" variable was non-ambiguous, and so we do not have to worry about all the various ways the same school could be coded in the dataset.
+As a result, we now do not have to search for various substrings; we can simply use a regular expression that equates to one value.
+
+It may be interesting to examine the school that is represented the most among all descendants of \code{David Cox}.
+To determine what school this is, we use the \code{getDescendants()} function to create a data frame called \code{desDC} that contains the names of all 159 ``descendants" of \code{David Cox}.
+Then, we use the base R function \code{match()} to match the school names from the original dataset to each of the 159 ``descendants" in the \code{desDC} data frame.
+After that, we use the base R functions \code{sort()} and \code{table()} to examine the five schools that were represented the most throughout the 159 ``descendants".
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> desDC <- getDescendants("David Cox", statGeneal, 15)
+R> tableDC <- table(statGeneal[match(desDC$label, statGeneal$child), ]$school)
+R> tail(sort(tableDC), 5)
+\end{CodeInput}
+\begin{CodeOutput}
+          The Johns Hopkins University 
+                                     9 
+Universidade Federal do Rio de Janeiro 
+                                    17 
+       North Carolina State University 
+                                    18 
+             Universidade de São Paulo 
+                                    28 
+                  University of London 
+                                    35
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see from this table that the most common school of the 159 ``descendants" of \code{David Cox} was the University of London with a count of 35.
+We can now determine which of the branches from the 42 ``children" of \code{David Cox} have the largest proportion of ``descendants" graduating from the University of London.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> colName = "school"
+R> rExpr = "geneal$colName=='University of London'"
+R> DCBranchUL <- getBranchQual(v1, geneal, colName, rExpr, gen)
+R> head(select(DCBranchUL, -DesNames))
+\end{CodeInput}
+\begin{CodeOutput}
+                        Name CountTrue Count NACount
+1            Peter McCullagh         1    26       0
+2           Adelchi Azzalini         0     0       0
+3 Amy Berrington de Gonzales         0     0       0
+4              Andrew Roddam         0     0       0
+5            Angela Mariotto         0     0       0
+6             Basil Springer         0     0       0
+\end{CodeOutput}
+\end{CodeChunk}
+
+We see that \code{Peter McCullagh} is the only ``child" of \code{David Cox} that has a ``descendant" branch with one student graduating from the University of London; the rest of the 41 children of \code{David Cox} have ``descendant" branches with zero students graduating from the University of London.
+This must mean the other 34 ``descendants" of \code{David Cox} that graduated from the University of London were direct ``children" of \code{David Cox}.
+We can verify this below:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> DCChild <- statGeneal[match(getChild("David Cox", statGeneal),
++    statGeneal$child), ]
+R> sum(DCChild$school == "University of London")
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 34
+\end{CodeOutput}
+\end{CodeChunk}
+
+The examples above demonstrate that users can quickly and flexibly parse descendant branches.
+The swiftness comes from \pkg{ggenealogy} functions that allow for fast parent-child traversals, such as \code{getChild()}, \code{getDescendants()}, \code{getBranchQuant()}, and \code{getBranchQual()}.
+The flexibility comes from data frame manipulation functions in base R that can be used in conjunction with the parent-child traversals.
 

### 2016-11-22 article.tex
ReplicationScript testing

@@
 \begin{Code}
-plotPath(pathTN, sbGeneal, "devYear")
+R> plotPath(pathTN, sbGeneal, "devYear")
 \end{Code}
@@
 R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
-R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) +
-+    ggplot2::xlab("Development Year")
+R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab
++    ("Development Year")
 \end{Code}
@@
 +    "Pella", "Tokyo", "Young", "Zane")
->R plotDegMatrix(varieties, sbIG, sbGeneal, "devYear") +
+>R plotDegMatrix(varieties, sbIG, sbGeneal) +
 +    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
 +    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
-+    legend.text = ggplot2::element_text(size = 15)) +
-+    ggplot2::labs(x = "Variety", y = "Variety")
++    legend.text = ggplot2::element_text(size = 15)) + ggplot2::labs(x = 
++    "Variety", y = "Variety")
 \end{Code}
@@
 +    "gradYear", isDirected = FALSE)
-R> plotPath(pathCB, "gradYear", fontFace = 4) + ggplot2::xlab("Graduation
-+    Year") + ggplot2::theme(axis.text = ggplot2::element_text(size = 10),
-+    axis.title = ggplot2::element_text(size = 10)) +
-+    ggplot2::scale_x_continuous(expand = c(.1, .2))
+R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) +
++    ggplot2::xlab("Graduation Year") + ggplot2::theme(axis.text = 
++    ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(
++    (size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200)
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200) +
 +    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
@@
 The first call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis.
-When hovering over labels, the label name and yield value is revealed.
+When hovering over labels, the label name and yield value would be revealed.
 The second call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis and development year on the vertical axis.
-When hovering over labels, the label name, yield value, and development year is revealed.
+When hovering over labels, the label name, yield value, and development year would be revealed.
+We do not include the animated videos that the below code creates in this document, but readers can use the below code to create them on their own.
 
@@
 R> AKBranchYield <- getBranchQuant("A.K.", sbGeneal, "yield", 15)
-R> select(AKBranchYield,-DesNames)
+R> select(AKBranchYield, -DesNames)
 \end{CodeInput}
@@
 R> CoxBranchYear <- getBranchQuant("David Cox", statGeneal, "gradYear", 15)
-R> head(select(CoxBranchYear,-DesNames),10)
+R> head(select(CoxBranchYear, -DesNames), 10)
 \end{CodeInput}

### 2016-11-22 JSS-ggenealogy/ReplicationScript.R

@@ -33,31 +33,30 @@ getVariable("Zane", sbGeneal, "devYear")
 pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
 plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab("Development Year")
 
-# In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
-plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol = "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
-
-# We can repeat this process, only now instead of setting the bin variable to 1:3 (as we did earlier), we can set it to 1:6.
-plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol = "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development Year")
-
-##########################################################
-
+# Here, we demonstrate plotting paths on a filtered dataset, where all observations contain a non-NA value for yield. We show one example where there is one variable plotted for the path (yield) and another example where there are two variables plotted for the path (yield and devYear)
 sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
 sbFiltIG <- dfToIG(sbFilt)
 pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
 plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
 pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
 plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
+
+# In the previous section, we obtained the shortest path between the the pair of labels "Tokyo" and "Narow" and saved it as a variable pathTN. Here, we can plot that path superimposed over all labels in the example soybean genealogy dataset.
[…]

### 2016-11-22 article.tex
Formatting

@@
 \Abstract{
-  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} software package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, and displaying the results.
+  This paper introduces \pkg{ggenealogy} (\citealt{ggen}), a developing \proglang{R} software package that provides tools for searching through genealogical data, generating basic statistics on their graphical structures using parent and child connections, parsing and performing calculations on branches of interest, and displaying the results.
 It is possible to draw the genealogy in relation to variables related to the nodes, and to determine and display the shortest path distances between the nodes.
@@
 Most of these plotting tools were developed by building upon the \pkg{ggplot2} package (\citealt{ggplot2}).
-As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in demonstrative code throughout this paper.
+As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in example code throughout this paper.
 Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Subsection \ref{interaction}).
@@
 Each case in this quantitative variable column should be of type numeric.
-In the first example dataset, columns that could be used for this purpose (once subsetted to filter out values of ``'' and NA) include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear".
-At this point, any \pkg{ggenealogy} plot can be used.
+In the first example dataset, columns that could be used for this purpose include ``devYear", ``yield", and ``yearImputed"; in the second example dataset, the column that could be used for this purpose is ``gradYear".
+However, for these quantitative variable columns to successfully be used in plots, we would first need to assure that each row within them contains a numeric value (not NA).
+We could achieve this by filtering out or imputing certain rows that contain NA values for this column of interest.
+We demonstrate a filtering process for this purpose at the end of Section \ref{psp}.
+After that, any \pkg{ggenealogy} plot can be used.
 
@@
 
+\clearpage
+
 In the code above, \code{bin} was assigned a value of 3, and \code{pathEdgeCol} was assigned a value of ``red".
@@
 
+\clearpage
+
 We can immediately see that Figure \ref{fig:plotTNBin6} more successfully mitigates text overlap compared to Figure \ref{fig:plotTNBin3}.
@@
 We would like to emphasize that the \code{plotPathOnAll()} method can be created with one or two quantitative variables that the user hard-codes.
-We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created by subsetting the soy bean genealogy data to only include observations that contain a value for the quantitative variable ``yield".
-Then, we can examine the path across the entire genealogical structure and how the quantitative variable ``yield" changes along the parent-child relationships of that path.
+We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created to assure that there were no NA values in the quantitative variable ``yield".
+Then, we can examine the path across the remaining genealogical structure and how the quantitative variable ``yield" changes along the parent-child relationships of that path.
 Likewise, we can plot how the quantitative variables ``devYear" and ``yield" both change along the parent-child relationships of that path.
@@
 
-\clearpage
+The bottom plot of Figure \ref{fig:Lee} is not an output plot of \pkg{ggenealogy}.
+Instead, it was simply created for didactic purposes.
+Here, the paths that were manually highlighted in blue in the top plot produced by \pkg{ggenealogy} are shown again, only now nodes cannot be repeated.
+The parenthetical number above each node represents the set of generation counts distancing that node from the center node \code{Lee}; green parentheses indicate that the node could be successfully placed in one horizontal position, but red parentheses indicate that the node could not be successfully placed in one horizontal position.
+We see that node \code{TN89-39} cannot simultaneously be represented as both a third and fourth descendent of node \code{Lee}, and node \code{5601T} cannot simultaneously be represented as a third, fourth, and fifth descendent of node \code{Lee}.
+Hence, without allowing nodes to repeat, this dataset cannot be presented in the graph on the bottom as it can be in the \pkg{ggenealogy} graph on the top.
+This is a current limitation in other genealogy and graphical software that \pkg{ggenealogy} can now provide.
+
+\section{Plotting distance matrix}
+\label{pdm}
+
+It may also be of interest to generate matrices where the colors indicates a variable between all pairwise combinations of inputted varieties.
+The package \pkg{ggenealogy} also provides a function \code{plotDegMatrix()} for that purpose.
+We can demonstrate this function with the variable being the shortest path degree between a given pair of varieties.
+The shortest path degree is calculated as the smallest number of parent-child edges needed to traverse between two varieties of interest.
+For instance, in Figure \ref{fig:pathTNZB}, the shortest path degree between \code{Tokyo} and \code{Narow} is four and the shortest path degree between \code{Bedford} and \code{Zane} is ten.
 
@@
 
-The bottom plot of Figure \ref{fig:Lee} is not an output plot of \pkg{ggenealogy}.
-Instead, it was simply created for didactic purposes.
-Here, the paths that were manually highlighted in blue in the top plot produced by \pkg{ggenealogy} are shown again, only now nodes cannot be repeated.
-The parenthetical number above each node represents the set of generation counts distancing that node from the center node \code{Lee}; green parentheses indicate that the node could be successfully placed in one horizontal position, but red parentheses indicate that the node could not be successfully placed in one horizontal position.
-We see that node \code{TN89-39} cannot simultaneously be represented as both a third and fourth descendent of node \code{Lee}, and node \code{5601T} cannot simultaneously be represented as a third, fourth, and fifth descendent of node \code{Lee}.
-Hence, without allowing nodes to repeat, this dataset cannot be presented in the graph on the bottom as it can be in the \pkg{ggenealogy} graph on the top.
-This is a current limitation in other genealogy and graphical software that \pkg{ggenealogy} can now provide.
-
-\section{Plotting distance matrix}
-\label{pdm}
-
-It may also be of interest to generate matrices where the colors indicates a variable between all pairwise combinations of inputted varieties.
-The package \pkg{ggenealogy} also provides a function \code{plotDegMatrix()} for that purpose.
-We can demonstrate this function with the variable being the shortest path degree between a given pair of varieties.
-The shortest path degree is calculated as the smallest number of parent-child edges needed to traverse between two varieties of interest.
-For instance, in Figure \ref{fig:pathTNZB}, the shortest path degree between \code{Tokyo} and \code{Narow} is four and the shortest path degree between \code{Bedford} and \code{Zane} is ten.
-
 Here we generate a distance matrix for a set of 10 varieties, setting the x-label and y-label as ``Variety" and the legend label as ``Degree".
@@
 +    "Pella", "Tokyo", "Young", "Zane")
->R plotDegMatrix(varieties, sbIG, sbGeneal) +
-+    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
-+    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
-+    legend.text = ggplot2::element_text(size = 15)) + ggplot2::labs(x = 
-+    "Variety", y = "Variety")
+>R plotDegMatrix(varieties, sbIG, sbGeneal) + ggplot2::scale_fill_continuous
++    (low = "white", high = "darkgreen") + ggplot2::theme(legend.title =
++    ggplot2::element_text(size = 15), legend.text = ggplot2::element_text
++    (size = 15)) + ggplot2::labs(x = "Variety", y = "Variety")
 \end{Code}
@@
 \end{figure}
-
-\clearpage
 
@@
 
-We see from Figure \ref{fig:dCox} that Sir \code{David Cox} had 42 ``children", many of them becoming notable statisticians themselves, such as Basilio Pereira, Valerie Isham, Gauss Cordeiro, Peter McCullagh, and Henry Wynn.
-Of his ``children", the one who produced the most ``children" of their own was Peter Bloomfield, who has 26 ``children" and 49 ``descendants".
+We see from Figure \ref{fig:dCox} that Sir \code{David Cox} had 42 ``children", many of them becoming notable statisticians themselves, such as \code{Basilio Pereira}, \code{Valerie Isham}, \code{Gauss Cordeiro}, \code{Peter McCullagh}, and \code{Henry Wynn}.
+Of his ``children", the one who produced the most ``children" of their own was \code{Peter Bloomfield}, who has 26 ``children" and 49 ``descendants".
 In total, Sir \code{David Cox} had five generations of academic statistics mentees in this dataset.
@@
 +    "gradYear", isDirected = FALSE)
-R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) +
-+    ggplot2::xlab("Graduation Year") + ggplot2::theme(axis.text = 
-+    ggplot2::element_text(size = 10), axis.title = ggplot2::element_text(
-+    (size = 10)) + ggplot2::scale_x_continuous(expand = c(.1, .2))
-\end{CodeInput}
-\end{CodeChunk}
-
-\clearpage
+R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) + ggplot2::xlab
++    ("Graduation Year") + ggplot2::theme(axis.text = ggplot2::element_text
++    (size = 10), axis.title = ggplot2::element_text(size = 10)) +
++    ggplot2::scale_x_continuous(expand = c(.1, .2))
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-\clearpage
-
 \begin{CodeChunk}
@@
 
-\begin{figure}[H]
-    \centering
-    \fbox{\includegraphics[width=\textwidth]{plotCBNoText}}
-    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.
-Individuals on the shortest path are labeled in large and black text and connected by dark green edges; all other individuals are labeled in small and gray text and connected by light gray edges.}
-    \label{fig:plotCBNoText}
-\end{figure}
-
 In Figure \ref{fig:plotCBNoText}, we can now see each individual on the path of interest, and how their values for the variable year are overlaid on the entire genealogy structure.
@@
 This interactive functionality comes from methods in the \pkg{plotly} package (\citealt{plotly}).
-A short video demonstration of these interactive features can be viewed below.
+A short video demonstration of these interactive features can be viewed in Figure \ref{fig:plotAnimate}.
+
+\begin{figure}[H]
+    \centering
+    \fbox{\includegraphics[width=\textwidth]{plotCBNoText}}
+    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.
+Individuals on the shortest path are labeled in large and black text and connected by dark green edges; all other individuals are labeled in small and gray text and connected by light gray edges.}
+    \label{fig:plotCBNoText}
+\end{figure}
 
@@
 \end{CodeChunk}
+
+We would like to state that users can still hard code one or two variables in the \code{plotPathOnAll()} function, even with the \code{animate} option set to a value of ``TRUE".
+The first call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis.
+When hovering over labels, the label name and yield value would be revealed.
+The second call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis and development year on the vertical axis.
+When hovering over labels, the label name, yield value, and development year would be revealed.
+We do not include the animated videos that the below code creates in this document, but readers can use the below code to create them on their own.
+
+\begin{Code}
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange",
++    animate = TRUE)
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol =
++    "orange", animate = TRUE)
+\end{Code}
 
@@
 
-We would like to state that users can still hard code one or two variables in the \code{plotPathOnAll()} function, even with the \code{animate} option set to a value of ``TRUE".
-The first call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis.
-When hovering over labels, the label name and yield value would be revealed.
-The second call to the \code{plotPathOnAll()} function below would produce an interactive plot with the yield on the horizontal axis and development year on the vertical axis.
-When hovering over labels, the label name, yield value, and development year would be revealed.
-We do not include the animated videos that the below code creates in this document, but readers can use the below code to create them on their own.
-
-\begin{Code}
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange",
-+    animate = TRUE)
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol =
-+    "orange", animate = TRUE)
-\end{Code}
-
 \section{Branch parsing and calculations}
@@
 The \pkg{ggenealogy} package has two methods that can answer these questions using branch traversal.
-The\code{getBranchQuant()} function can be used to track a quantitative variable across branches and the \code{getBranchQual()} method can be used to track a qualitative variable across branches.
-
+The \code{getBranchQuant()} function can be used to track a quantitative variable across branches and the \code{getBranchQual()} method can be used to track a qualitative variable across branches.
+
+\clearpage
 \subsection{Quantitative variable parsing and calculations}
 
-We can demonstrate the \code{getBranchQuant()} function by examining the quantitative variable ``yield" across the descendant branches of the soybean variety \code{A.K.}.
-To understand more about the output of this function, please consult the \pkg{ggenealogy} package documentation.
+We can demonstrate the \code{getBranchQuant()} function by examining the quantitative variable ``yield" across the descendant branches of the soybean variety \code{A.K.} To understand more about the output of this function, please consult the \pkg{ggenealogy} package documentation.
 In the code below, we remove the output column ``DesNames" because it verbosely lists all descendant names, which is not necessary for this demonstration.
@@
 However, we should recognize that even though the branches contain a large number of descendants (54 and 131), most of these descendants did not come with a yield value (41 and 125).
-As a result, the mean values are calculated from a small proportion of the descendants.
+As a result, the mean values were calculated from a small proportion of the descendants.
 
@@
 
+\clearpage
 \subsection{Qualitative variable parsing and calculations}
 
-The \code{getBranchQual()} function requires similar inputs as the \code{getBranchQuant()} function, except that it also requires an input parameter called \code{rExpr}.
+The \code{getBranchQual()} function requires similar inputs as the \code{getBranchQuant()} function above, except that it also requires an input parameter called \code{rExpr}.
 The user must initialize this input parameter to a regular expression that can be applied to the column containing the qualitative variable of interest.
@@
 It must be saved as a double quotation string, and any quotation marks within it must be single quotations.
-The term ``geneal\$colName" must be used in the regular expression.
+The term \code{geneal\$colName} must be used in the regular expression.
 
@@
 The Mathematics Genealogy Project coding system for the ``school" variable was non-ambiguous, and so we do not have to worry about all the various ways the same school could be coded in the dataset.
-As a result, we now do not have to search for various substrings; we can simply use a regular expression that equates to one value.
+As a result, we no longer have to search for various substrings; we can simply use a regular expression that equates to one value.
 
@@
 To determine what school this is, we use the \code{getDescendants()} function to create a data frame called \code{desDC} that contains the names of all 159 ``descendants" of \code{David Cox}.
-Then, we use the base R function \code{match()} to match the school names from the original dataset to each of the 159 ``descendants" in the \code{desDC} data frame.
+Then, we use the base R function \code{match()} to match the school names from the original genealogy dataset to each of the 159 ``descendants" in the \code{desDC} data frame.
 After that, we use the base R functions \code{sort()} and \code{table()} to examine the five schools that were represented the most throughout the 159 ``descendants".
@@
 The swiftness comes from \pkg{ggenealogy} functions that allow for fast parent-child traversals, such as \code{getChild()}, \code{getDescendants()}, \code{getBranchQuant()}, and \code{getBranchQual()}.
-The flexibility comes from data frame manipulation functions in base R that can be used in conjunction with the parent-child traversals.
+The flexibility comes from data frame manipulation functions in base R that can be used in conjunction with the parent-child traversal methods.
 
@@
 
+\clearpage
 \section*{Acknowledgments}

### 2017-02-20 JSS-ggenealogy/testBranches.R
Editor update from 2/16/17

@@ -1,39 +1,0 @@
-rm(list=ls())
-library("devtools")
-load_all("SourceCode")
-library("ggenealogy")
-data("sbGeneal")
-data("statGeneal")
-library(stringr) # for str_count
-
-v1="Tokyo";geneal=sbGeneal;colName="devYear";gen=15
-v1="David Cox";geneal=statGeneal;colName="gradYear";gen=15
-
-TokyoBranches <- getBranchQuant("Tokyo", sbGeneal, "devYear", 15)
-AKBranches <- getBranchQuant("A.K.", sbGeneal, "yield", 15)
-CoxBranches <- getBranchQuant("David Cox", statGeneal, "gradYear", 15)
-
-##############################################################################
-##############################################################################
-
-v1="David Cox";geneal=statGeneal;colName="thesis";gen=15
-rExpr = "grepl('Stochastic', geneal$colName)"
-DCBranches_Sto <- getBranchQual(v1, geneal, colName, rExpr, gen)
-
-rExpr = "grepl('(?i)Stochastic', geneal$colName)"
-DCBranches_Stochastic <- getBranchQual(v1, geneal, colName, rExpr, gen)
-
-colName = "country"
-rExpr = "geneal$colName=='UnitedKingdom'" #works
-DCBranches_UK <- getBranchQual(v1, geneal, colName, rExpr, gen)
-
-desDC <- getDescendants("David Cox", statGeneal,15)
-tail(sort(table(statGeneal[match(desDC$label,statGeneal$child),]$school)),5)
-sum(statGeneal[match(getChild("David Cox",statGeneal),statGeneal$child),]$school=="University of London")
-
-colName = "school"
-rExpr = "geneal$colName=='University of London'" #works
-DCBranches_UL <- getBranchQual(v1, geneal, colName, rExpr, gen)
-
-rExpr = "geneal$colName=='Universidade de São Paulo'"
-DCBranches_USP <- getBranchQual(v1, geneal, colName, rExpr, gen)

### 2017-02-20 testBranches.R

@@ -21,7 +21,7 @@ rExpr = "grepl('Stochastic', geneal$colName)"
 DCBranches_Sto <- getBranchQual(v1, geneal, colName, rExpr, gen)
 
 rExpr = "grepl('(?i)Stochastic', geneal$colName)"
-DCBranches_StoCaps <- getBranchQual(v1, geneal, colName, rExpr, gen)
+DCBranches_Stochastic <- getBranchQual(v1, geneal, colName, rExpr, gen)
 
 colName = "country"
 rExpr = "geneal$colName=='UnitedKingdom'" #works

### 2017-04-15 article.tex
Editor update from 2/16/17
Editor code indentation corrections

@@
 As such, most of these plotting tools can be customized by appending syntax from the \pkg{ggplot2} package, as will be demonstrated in example code throughout this paper.
-Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Subsection \ref{interaction}).
+Moreover, some of these plotting tools have interactive capabilities that were developed with the use of the \pkg{plotly} package (\citealt{plotly}) (see Section \ref{interaction}).
 
@@
 R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
-R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab
-+    ("Development Year")
+R> plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) +
++    ggplot2::xlab("Development Year")
 \end{Code}
@@
 R> pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
-R> plotPath(pathCL2, sbFilt, "devYear", "yield") + ggplot2::xlab("Development
-+    Year") + ggplot2::ylab("Yield")
+R> plotPath(pathCL2, sbFilt, "devYear", "yield") +
++    ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
 \end{Code}
@@
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol =
-+    "red", nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text =
-+    ggplot2::element_text(size = 12), axis.title = ggplot2::element_text(size
-+    = 12))
++    "red", nodeSize = 2.5, pathNodeSize = 4) +
++    ggplot2::theme(axis.text = ggplot2::element_text(size = 12),
++    axis.title = ggplot2::element_text(size = 12))
 \end{Code}
@@
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol =
-+    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Development
-+    Year")
++    "seagreen2", nodeSize = 1, pathNodeSize = 3) +
++    ggplot2::xlab("Development Year")
 \end{Code}
@@
 +    "purple") + ggplot2::xlab("Yield")
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol =
-+    "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear",
++    pathEdgeCol = "orange") + ggplot2::xlab("Yield") +
++    ggplot2::ylab("Development Year")
 \end{Code}
@@
 +    "Pella", "Tokyo", "Young", "Zane")
->R plotDegMatrix(varieties, sbIG, sbGeneal) + ggplot2::scale_fill_continuous
-+    (low = "white", high = "darkgreen") + ggplot2::theme(legend.title =
-+    ggplot2::element_text(size = 15), legend.text = ggplot2::element_text
-+    (size = 15)) + ggplot2::labs(x = "Variety", y = "Variety")
+>R plotDegMatrix(varieties, sbIG, sbGeneal) +
++    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
++    ggplot2::theme(legend.title = ggplot2::element_text(size = 15),
++    legend.text = ggplot2::element_text(size = 15)) +
++    ggplot2::labs(x = "Variety", y = "Variety")
 \end{Code}
@@
 +    "gradYear", isDirected = FALSE)
-R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) + ggplot2::xlab
-+    ("Graduation Year") + ggplot2::theme(axis.text = ggplot2::element_text
-+    (size = 10), axis.title = ggplot2::element_text(size = 10)) +
+R> plotPath(pathCB, statGeneal, "gradYear", fontFace = 4) +
++    ggplot2::xlab("Graduation Year") + ggplot2::theme(axis.text =
++    ggplot2::element_text(size = 10), axis.title =
++    ggplot2::element_text(size = 10)) +
 +    ggplot2::scale_x_continuous(expand = c(.1, .2))
@@
 R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200) +
-+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
-+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-+    (expand = c(.1, .2)) + ggplot2::xlab("Graduation Year")
++    ggplot2::theme(axis.text = ggplot2::element_text(size = 8),
++    axis.title = ggplot2::element_text(size = 8)) +
++    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
++    ggplot2::xlab("Graduation Year")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize =
-+    .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
-+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
-+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
-+    (expand = c(.1, .2)) + ggplot2::xlab("Graduation Year")
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200,
++    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
++    "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = +    8), axis.title = ggplot2::element_text(size = 8)) +
++    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
++    ggplot2::xlab("Graduation Year")
 \end{CodeInput}
@@
 \begin{CodeInput}
-R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200, nodeSize =
-+    .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
-+    animate = TRUE)
+R> plotPathOnAll(pathCB, statGeneal, statIG, "gradYear", bin = 200,
++    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
++    "lightgray", animate = TRUE)
 \end{CodeInput}
@@
 +    animate = TRUE)
-R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol =
-+    "orange", animate = TRUE)
+R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear",
++    pathEdgeCol = "orange", animate = TRUE)
 \end{Code}
@@
 R> desDC <- getDescendants("David Cox", statGeneal, 15)
-R> tableDC <- table(statGeneal[match(desDC$label, statGeneal$child), ]$school)
+R> tableDC <- table(statGeneal[match(desDC$label,
++    statGeneal$child), ]$school)
 R> tail(sort(tableDC), 5)

### 2017-04-15 article.tex
Ready for resubmission

@@
 Users can obtain and plot the shortest path between two nodes of interest constrained on a variable of interest (see Section \ref{psp}), superimpose a shortest path between two nodes of interest with the entire genealogical structure constrained on a variable of interest (see Section \ref{sspt}), plot the ancestors and descendants of a node of interest constrained on generation count (see Section \ref{remedy}), and plot distance matrices based on a variable of interest (see Section \ref{pdm}).
-
 Most of these plotting tools were developed by building upon the \pkg{ggplot2} package (\citealt{ggplot2}).
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> plotPath(pathTN, sbGeneal, "devYear")
-\end{Code}
+\end{CodeInput}
 
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")
@@
 +    ggplot2::xlab("Development Year")
-\end{Code}
+\end{CodeInput}
 
@@
 
-\begin{Code}
-R> sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]),]
+\begin{CodeInput}
+R> sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]), ]
 R> sbFiltIG <- dfToIG(sbFilt)
@@
 +    ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
-\end{Code}
+\end{CodeInput}
 
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, pathEdgeCol =
@@
 +    axis.title = ggplot2::element_text(size = 12))
-\end{Code}
+\end{CodeInput}
 
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6, pathEdgeCol =
@@
 We would like to emphasize that the \code{plotPathOnAll()} method can be created with one or two quantitative variables that the user hard-codes.
-We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) that we previously created to assure that there were no NA values in the quantitative variable ``yield".
+We demonstrate this using the filtered data frame and igraph objects (\code{sbFilt} and \code{sbFiltIG}) to assure that there were no NA values in the quantitative variable ``yield".
 Then, we can examine the path across the remaining genealogical structure and how the quantitative variable ``yield" changes along the parent-child relationships of that path.
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol =
@@
 +    ggplot2::ylab("Development Year")
-\end{Code}
+\end{CodeInput}
 
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> plotAncDes("Lee", sbGeneal, mAnc = 6, mDes = 6, vCol = "blue")
-\end{Code}
+\end{CodeInput}
 
@@
 
-\begin{Code}
->R varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
+\begin{CodeInput}
+R> varieties <- c("Brim", "Bedford", "Calland", "Dillon", "Hood", "Narow",
 +    "Pella", "Tokyo", "Young", "Zane")
->R plotDegMatrix(varieties, sbIG, sbGeneal) +
+R> plotDegMatrix(varieties, sbIG, sbGeneal) +
 +    ggplot2::scale_fill_continuous(low = "white", high = "darkgreen") +
@@
 +    ggplot2::labs(x = "Variety", y = "Variety")
-\end{Code}
+\end{CodeInput}
 
@@
 \begin{CodeInput}
->R library("dplyr")
->R indVec <- getNodes(statGeneal)
->R indVec <- indVec[which(indVec != "", )]
->R dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
->R numDesc <- sapply(indVec, dFunc)
->R table(numDesc)
+R> library("dplyr")
+R> indVec <- getNodes(statGeneal)
+R> indVec <- indVec[which(indVec != "", )]
+R> dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
+R> numDesc <- sapply(indVec, dFunc)
+R> table(numDesc)
 \end{CodeInput}
@@
 
-\begin{CodeChunk}
 \begin{CodeInput}
@@
 \end{CodeInput}
-\end{CodeChunk}
 
@@
 
+\begin{figure}[H]
+    \centering
+    \fbox{\includegraphics[width=\textwidth]{plotCBText}}
+    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.}
+    \label{fig:plotCBText}
+\end{figure}
+
 We can achieve this using the below code.
@@
 \end{CodeInput}
-\end{CodeChunk}
-
-\begin{figure}[H]
-    \centering
-    \fbox{\includegraphics[width=\textwidth]{plotCBText}}
-    \caption{The shortest path between Sir \code{David Cox} and \code{Petra Buzkova}, superimposed over the data structure, using a bin size of 200.}
-    \label{fig:plotCBText}
-\end{figure}
 
@@
 +    nodeSize = .5, pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol =
-+    "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size = +    8), axis.title = ggplot2::element_text(size = 8)) +
++    "lightgray") + ggplot2::theme(axis.text = ggplot2::element_text(size =
++    8), axis.title = ggplot2::element_text(size = 8)) +
 +    ggplot2::scale_x_continuous(expand = c(.1, .2)) +
@@
 
-\begin{Code}
+\begin{CodeInput}
 R> plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", pathEdgeCol = "orange",
@@
 +    pathEdgeCol = "orange", animate = TRUE)
-\end{Code}
+\end{CodeInput}
 

### 2017-04-15 JSS-ggenealogy/ReplicationScript.R

@@ -1,7 +1,7 @@
 # Load and examine the structure of the example soybean genealogy dataset (called sbGeneal)
 rm(list=ls())
 library("devtools")
-load_all("SourceCode")
+load_all("Code")
 library("ggenealogy")
 data("sbGeneal")
 str(sbGeneal)

### 2017-04-15 JSS-ggenealogy/SourceCode/R/allFunctions.R

@@ -1,1458 +1,0 @@
-#' Returns the coordinate positions of all ancestors and descendants of a variety.
-#' 
-#' Calculates coordinates to plot each ancestors and descendant of a variety in a lineage. The x and y values describe the coordinates of the label, while the xstart, ystart, xend, and yend values describe the edges of the label.
-#' 
-#' @param df the data frame of the ancestors and descendants of a variety (from function buildAncDesTotalDF)
-#' @seealso \code{\link{buildAncList}} for information on determining ancestors
-#' @seealso \code{\link{buildDesList}} for information on determining descendants
-buildAncDesCoordDF = function(df){
-  id.offset <- NULL
-  root.gen <- gen <- par.id <- label <- NULL 
-  # This gets rid of redundancy and creates a "center"
-  if(nrow(subset(df, root.gen==0 & gen==0))>1){
-    temp = subset(df, root.gen==0 & gen==0)
-    temp$type = "center"
-    old.ids = temp$id[-1]
-    df$par.id[which(df$par.id%in%old.ids)] = temp$id[1]
-    temp$id[-1] = temp$id[1]
-    temp = unique(temp)
-    df = plyr::rbind.fill(subset(df, !(root.gen==0 & gen==0)), temp)
-  }
-  # This adds a leaf boolean column
-  df$leaf = !df$id%in%df$par.id
-  # Initialize x, y coords, and genside
-  df$genside = df$gen*(df$type=="descendant")-df$gen*(df$type=="ancestor")
-  df$x = 0
-  df$y = 0
-  # Ensure y coords are spread among all branches
-  df$y[df$leaf & df$genside<0] = seq(1, sum(df$leaf), length.out=sum(df$leaf & df$genside<0))
-  # but later, 2 ancestors are taken off b/c too previous
-  df$y[df$leaf & df$genside>0] = seq(1, sum(df$leaf), length.out=sum(df$leaf & df$genside>0))
-  # Ensure a single branch on one side is centrally placed (what if only 2 or 3?)
-  if(sum(df$leaf & df$genside<0)==1) df$y[df$leaf & df$genside<0] = sum(df$leaf)/2
-  if(sum(df$leaf & df$genside>0)==1) df$y[df$leaf & df$genside>0] = sum(df$leaf)/2
-  
-  df$xstart = 0
-  df$ystart = 0
-  df$xend = 0
-  df$yend = 0
-  df$branchx = 0
-  df$branchy = 0
[…]

### 2017-04-15 JSS-ggenealogy/SourceCode/R/data-sbGeneal.R

@@ -1,25 +1,0 @@
-#' Soybean genealogical data
-#' 
-#' This data contains information on copy number variants, single nucleotide polymorphisms, protein content, and yield, of soybeans. The available data consists of a data frame structure that contains 412 direct child-parent relationships between pairs of soybean varieties. These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-
-#' @name sbGeneal
-#' @title Soybean data
-#' @description This data set contains soy bean genealogical information maintained by the United States Department of Agriculture to be used by plant breeders, geneticists, bioinformaticians, pathologists,and many other research workers.
-#' @docType data
-#' @format a \code{RData} instance, 1 row per each child-parent relationship between soybean varieties
-#' @details \itemize{
-  #' \item child name of child soybean variety
-  #' \item devYear year child variety was introduced
-  #' \item yield protein yield
-  #' \item yearImputed whether or not the introduced year of the child variety was imputed
-  #' \item parent name of parent soybean variety
-  #' }
-  #'
-#' @docType data
-#' @keywords datasets
-#' @name sbGeneal
-#' @usage data(sbGeneal)
-#' @format A data frame with 412 rows and 7 variables
-#' @references
-#' "Pedigrees of soybean cultivars released in the United States and Canada." Theodore Hyivitz, C.A. Newell, S.G. Carmer. College of Agriculture, University of Illinois at Urbana-Champaign (1977).
-NULL

### 2017-04-15 JSS-ggenealogy/SourceCode/R/data-statGeneal.R

@@ -1,26 +1,0 @@
-#' Academic statistics genealogical data
-#' 
-#' This data contains information on copy number variants, single nucleotide polymorphisms, protein content, and yield, of soybeans. The available data consists of a data frame structure that contains 412 direct child-parent relationships between pairs of soybean varieties. These data were collected from field trials, genetic studies, and United States Department of Agriculture (USDA) bulletins, and date as early as the first decade of the 1900s.
-
-#' @name statGeneal
-#' @title Academic statistics data
-#' @description This data set contains academic genealogical information from the Mathematics Genealogy Project, a web-based database serviced by the North Dakota State University Department of Mathematics and the American Mathematical Society. Specifically, this data set represents a subset of the Mathematical Genealogy Project that contains all the parent-child relationships where both parent and child received an advanced degree of statistics as of June 6, 2015.
-#' @docType data
-#' @format a \code{RData} instance, 1 row per each individual in the Mathematics Genealogy Project with an advanced degree in statistics. If the child had a parent who was also listed as having received an advanced degree in statistics in the Mathematics Genealogy project, then the row also contains the parental information.
-#' @details \itemize{
-  #' \item child name of the individual who received an advanced degree in statistics
-  #' \item parent name of the individual who mentored the child and also received an advanced degree in statistics. If the child has no such parent, then this field is an empty string
-  #' \item gradYear year the child received their advanced degree in statistics
-  #' \item country country from which the child received their advanced degree in statistics
-  #' \item school school from which the child received their advanced degree in statistics
-  #' \item thesis title of the thesis the child submitted to receive their advanced degree in statistics. If this information is not available, then this field is an empty string
-#' }
-#'
-#' @docType data
-#' @keywords datasets
-#' @name statGeneal
-#' @usage data(statGeneal)
-#' @format A data frame with 8165 rows and 6 variables
-#' @references
-#' North Dakota State University and American Mathematical Society (2010). The Mathematics Genealogy Project. Archived Web Site. Retrieved from the Library of Congress, Accessed on March 6, 2015, URL http://www.genealogy.math.ndsu.nodak.edu.
-NULL
